### 2014-04-29 src/paper.tex
Initial commit

\documentclass[conference,a4paper]{IEEEtran}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{A secondary screen architecture to improve recommendations in DTV environments by capturing individual user interactions}

\author{\IEEEauthorblockN{Ricardo E. V. de S. Rosa}
\IEEEauthorblockA{Graduate Program in Electrical Engineering \\ Federal University of Minas Gerais \\ Av. Antônio Carlos 6627, 31270-901, \\Belo Horizonte, MG, Brazil\\
Email: ricardoerikson@ufmg.br}
\and
\IEEEauthorblockN{Vicente F. de Lucena Junior}
\IEEEauthorblockA{Graduate Program in Electrical Engineering \\
Federal University of Amazonas \\
Manaus, AM, Brazil\\
Email: vicente@ufam.edu.br}
}
\maketitle

\begin{abstract}
%\boldmath
The abstract goes here.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% no \IEEEPARstart
This demo file is intended to serve as a ``starter file''
for IEEE conference papers produced under \LaTeX\ using
IEEEtran.cls version 1.7 and later.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
I wish you the best of success.

\hfill mds

\hfill January 11, 2007

\subsection{Subsection Heading Here}
Subsection text here.

\subsubsection{Subsubsection Heading Here}
Subsubsection text here.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

\section{Conclusion}
The conclusion goes here \cite{IEEEexample:confwithvolume}.

% conference papers do not normally have an appendix

% use section* for acknowledgement
\section*{Acknowledgment}

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\bibliographystyle{IEEEtran}
\bibliography{IEEEexample}

% that's all folks
\end{document}

### 2014-04-30 src/paper.tex
Added sublime project

@@
 
+\usepackage{cite}
+
 \begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{A secondary screen architecture to improve recommendations in DTV environments by capturing individual user interactions}
+
+\title{A secondary screen architecture to accurately capture viewer interactions in interactive TV environments}
 
@@
 \begin{abstract}
-%\boldmath
-The abstract goes here.
+Watching TV is frequently seen as a social activity.
+Groups of people gather around the television for entertainment or information.
+One important application for TV viewers is content personalization, which consists in generating personalized content based on viewer interactions.
+Since the TV set is shared by many people, it is difficult to accurately capture the individual interactions from every viewer using a traditional remote control.
+As a new trend, the proliferation of personal mobile devices has changed the habits on audiovisual content consumption in interactive TV environments.
+In many situations people use their personal devices as a secondary screen to perform activities related to television watching.
+In this paper, we present an architecture for secondary screen devices that facilitates the capture of individual viewer interactions to use in content recommendation.
+In addition, this architecture can enhance the viewer interaction without interrupting the collective experience while watching TV.
 \end{abstract}
-% IEEEtran.cls defaults to using nonbold math in the Abstract.
-% This preserves the distinction between vectors and scalars. However,
-% if the conference you are submitting to favors bold math in the abstract,
-% then you can use LaTeX's standard command \boldmath at the very start
-% of the abstract to achieve this. Many IEEE journals/conferences frown on
-% math in the abstract anyway.
 
-% no keywords
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
 \IEEEpeerreviewmaketitle
@@
 \section{Introduction}
-% no \IEEEPARstart
-This demo file is intended to serve as a ``starter file''
-for IEEE conference papers produced under \LaTeX\ using
-IEEEtran.cls version 1.7 and later.
-% You must have at least 2 lines in the paragraph with the drop letter
-% (should never be an issue)
-I wish you the best of success.
 
-\hfill mds
+Watching TV is essentially a social activity, where family, friends, or people with a common interest share the same space and TV set for entertainment or information~\cite{Masthoff2004}.
+After many advances in technologies for interactive and digital TV it is possible to develop high level applications to enrich the TV experience.
+One important application is content personalization, in which content providers deliver personalized content to viewers based on their preferences.
 
-\hfill January 11, 2007
+The viewer preferences can be obtained by observing the viewing habits as interactions between the viewers and the TV system.
+Such interactions can be classified as implicit and explicit.
+Implicit interactions are part of the natural actions of watching TV e.g., change channel and volume.
+In explicit interactions viewers contribute by evaluating, commenting or sharing their opinion about a given content.
+Considering the data of thousands of viewers, the data captured can provide valuable information for content personalization services~\cite{Teixeira2010}.
 
-\subsection{Subsection Heading Here}
-Subsection text here.
+The traditional remote control as a mechanism of interaction between viewer and TV can present some problems.
+Since watching TV is a social and collective experience it is difficult to automatically identify the viewer in charge of the remote control.
+In addition, explicit interactions that require the use of the main screen of the TV set (e.g., evaluate the current content or read comments) ends up interfering in the collective experience.
+Some researchers tried to address this problem
 
-\subsubsection{Subsubsection Heading Here}
-Subsubsection text here.
+provide services which deliver personalized content to both individual users and groups of users.
+Such services depend on user preferences, which can be obtained by observing past viewers interactions with the TV system.
 
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation Results}
-%\label{fig_sim}
-%\end{figure}
+to perform secondary tasks while watching TV
 
-% Note that IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
+These tasks are frequently related to television watching e.g., evaluating content according to personal tastes, sharing, or even recommending content to other people.
+This paper presents an architecture that
 
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command, the
-% \label for the overall figure must come after \caption.
-% \hfil must be used as a separator to get equal spacing.
-% The subfigure.sty package works much the same way, except \subfigure is
-% used instead of \subfloat.
-%
-%\begin{figure*}[!t]
-%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
-%\label{fig_second_case}}}
-%\caption{Simulation results}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
+TV family behaviors towards television.
+potential of mobile devices to capture individual user interactions.
 
-% An example of a floating table. Note that, for IEEE style tables, the 
-% \caption command should come BEFORE the table. Table text will default to
-% \footnotesize as IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
+opens up new possibilities to accurately capture individual interactions from TV users. by using secondary screen devices.
 
-% Note that IEEE does not put floats in the very first column - or typically
-% anywhere on the first page for that matter. Also, in-text middle ("here")
-% positioning is not used. Most IEEE journals/conferences use top floats
-% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
-% command of the stfloats package.
+\bibliographystyle{IEEEtran}
+\bibliography{biblio.bib}
 
-\section{Conclusion}
-The conclusion goes here \cite{IEEEexample:confwithvolume}.
-
-% conference papers do not normally have an appendix
-
-% use section* for acknowledgement
-\section*{Acknowledgment}
-
-The authors would like to thank...
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-\bibliographystyle{IEEEtran}
-\bibliography{IEEEexample}
-
-% that's all folks
 \end{document}

### 2014-04-30 src/paper.tex
Paper updated

@@
 
-The traditional remote control as a mechanism of interaction between viewer and TV can present some problems.
-Since watching TV is a social and collective experience it is difficult to automatically identify the viewer in charge of the remote control.
-In addition, explicit interactions that require the use of the main screen of the TV set (e.g., evaluate the current content or read comments) ends up interfering in the collective experience.
-Some researchers tried to address this problem
+The traditional remote control as a mechanism of interaction between viewer and TV can present notable problems.It is difficult to automatically identify the viewer in charge of the remote control.
+Thus, in practice the captured interactions are assigned to the whole group, rather than to individual viewers.
+As a result, in recommender systems applications for example, the tastes of the user in charge of the remote control can be imposed on the tastes of others.
 
-provide services which deliver personalized content to both individual users and groups of users.
-Such services depend on user preferences, which can be obtained by observing past viewers interactions with the TV system.
+Considering the collective experience of watching TV, the explicit interactions of one viewer can potencialy interfere in the experience of other viewers.
+Since there is only one remote control and a unique screen (TV screen), explicit interactions that require the use this screen (e.g., evaluate the current content or read viewers comments) end up in an undesired experience for users that are not actively participating of that interactions.
+
+The proliferation of personal mobile devices has changed the people behaviors towards TV.
+Frequently people perform secondary tasks while watching TV such as use of email, use of social networks or surf the web.
+Many times these tasks are related to television watching where the viewer can evaluate, share or even recommend content to other people.
+Following this trend, more and more TV providers are enabling the audience to use personal devices as an extension of TV, i.e., as a secondary screen to enrich and immerse the user into the TV experience.
+
+Identifying the viewers and capturing their interactions in TV environments are challenging activities~\cite{Hwang2007,Teixeira2010}.
 
@@
 TV family behaviors towards television.
-potential of mobile devices to capture individual user interactions.
+potential use of mobile devices to capture individual user interactions.
 
 opens up new possibilities to accurately capture individual interactions from TV users. by using secondary screen devices.
+
+\section{Architecture}
+
+\section{Conclusions}
 

### 2014-04-30 src/paper.tex
WIP: Added related work in introduction

@@
 
-The traditional remote control as a mechanism of interaction between viewer and TV can present notable problems.It is difficult to automatically identify the viewer in charge of the remote control.
+The traditional remote control as a mechanism of interaction between viewer and TV can present notable problems.
+It is difficult to automatically identify the viewer in charge of the remote control.
 Thus, in practice the captured interactions are assigned to the whole group, rather than to individual viewers.
@@
 The proliferation of personal mobile devices has changed the people behaviors towards TV.
-Frequently people perform secondary tasks while watching TV such as use of email, use of social networks or surf the web.
-Many times these tasks are related to television watching where the viewer can evaluate, share or even recommend content to other people.
-Following this trend, more and more TV providers are enabling the audience to use personal devices as an extension of TV, i.e., as a secondary screen to enrich and immerse the user into the TV experience.
+Frequently people perform secondary tasks while watching TV such as use of email and social networks or surf the web.
+Many times these tasks are related to television so that the viewer can evaluate, share or even recommend content to other people.
+Following this trend, TV providers are enabling the audience to use personal devices as an extension of TV, i.e., as a secondary screen to enrich and immerse the user into the TV experience.
 
-Identifying the viewers and capturing their interactions in TV environments are challenging activities~\cite{Hwang2007,Teixeira2010}.
+In this scenario, secondary screen devices arise as a powerful mechanism to accurately capture viewer interactions.
+Other works addressed this separately: identifying viewers~\cite{Cabarcos2011,Hwang2007} and capturing interactions~\cite{Teixeira2010}.
+In this paper, we present a mechanism that uses
 
-to perform secondary tasks while watching TV
+ without interfering in the viewing experience of other people in the TV environment
 
-These tasks are frequently related to television watching e.g., evaluating content according to personal tastes, sharing, or even recommending content to other people.
-This paper presents an architecture that
-
-TV family behaviors towards television.
 potential use of mobile devices to capture individual user interactions.

### 2014-05-01 src/paper.tex
Introduction first version finished

@@
 
-Considering the collective experience of watching TV, the explicit interactions of one viewer can potencialy interfere in the experience of other viewers.
+Considering the collective experience of watching TV, the explicit interactions of one viewer can potentially interfere in the experience of other viewers.
 Since there is only one remote control and a unique screen (TV screen), explicit interactions that require the use this screen (e.g., evaluate the current content or read viewers comments) end up in an undesired experience for users that are not actively participating of that interactions.
@@
 
-In this scenario, secondary screen devices arise as a powerful mechanism to accurately capture viewer interactions.
+In this scenario, secondary screen devices arise as a powerful mechanism to accurately identify viewers and  capture their interactions.
 Other works addressed this separately: identifying viewers~\cite{Cabarcos2011,Hwang2007} and capturing interactions~\cite{Teixeira2010}.
-In this paper, we present a mechanism that uses
-
- without interfering in the viewing experience of other people in the TV environment
-
-potential use of mobile devices to capture individual user interactions.
-
-opens up new possibilities to accurately capture individual interactions from TV users. by using secondary screen devices.
+In this paper, we present an architecture that enables both identifying viewers and capture their interactions.
+In addition, the architecture described in this paper enables viewers to use the screen of their personal devices to perform explicit interactions without interfering in the experience of the audience in the TV environment.
 

### 2014-05-02 src/paper.tex
Abstract rewriting

@@
 \begin{abstract}
-Watching TV is frequently seen as a social activity.
-Groups of people gather around the television for entertainment or information.
-One important application for TV viewers is content personalization, which consists in generating personalized content based on viewer interactions.
+TV watching is frequently seen as a social activity.
+People gather around the television for entertainment or information.
+Advances in TV technology have enabled the viewer to actively interact with applications and providers instead of just passively watch TV.
+One important application for the viewers is content personalization, which consists in generating content recommendations based on viewers' interactions.
 Since the TV set is shared by many people, it is difficult to accurately capture the individual interactions from every viewer using a traditional remote control.
-As a new trend, the proliferation of personal mobile devices has changed the habits on audiovisual content consumption in interactive TV environments.
-In many situations people use their personal devices as a secondary screen to perform activities related to television watching.
-In this paper, we present an architecture for secondary screen devices that facilitates the capture of individual viewer interactions to use in content recommendation.
-In addition, this architecture can enhance the viewer interaction without interrupting the collective experience while watching TV.
+In this paper, we present an architecture that facilitates the identification of viewers and the capture of their individual interactions in shared TV environments.
+In addition, we present an area of application of this architecture in the recommender systems domain.
 \end{abstract}
@@
 
-Watching TV is essentially a social activity, where family, friends, or people with a common interest share the same space and TV set for entertainment or information~\cite{Masthoff2004}.
+TV watching is essentially a social activity, where family, friends, or people with a common interest share the same space and TV set for entertainment or information~\cite{Masthoff2004}.
 After many advances in technologies for interactive and digital TV it is possible to develop high level applications to enrich the TV experience.
@@
 Considering the data of thousands of viewers, the data captured can provide valuable information for content personalization services~\cite{Teixeira2010}.
+
+people rarely just sit passively and watch TV
+
+As a new trend, the proliferation of personal mobile devices has changed the habits on audiovisual content consumption in interactive TV environments.
+In many situations people use their personal devices as a secondary screen to perform activities related to television watching.
+
+In addition, this architecture can enhance the viewer interaction without interrupting the collective experience while watching TV.
 

### 2014-05-02 src/paper.tex
Rewrited introduction section

@@
 
-\title{A secondary screen architecture to accurately capture viewer interactions in interactive TV environments}
+\title{A secondary screen architecture to accurately capture TV viewer interactions for content personalization in TV environments}
 
@@
 In this paper, we present an architecture that facilitates the identification of viewers and the capture of their individual interactions in shared TV environments.
-In addition, we present an area of application of this architecture in the recommender systems domain.
 \end{abstract}
@@
 After many advances in technologies for interactive and digital TV it is possible to develop high level applications to enrich the TV experience.
-One important application is content personalization, in which content providers deliver personalized content to viewers based on their preferences.
+Thus, the interaction between viewers and TV became more elaborated than simply change channel and volume.
 
-The viewer preferences can be obtained by observing the viewing habits as interactions between the viewers and the TV system.
-Such interactions can be classified as implicit and explicit.
+The interactions between viewers and TV can be classified as implicit and explicit.
 Implicit interactions are part of the natural actions of watching TV e.g., change channel and volume.
-In explicit interactions viewers contribute by evaluating, commenting or sharing their opinion about a given content.
-Considering the data of thousands of viewers, the data captured can provide valuable information for content personalization services~\cite{Teixeira2010}.
+In explicit interactions viewers contribute by actively evaluating, commenting or sharing their opinion about a given content.
+Both kinds of interaction are important for applications that rely on audience generated data.
 
-people rarely just sit passively and watch TV
+Considering the interactions of millions of viewers, the captured data can provide valuable information~\cite{Teixeira2010}.
+For example, the content producers can obtain the user ratings given to TV programs, amount of viewers and user profiles.
+One important application that makes use of viewers interaction is content personalization.
+Content providers can infer user profiles from viewers' interactions in order to deliver personalized content.
 
-As a new trend, the proliferation of personal mobile devices has changed the habits on audiovisual content consumption in interactive TV environments.
-In many situations people use their personal devices as a secondary screen to perform activities related to television watching.
+The traditional remote control (RC) is the input device that viewers typically use to interact with the TV.
+However, the RC presents two notable problems for capturing interactions from viewers: (1) to automatically identify the viewers in the TV environment; and (2) to individually capture interactions of the viewer in charge of the RC.
+Thus, considering a scenario with the traditional RC, all interactions are assigned to the whole group of viewers (e.g., a family), rather than to the viewer in charge of the RC.
+As a result, the tastes that were inferred from the users in charge of the RC can be imposed on the tastes of others.
 
-In addition, this architecture can enhance the viewer interaction without interrupting the collective experience while watching TV.
+As a potential mean to handle the limitations of the RC for capturing interactions from viewers, the proliferation of personal mobile devices has changed the people behavior towards audiovisual content consumption in interactive TV environments.
+Nowadays, viewers rarely just sit passively and watch TV.
+In many situations they use their personal devices as a secondary screen to perform activities related to TV watching, which end up enriching and immersing the user into the TV experience.
 
-The traditional remote control as a mechanism of interaction between viewer and TV can present notable problems.
-It is difficult to automatically identify the viewer in charge of the remote control.
-Thus, in practice the captured interactions are assigned to the whole group, rather than to individual viewers.
-As a result, in recommender systems applications for example, the tastes of the user in charge of the remote control can be imposed on the tastes of others.
-
-Considering the collective experience of watching TV, the explicit interactions of one viewer can potentially interfere in the experience of other viewers.
-Since there is only one remote control and a unique screen (TV screen), explicit interactions that require the use this screen (e.g., evaluate the current content or read viewers comments) end up in an undesired experience for users that are not actively participating of that interactions.
-
-The proliferation of personal mobile devices has changed the people behaviors towards TV.
-Frequently people perform secondary tasks while watching TV such as use of email and social networks or surf the web.
-Many times these tasks are related to television so that the viewer can evaluate, share or even recommend content to other people.
-Following this trend, TV providers are enabling the audience to use personal devices as an extension of TV, i.e., as a secondary screen to enrich and immerse the user into the TV experience.
-
-In this scenario, secondary screen devices arise as a powerful mechanism to accurately identify viewers and  capture their interactions.
-Other works addressed this separately: identifying viewers~\cite{Cabarcos2011,Hwang2007} and capturing interactions~\cite{Teixeira2010}.
-In this paper, we present an architecture that enables both identifying viewers and capture their interactions.
-In addition, the architecture described in this paper enables viewers to use the screen of their personal devices to perform explicit interactions without interfering in the experience of the audience in the TV environment.
+Secondary screen devices arise as a powerful mechanism to accurately identify viewers and  capture their interactions.
+In this research field, a typical approach is to use personal devices (e.g., smartphones and tablets) as secondary screens~\cite{Courtois2012}, since they are almost ubiquitous and present good graphical capabilities and Internet connection.
+The connectivity capabilities of smartphones were used in~\cite{Cabarcos2011}, where the Bluetooth technology was used to detect nearby users.
+With respect to interaction capture, an approach described in~\cite{Teixeira2010} shown an architecture to capture contextualized viewers interactions through a conventional remote control.
+In this paper, we present an architecture that enables identifying viewers and capturing their interactions using a personal device as secondary screen.
 

### 2014-05-02 src/paper.tex
WIP: architecture

@@
 
+This architecture consists of a secondary screen device (which might be implemented in personal devices such as smartphones and tablets) a set-top box (STB) with return channel, and a content provider (which is responsible for providing the interactive applications and for storing the data that was captured from viewers).
+
+The secondary screen is used to identify the users that are present in the environment.
+That is, the viewer can authenticate to the local set-top box by using a transparent authentication mechanism.
+Thus, every time a viewer is nearby the STB using a a TV application with its
+
+The second screen devices are Internet capable devices, which have TV related applications. The viewers use these applications to interact with the TV by evaluating, commenting or sharing TV content.
+
+The STB is synchronized with the TV in order to contextualize the data that is captured.
+\cite{Lee2010}
+
 \section{Conclusions}

### 2014-05-02 src/paper.tex
WIP: Architecture

@@
 \usepackage{cite}
+\usepackage{graphicx}
 
@@
 Considering the interactions of millions of viewers, the captured data can provide valuable information~\cite{Teixeira2010}.
-For example, the content producers can obtain the user ratings given to TV programs, amount of viewers and user profiles.
+For example, the content providers (CP) can obtain the user ratings given to TV programs, amount of viewers and user profiles.
 One important application that makes use of viewers interaction is content personalization.
@@
 
-This architecture consists of a secondary screen device (which might be implemented in personal devices such as smartphones and tablets) a set-top box (STB) with return channel, and a content provider (which is responsible for providing the interactive applications and for storing the data that was captured from viewers).
+This architecture consists of secondary screen devices (which might be implemented in personal devices such as smartphones and tablets) a set-top box (STB) with return channel, and a CP (which is responsible for providing the interactive applications and for storing the data that is captured from viewers).
 
-The secondary screen is used to identify the users that are present in the environment.
-That is, the viewer can authenticate to the local set-top box by using a transparent authentication mechanism.
-Thus, every time a viewer is nearby the STB using a a TV application with its
+Initially, the CP must provide an infrastructure for viewers and the STB so that they can be authenticated to use the application services.
+In addition, the infrastructure must also provide data storage capabilities to store the interaction data captured from viewers.
+Then the STB must be authenticated in the application hosted on the CP infrastructure.
+Once the STB is authenticated, the viewers in the TV environment must also authenticate in the application service and register the STB in their accounts.
+As a result, the viewer can use their personal device as a secondary screen for that STB.
+
+Figure~\ref{fig_architecture} Green arrows represent data flowing from the viewer to the CP infrastructure, while blue arrows represent the flow from CP to viewer.
+The black arrows represent data flow inside the CP infrastructure.
+
+The secondary screen device must be used in conjunction with the STB to identify the users that are present in the environment and update context information. In the first place, the STB
+
+That is, the viewer can authenticate to the local set-top box by using a transparent authentication mechanism. Thus, every time a viewer is nearby the STB using a a TV application with its
 
 The second screen devices are Internet capable devices, which have TV related applications. The viewers use these applications to interact with the TV by evaluating, commenting or sharing TV content.
+
+\begin{figure}[!t]
+	\centering
+	\includegraphics[width=3.5in]{img/architecture.pdf}
+	\caption{Architecture to capture viewers insteractions. }
+	\label{fig_architecture}
+\end{figure}
 
@@
 
+Since applications in the TV domain have potential to deal with millions of users at the same time, it is interesting that this infrastructure has the ability to handle a growing amount of data~\cite{Lee2010}.
+
 \bibliographystyle{IEEEtran}

### 2014-05-04 src/paper.tex
Unfinished version of the paper.

--- 2014-05-02 src/paper.tex
@@
 
-\title{A secondary screen architecture to accurately capture TV viewer interactions for content personalization in TV environments}
+\title{A secondary screen architecture to accurately capture TV viewers' interactions and contextual information in TV environments}
 
@@
 Considering the interactions of millions of viewers, the captured data can provide valuable information~\cite{Teixeira2010}.
-For example, the content producers can obtain the user ratings given to TV programs, amount of viewers and user profiles.
+For example, the content providers (CPs) can obtain the user ratings given to TV programs, amount of viewers and user profiles.
+
 One important application that makes use of viewers interaction is content personalization.
@@
 
+Initially, the CP infrastructure must provide an authentication mechanism so that viewers and STBs can be authenticated to use the secondary screen services.
+The STB must be authenticated in the application hosted on the CP infrastructure.
+Once the STB is authenticated, the viewers in the TV environment must link their user accounts with the STB.
+As a result, the viewer can use their personal device as a secondary screen for that STB, enabling the capture of viewers' interactions.
+
+The architecture to accurately capture viewers interactions is shown in Figure~\ref{fig_architecture}.
+This architecture consists of secondary screen devices (which might be implemented in personal devices such as smartphones and tablets) a set-top box (STB), and the CP infrastructure (which includes an application hosting infrastructure).
+The green arrows represent the data flowing from the viewer to the CP infrastructure, while blue arrows denote the flow from CP infrastructure to the viewers.
+The black arrows represent the data flow inside the CP infrastructure.
+
+\begin{figure}[!t]
+	\centering
+	\includegraphics[width=3.5in]{img/architecture.pdf}
+	\caption{Architecture to capture audience data. The viewers' interactions and the contextual information are captured and stored in the CP infrastructure.}
+	\label{fig_architecture}
+\end{figure}
+
+The STB has a presence service which detects nearby devices with viewer accounts linked to the STB.
+Once the STB and a linked secondary screen device are in the same environment, they are synchronized in order to update contextual data about viewer profile and amount of people using a secondary device in the room.
+This contextual information can be used to learn how a viewer behave toward other viewers.
+For example, a viewer behaves in a particular way while watching TV alone and the same viewer can behave in a different way while watching in group.
+
+Viewers can use their secondary screen device to interact with TV by evaluating, commenting, sharing or even recommending TV content to other viewers.
+These interactions and the contextual information are represented by the audience data that is then sent to CP infrastructure through Internet.
+Then, a Data Writer service stores the audience data related to each individual viewer and STB device that are authenticated in the TV environment.
+
+Once the audience data is stored, the CP can use this data to derive valuable information about users' tastes.
+This information can be used to provide personalized content to viewers based on the current context.
+Thus, the personalized content can be delivered via Internet to the secondary screen device or even to the TV depending on the context.
+
 \section{Conclusions}
 
+This is paper presents an architecture to accurately capture audience data in interactive TV by using secondary screen devices.
+The implementation of this architecture is a work in progress and its main goal is to capture viewers' interactions and contextual information to use in applications for interactive TV environments.
+Potential applications for this data are in the field of content personalization for TV.
+
+Since applications in the TV domain have potential to deal with millions of users at the same time, it is interesting that this CP infrastructure has the ability to handle a growing amount of data~\cite{Lee2010,Kim2012,Shin2009}.
+The data from millions of viewers obtained from the implementation of this architecture can provide valuable information about viewers.
+To obtain this information, CPs can employ machine learning algorithms or other techniques on the raw audience data aiming to create viewer profiles for content personalization.
+As a result, CPs can improve their services aiming to increase the audience or enrich user experience.
+
 \bibliographystyle{IEEEtran}
-\bibliography{biblio.bib}
+\bibliography{biblio}
 

### 2014-05-05 src/paper.tex
Finished the conclusion

@@
 The implementation of this architecture is a work in progress and its main goal is to capture viewers' interactions and contextual information to use in applications for interactive TV environments.
-Potential applications for this data are in the field of content personalization for TV.
 
-Since applications in the TV domain have potential to deal with millions of users at the same time, it is interesting that this CP infrastructure has the ability to handle a growing amount of data~\cite{Lee2010,Kim2012,Shin2009}.
-The data from millions of viewers obtained from the implementation of this architecture can provide valuable information about viewers.
-To obtain this information, CPs can employ machine learning algorithms or other techniques on the raw audience data aiming to create viewer profiles for content personalization.
-As a result, CPs can improve their services aiming to increase the audience or enrich user experience.
+Since applications in the TV domain have the potential to reach out to millions of users at the same time, it is interesting that the CP infrastructure has the ability to deal with a large amount of incoming data~\cite{Lee2010,Kim2012,Shin2009}.
+The captured data is individual and when it is associated with the context it can provide valuable information about viewers.
+By using proper machine learning algorithms this data can provide valuable information about viewers.
+Knowing this information can be very useful for CPs to improve their services aiming to increase the audience and enrich user experience.
+As an example, CPs can deliver personalized content for a single person as well as for groups of people by using content recommendation algorithms.
+
+The use of personal devices as secondary screens arises as a new trend in TV environments.
+The architecture described in this paper take advantage of the ubiquity and computational capabilities of secondary screen devices to identify viewers and provide feedback for CPs through viewers' interactions and contextual information.
 

### 2014-05-05 src/paper.tex
Updated abstract

@@
 
-\title{A secondary screen architecture to accurately capture TV viewers' interactions and contextual information in TV environments}
+\title{A secondary screen architecture to accurately capture viewers' interactions and contextual information in TV environments}
 
@@
 People gather around the television for entertainment or information.
-Advances in TV technology have enabled the viewer to actively interact with applications and providers instead of just passively watch TV.
-One important application for the viewers is content personalization, which consists in generating content recommendations based on viewers' interactions.
-Since the TV set is shared by many people, it is difficult to accurately capture the individual interactions from every viewer using a traditional remote control.
-In this paper, we present an architecture that facilitates the identification of viewers and the capture of their individual interactions in shared TV environments.
+Advances in TV technology have enabled the viewer to actively interact with the TV through interactive applications instead of just passively watch TV.
+Since the TV set is shared by many people and the interaction occur by using a single remote control, it is difficult to accurately capture the individual interactions from every viewer.
+It is also difficult to identify the context in which the interactions occur, e.g., the people that are present in the environment.
+In this paper, we present an architecture that facilitates viewers' individual interactions and contextual information in shared TV environments.
+One important application of this data consists in generating personalized content for single viewers and groups of viewers.
 \end{abstract}

### 2014-05-05 src/paper.tex
First version for revision

@@
 In this paper, we present an architecture that facilitates viewers' individual interactions and contextual information in shared TV environments.
-One important application of this data consists in generating personalized content for single viewers and groups of viewers.
+One important application for this data consists in generating personalized content for single viewers and groups of viewers.
 \end{abstract}
@@
 TV watching is essentially a social activity, where family, friends, or people with a common interest share the same space and TV set for entertainment or information~\cite{Masthoff2004}.
-After many advances in technologies for interactive and digital TV it is possible to develop high level applications to enrich the TV experience.
+After many advances in technologies for interactive and digital TV it is possible to develop high level applications to enrich the watching experience.
 Thus, the interaction between viewers and TV became more elaborated than simply change channel and volume.
@@
 Implicit interactions are part of the natural actions of watching TV e.g., change channel and volume.
-In explicit interactions viewers contribute by actively evaluating, commenting or sharing their opinion about a given content.
+In explicit interactions, viewers contribute by actively evaluating, commenting or sharing their opinion about a given content.
 Both kinds of interaction are important for applications that rely on audience generated data.
 
-Considering the interactions of millions of viewers, the captured data can provide valuable information~\cite{Teixeira2010}.
-For example, the content providers (CPs) can obtain the user ratings given to TV programs, amount of viewers and user profiles.
+Considering the interactions of millions of viewers, the captured data can provide valuable information for content providers (CPs)~\cite{Teixeira2010}.
+For example, the CPs can obtain the user ratings given to TV programs, amount of viewers and user profiles.
+By using proper machine learning algorithms, CPs can infer viewer tastes from audience data in order to deliver personalized content~\cite{Kim2012,Shin2009}.
+However, identifying viewers and capturing their individual interactions is still a challenging activity in the traditional TV environment, which consists of the TV and viewers interacting through a remote control (RC).
 
-One important application that makes use of viewers interaction is content personalization.
-Content providers can infer user profiles from viewers' interactions in order to deliver personalized content.
-
-The traditional remote control (RC) is the input device that viewers typically use to interact with the TV.
-However, the RC presents two notable problems for capturing interactions from viewers: (1) to automatically identify the viewers in the TV environment; and (2) to individually capture interactions of the viewer in charge of the RC.
-Thus, considering a scenario with the traditional RC, all interactions are assigned to the whole group of viewers (e.g., a family), rather than to the viewer in charge of the RC.
+The conventional RC is the input device that viewers typically use to interact with the TV.
+However, the RC presents two notable problems for capturing interactions from viewers: (1) to automatically identify the viewers that are in the TV environment in order to contextualize the interaction; and (2) to individually capture interactions from the viewer in charge of the RC.
+Thus, considering a scenario with the traditional TV environment and a conventional TV, all interactions are assigned to the whole group of viewers (e.g., a family), rather than to the viewer in charge of the RC.
 As a result, the tastes that were inferred from the users in charge of the RC can be imposed on the tastes of others.
 
-As a potential mean to handle the limitations of the RC for capturing interactions from viewers, the proliferation of personal mobile devices has changed the people behavior towards audiovisual content consumption in interactive TV environments.
+As a potential alternative to handle the limitations of the RC for capturing interactions from viewers, the proliferation of personal mobile devices has changed the people behavior towards audiovisual content consumption in interactive TV environments.
 Nowadays, viewers rarely just sit passively and watch TV.
-In many situations they use their personal devices as a secondary screen to perform activities related to TV watching, which end up enriching and immersing the user into the TV experience.
+In many situations, they use their personal devices as a secondary screen to perform activities related to TV watching, which end up enriching and immersing the user into the TV experience.
 
 Secondary screen devices arise as a powerful mechanism to accurately identify viewers and  capture their interactions.
-In this research field, a typical approach is to use personal devices (e.g., smartphones and tablets) as secondary screens~\cite{Courtois2012}, since they are almost ubiquitous and present good graphical capabilities and Internet connection.
+In this research field, a typical approach is to use personal devices (e.g., smartphones and tablets) as secondary screens~\cite{Courtois2012}, since they are almost ubiquitous and present good graphical capabilities.
 The connectivity capabilities of smartphones were used in~\cite{Cabarcos2011}, where the Bluetooth technology was used to detect nearby users.
-With respect to interaction capture, an approach described in~\cite{Teixeira2010} shown an architecture to capture contextualized viewers interactions through a conventional remote control.
-In this paper, we present an architecture that enables identifying viewers and capturing their interactions using a personal device as secondary screen.
+With respect to interaction capture, an approach described in~\cite{Teixeira2010} shown an architecture to capture contextualized viewers interactions through a conventional RC.
+In this paper, we present an architecture that facilitates the identification of viewers, which is useful to contextualize the interactions.
+In addition, the identification of viewers facilitates the capture of their individual interactions.
 
@@
 
-Since applications in the TV domain have the potential to reach out to millions of users at the same time, it is interesting that the CP infrastructure has the ability to deal with a large amount of incoming data~\cite{Lee2010,Kim2012,Shin2009}.
+Since applications in the TV domain have the potential to reach out to millions of users at the same time, it is interesting that the CP infrastructure has the ability to deal with a large amount of incoming data~\cite{Lee2010}.
 The captured data is individual and when it is associated with the context it can provide valuable information about viewers.

### 2014-05-14 src/paper.tex
Changes in title and abstract

@@
 
-\title{A secondary screen architecture to accurately capture viewers' interactions and contextual information in TV environments}
+\title{A secondary screen architecture to accurately capture viewers' interactions in an iTV environment}
 
@@
 Advances in TV technology have enabled the viewer to actively interact with the TV through interactive applications instead of just passively watch TV.
-Since the TV set is shared by many people and the interaction occur by using a single remote control, it is difficult to accurately capture the individual interactions from every viewer.
-It is also difficult to identify the context in which the interactions occur, e.g., the people that are present in the environment.
+Since the TV set is shared by many people and the interaction occur by using a single remote control, it is difficult to accurately capture the individual needs from each viewer.
+It is also difficult to identify the context in which the interactions occur, e.g., to manage uniquely who is present in the environment and what is being watched on TV.
 In this paper, we present an architecture that facilitates viewers' individual interactions and contextual information in shared TV environments.

### 2014-05-14 src/paper.tex
Updated introduction with the revision from Prof. Vicente

@@
 After many advances in technologies for interactive and digital TV it is possible to develop high level applications to enrich the watching experience.
-Thus, the interaction between viewers and TV became more elaborated than simply change channel and volume.
+Thus, the interaction between viewers and TV became more elaborated than simply change channels and adjust sound volume.
 
 The interactions between viewers and TV can be classified as implicit and explicit.
-Implicit interactions are part of the natural actions of watching TV e.g., change channel and volume.
+Implicit interactions are part of the natural actions of watching TV e.g., changing channels.
 In explicit interactions, viewers contribute by actively evaluating, commenting or sharing their opinion about a given content.
@@
 Considering the interactions of millions of viewers, the captured data can provide valuable information for content providers (CPs)~\cite{Teixeira2010}.
-For example, the CPs can obtain the user ratings given to TV programs, amount of viewers and user profiles.
-By using proper machine learning algorithms, CPs can infer viewer tastes from audience data in order to deliver personalized content~\cite{Kim2012,Shin2009}.
+For example, the CPs can obtain the user ratings given to TV programs, amount of viewers, and user profiles.
+By using proper machine learning algorithms, CPs can infer viewers' tastes from audience data in order to deliver personalized content~\cite{Kim2012,Shin2009}.
 However, identifying viewers and capturing their individual interactions is still a challenging activity in the traditional TV environment, which consists of the TV and viewers interacting through a remote control (RC).
 
-The conventional RC is the input device that viewers typically use to interact with the TV.
-However, the RC presents two notable problems for capturing interactions from viewers: (1) to automatically identify the viewers that are in the TV environment in order to contextualize the interaction; and (2) to individually capture interactions from the viewer in charge of the RC.
+The conventional RC presents two notable problems for capturing interactions from viewers: (1) to automatically identify the viewers that are in the TV environment in order to contextualize the interaction; and (2) to individually capture interactions from the viewer in charge of the RC.
 Thus, considering a scenario with the traditional TV environment and a conventional TV, all interactions are assigned to the whole group of viewers (e.g., a family), rather than to the viewer in charge of the RC.
-As a result, the tastes that were inferred from the users in charge of the RC can be imposed on the tastes of others.
+As a result, the tastes that were inferred from the users in charge of the RC can be imposed on the tastes of the others.
 
-As a potential alternative to handle the limitations of the RC for capturing interactions from viewers, the proliferation of personal mobile devices has changed the people behavior towards audiovisual content consumption in interactive TV environments.
+As a potential alternative to handle the limitations of the RC, the proliferation of personal mobile devices has changed the people behavior towards audiovisual content consumption in interactive TV environments.
 Nowadays, viewers rarely just sit passively and watch TV.

### 2014-05-15 src/paper.tex
WIP: changes in figure and itv device definition

@@
 
-Initially, the CP infrastructure must provide an authentication mechanism so that viewers and STBs can be authenticated to use the secondary screen services.
-The STB must be authenticated in the application hosted on the CP infrastructure.
-Once the STB is authenticated, the viewers in the TV environment must link their user accounts with the STB.
-As a result, the viewer can use their personal device as a secondary screen for that STB, enabling the capture of viewers' interactions.
+Initially, the CP infrastructure must provide an authentication mechanism so that viewers and their interactive TV (iTV) devices (e.g., set-top boxes or smart TVs) can be authenticated by the CP in order to use the secondary screen services.
+Once one iTV device is authenticated, the viewers in the TV environment must link their user accounts with the iTV device.
+As a result, the viewer can use their personal device as a secondary screen for that iTV device, enabling the capture of viewers' interactions.
 
 The architecture to accurately capture viewers interactions is shown in Figure~\ref{fig_architecture}.
-This architecture consists of secondary screen devices (which might be implemented in personal devices such as smartphones and tablets) a set-top box (STB), and the CP infrastructure (which includes an application hosting infrastructure).
+This architecture consists of secondary screen devices (which might be implemented in personal devices such as smartphones and tablets) an iTV device, and the CP infrastructure (which includes an application hosting infrastructure).
 The green arrows represent the data flowing from the viewer to the CP infrastructure, while blue arrows denote the flow from CP infrastructure to the viewers.

### 2014-05-17 src/paper.tex
Finished text corrections and added Lucas as author

@@
 \and
-\IEEEauthorblockN{Vicente F. de Lucena Junior}
+\IEEEauthorblockN{Lucas C. Cordeiro and Vicente F. de Lucena Junior}
 \IEEEauthorblockA{Graduate Program in Electrical Engineering \\
@@
 
-TV watching is essentially a social activity, where family, friends, or people with a common interest share the same space and TV set for entertainment or information~\cite{Masthoff2004}.
+TV watching is essentially a social activity, where family, friends, or people with a common interest share the same space and TV set for entertainment or information.
 After many advances in technologies for interactive and digital TV it is possible to develop high level applications to enrich the watching experience.
@@
 Initially, the CP infrastructure must provide an authentication mechanism so that viewers and their interactive TV (iTV) devices (e.g., set-top boxes or smart TVs) can be authenticated by the CP in order to use the secondary screen services.
-Once one iTV device is authenticated, the viewers in the TV environment must link their user accounts with the iTV device.
+Once one iTV device is authenticated, the viewers in the TV environment must link their user accounts up with the iTV device.
 As a result, the viewer can use their personal device as a secondary screen for that iTV device, enabling the capture of viewers' interactions.
 
-The architecture to accurately capture viewers interactions is shown in Figure~\ref{fig_architecture}.
+The architecture to accurately capture viewers interactions is shown in \figurename~\ref{fig_architecture}.
 This architecture consists of secondary screen devices (which might be implemented in personal devices such as smartphones and tablets) an iTV device, and the CP infrastructure (which includes an application hosting infrastructure).
@@
 
-The STB has a presence service which detects nearby devices with viewer accounts linked to the STB.
-Once the STB and a linked secondary screen device are in the same environment, they are synchronized in order to update contextual data about viewer profile and amount of people using a secondary device in the room.
+A presence service detects nearby secondary screen devices that have the services necessary for establishing a communication with iTV device.
+Once the secondary screen device is linked up with the iTV device in the same environment, they are synchronized in order to update contextual data about viewer profile and amount of people using a secondary device in the room.
 This contextual information can be used to learn how a viewer behave toward other viewers.
@@
 These interactions and the contextual information are represented by the audience data that is then sent to CP infrastructure through Internet.
-Then, a Data Writer service stores the audience data related to each individual viewer and STB device that are authenticated in the TV environment.
+Then, a Data Writer service stores the audience data related to each particular viewer and the iTV device.
 
@@
 Since applications in the TV domain have the potential to reach out to millions of users at the same time, it is interesting that the CP infrastructure has the ability to deal with a large amount of incoming data~\cite{Lee2010}.
-The captured data is individual and when it is associated with the context it can provide valuable information about viewers.
+The captured data relates to a single viewer in a given context.
 By using proper machine learning algorithms this data can provide valuable information about viewers.
@@
 The use of personal devices as secondary screens arises as a new trend in TV environments.
-The architecture described in this paper take advantage of the ubiquity and computational capabilities of secondary screen devices to identify viewers and provide feedback for CPs through viewers' interactions and contextual information.
+The architecture described in this paper take advantage of the ubiquity and interactive capabilities of secondary screen devices to identify viewers and provide feedback for CPs through viewers' interactions and contextual information.
 

### 2014-05-17 src/paper.tex
Revision finished

@@
 This architecture consists of secondary screen devices (which might be implemented in personal devices such as smartphones and tablets) an iTV device, and the CP infrastructure (which includes an application hosting infrastructure).
-The green arrows represent the data flowing from the viewer to the CP infrastructure, while blue arrows denote the flow from CP infrastructure to the viewers.
-The black arrows represent the data flow inside the CP infrastructure.
+The dashed green arrows represent the data flowing from the viewer to the CP infrastructure, while dotted blue arrows denote the flow from CP infrastructure to the viewers.
+The continuous black arrows represent the data flow inside the CP infrastructure.
 
@@
 
-A presence service detects nearby secondary screen devices that have the services necessary for establishing a communication with iTV device.
-Once the secondary screen device is linked up with the iTV device in the same environment, they are synchronized in order to update contextual data about viewer profile and amount of people using a secondary device in the room.
-This contextual information can be used to learn how a viewer behave toward other viewers.
+A presence service detects physically nearby secondary screen devices that have the services necessary for establishing a communication with iTV device.
+Modern commercial operating systems for personal devices enable service discovery by using peer-to-peer or WiFi communication.
+This feature is particularly appropriate for the TV environment, once the secondary screen device must be wirelessly linked up with the iTV device.
+The devices are synchronized in order to update contextual data about viewer profile and amount of people using a secondary device in the room.
+This contextual data can be used to learn how a viewer behave toward other viewers.
 For example, a viewer behaves in a particular way while watching TV alone and the same viewer can behave in a different way while watching in group.
@@
 Viewers can use their secondary screen device to interact with TV by evaluating, commenting, sharing or even recommending TV content to other viewers.
-These interactions and the contextual information are represented by the audience data that is then sent to CP infrastructure through Internet.
+The data related to those actions are naturally provided by the viewers as part of the interaction when using interactive applications (developed by the CP) installed in their secondary screen devices.
+The interactions and the contextual data are represented by the audience data, which is sent to CP infrastructure through Internet.
 Then, a Data Writer service stores the audience data related to each particular viewer and the iTV device.
@@
 
-This is paper presents an architecture to accurately capture audience data in interactive TV by using secondary screen devices.
-The implementation of this architecture is a work in progress and its main goal is to capture viewers' interactions and contextual information to use in applications for interactive TV environments.
+The use of personal devices as secondary screens arises as a new trend in TV environments.
+This is paper presents an architecture to accurately capture audience data by using secondary screen devices and iTV devices.
+The implementation of this architecture is a work in progress, and take advantage of the ubiquity and interactive capabilities of secondary screen devices to identify viewers and provide feedback for CPs through viewers' interactions and contextual information.
 
@@
 
-The use of personal devices as secondary screens arises as a new trend in TV environments.
-The architecture described in this paper take advantage of the ubiquity and interactive capabilities of secondary screen devices to identify viewers and provide feedback for CPs through viewers' interactions and contextual information.
+The architecture described in this paper is also technically feasible to implement in Digital TV (DTV) systems and can be particularly interesting for CPs in DTV standards like the Brazilian (ISDB-TB) and the Japanese(ISDB-T).
+These standards support the use of mobile devices as part of the system.
+So, since the mobile devices support the implementation of the DTV middleware, a cooperative exhibition of the DTV application can be managed with the iTV device, thus enabling the capture of viewers' interactions and contextual data.
 

### 2014-05-17 src/paper.tex
minor changes in abstract

@@
 People gather around the television for entertainment or information.
-Advances in TV technology have enabled the viewer to actively interact with the TV through interactive applications instead of just passively watch TV.
-Since the TV set is shared by many people and the interaction occur by using a single remote control, it is difficult to accurately capture the individual needs from each viewer.
+Advances in TV technology have enabled the viewer to actively interact with the TV through interactive applications instead of just passively watching TV.
+Since the TV set is shared by many people and the interaction occurs by using a single remote control, it is difficult to accurately capture the individual needs from each viewer.
 It is also difficult to identify the context in which the interactions occur, e.g., to manage uniquely who is present in the environment and what is being watched on TV.
-In this paper, we present an architecture that facilitates viewers' individual interactions and contextual information in shared TV environments.
-One important application for this data consists in generating personalized content for single viewers and groups of viewers.
+In this paper, we present a novel architecture that facilitates viewers' individual interactions and contextual information in shared TV environments.
+One important application for this data consists of generating personalized content for single viewers and groups of viewers.
 \end{abstract}

### 2014-05-17 src/paper.tex
minor changes to the introduction

@@
 TV watching is essentially a social activity, where family, friends, or people with a common interest share the same space and TV set for entertainment or information.
-After many advances in technologies for interactive and digital TV it is possible to develop high level applications to enrich the watching experience.
+After many advances in technologies, for interactive and digital TV, it is possible to develop high-level applications to enrich the watching experience.
 Thus, the interaction between viewers and TV became more elaborated than simply change channels and adjust sound volume.
@@
 The interactions between viewers and TV can be classified as implicit and explicit.
-Implicit interactions are part of the natural actions of watching TV e.g., changing channels.
-In explicit interactions, viewers contribute by actively evaluating, commenting or sharing their opinion about a given content.
+Implicit interactions are part of the natural actions of watching TV (e.g., changing channels).
+In explicit interactions, viewers contribute by actively evaluating, commenting, or sharing their opinion about a given content.
 Both kinds of interaction are important for applications that rely on audience generated data.
@@
 By using proper machine learning algorithms, CPs can infer viewers' tastes from audience data in order to deliver personalized content~\cite{Kim2012,Shin2009}.
-However, identifying viewers and capturing their individual interactions is still a challenging activity in the traditional TV environment, which consists of the TV and viewers interacting through a remote control (RC).
+However, identifying viewers and capturing their individual interactions are still a challenging activity in the traditional TV environment, which consists of the TV and viewers interacting through a remote control (RC).
 
@@
 The connectivity capabilities of smartphones were used in~\cite{Cabarcos2011}, where the Bluetooth technology was used to detect nearby users.
-With respect to interaction capture, an approach described in~\cite{Teixeira2010} shown an architecture to capture contextualized viewers interactions through a conventional RC.
-In this paper, we present an architecture that facilitates the identification of viewers, which is useful to contextualize the interactions.
+With respect to interaction capture, an approach described in~\cite{Teixeira2010} has shown an architecture to capture contextualized viewers interactions through a conventional RC.
+In this paper, we present a novel architecture that facilitates the identification of viewers, which is useful to contextualize the interactions.
 In addition, the identification of viewers facilitates the capture of their individual interactions.

### 2014-05-17 src/paper.tex
comments related to the introduction

@@
 \usepackage{graphicx}
+
+\newcommand{\blurb}[1]{\marginpar{{\tt !!!}}{\tt [... #1 ...]}}
 
@@
 
+\blurb{Achei que o inicio da introducao esta com muito bla bla bla.
+Tente ir direto ao ponto, dando uma visao geral da sua abordagem, as vantagens e desvantages com as outras abordagens existentes na literatura e as principais contribuicoes do seu trabalho.
+Por que voce acredita que os revisores devem aceitar o seu trabalho?
+Qual eh a inovacao?
+Quais sao as contribuicoes?
+Quais sao as diferencas para os trabalhos que ja existem.
+Diminua o bla bla bla do inicio e tente responder estas perguntas na introducao do seu trabalho.}
+
 \section{Architecture}

### 2014-05-19 src/paper.tex
comments about the architecture

@@
 
+\blurb{Tente colocar um nome mais sugestivo para esta secao, em vez ``Architecture'' somente.}
+
 Initially, the CP infrastructure must provide an authentication mechanism so that viewers and their interactive TV (iTV) devices (e.g., set-top boxes or smart TVs) can be authenticated by the CP in order to use the secondary screen services.
 Once one iTV device is authenticated, the viewers in the TV environment must link their user accounts up with the iTV device.
-As a result, the viewer can use their personal device as a secondary screen for that iTV device, enabling the capture of viewers' interactions.
+As a result, the viewer can use their personal device as a secondary screen, for that iTV device, in order to capture the viewers' interactions.
 
-The architecture to accurately capture viewers interactions is shown in \figurename~\ref{fig_architecture}.
-This architecture consists of secondary screen devices (which might be implemented in personal devices such as smartphones and tablets) an iTV device, and the CP infrastructure (which includes an application hosting infrastructure).
+The architecture to accurately capture viewers interactions is shown in~\figurename~\ref{fig_architecture}.
+This architecture consists of secondary screen devices, which might be implemented in personal devices such as smartphones and tablets, an iTV device, and the CP infrastructure, which includes an application hosting infrastructure.
 The dashed green arrows represent the data flowing from the viewer to the CP infrastructure, while dotted blue arrows denote the flow from CP infrastructure to the viewers.
@@
 
-A presence service detects physically nearby secondary screen devices that have the services necessary for establishing a communication with iTV device.
+A presence service detects physically nearby secondary screen devices that have the services necessary to establish a communication with an iTV device.
 Modern commercial operating systems for personal devices enable service discovery by using peer-to-peer or WiFi communication.
 This feature is particularly appropriate for the TV environment, once the secondary screen device must be wirelessly linked up with the iTV device.
-The devices are synchronized in order to update contextual data about viewer profile and amount of people using a secondary device in the room.
-This contextual data can be used to learn how a viewer behave toward other viewers.
+The devices are synchronized in order to update the contextual data about the viewer profile and the amount of people using a secondary device in the room.
+This contextual data can be used to learn how a viewer behaves towards other viewers.
 For example, a viewer behaves in a particular way while watching TV alone and the same viewer can behave in a different way while watching in group.
 
-Viewers can use their secondary screen device to interact with TV by evaluating, commenting, sharing or even recommending TV content to other viewers.
-The data related to those actions are naturally provided by the viewers as part of the interaction when using interactive applications (developed by the CP) installed in their secondary screen devices.
+Viewers can use their secondary screen device to interact with TV by evaluating, commenting, sharing, or even recommending TV content to other viewers.
+The data related to those actions are naturally provided by the viewers, as part of the interaction, when using interactive applications (developed by the CP) installed in their secondary screen devices.
 The interactions and the contextual data are represented by the audience data, which is sent to CP infrastructure through Internet.
@@
 Thus, the personalized content can be delivered via Internet to the secondary screen device or even to the TV depending on the context.
+
+\blurb{Tens algum resumo dos resultados experimentais?}
 

### 2014-05-19 src/paper.tex
comments related to the conclusion

@@
 
-Since applications in the TV domain have the potential to reach out to millions of users at the same time, it is interesting that the CP infrastructure has the ability to deal with a large amount of incoming data~\cite{Lee2010}.
+Since applications in the TV domain have the potential to reach out millions of users at the same time, it is interesting that the CP infrastructure has the ability to deal with a large amount of incoming data~\cite{Lee2010}.
 The captured data relates to a single viewer in a given context.
-By using proper machine learning algorithms this data can provide valuable information about viewers.
+By using proper machine learning algorithms, this data can provide valuable information about viewers.
 Knowing this information can be very useful for CPs to improve their services aiming to increase the audience and enrich user experience.
@@
 These standards support the use of mobile devices as part of the system.
-So, since the mobile devices support the implementation of the DTV middleware, a cooperative exhibition of the DTV application can be managed with the iTV device, thus enabling the capture of viewers' interactions and contextual data.
+Since the mobile devices support the implementation of the DTV middleware, a cooperative exhibition of the DTV application can be managed with the iTV device, thus enabling the capture of viewers' interactions and contextual data.
+
+\blurb{Seria bom incluir algum resultado de experimento na conclusao}
 

### 2014-05-19 src/paper.tex
Removed some parts of the introduction

@@
 
-The interactions between viewers and TV can be classified as implicit and explicit.
+% The interactions between viewers and TV can be classified as implicit and explicit.
 Implicit interactions are part of the natural actions of watching TV (e.g., changing channels).
@@
 By using proper machine learning algorithms, CPs can infer viewers' tastes from audience data in order to deliver personalized content~\cite{Kim2012,Shin2009}.
-However, identifying viewers and capturing their individual interactions are still a challenging activity in the traditional TV environment, which consists of the TV and viewers interacting through a remote control (RC).
+However, identifying viewers and capturing their individual interactions are still a challenging activity in the traditional TV environment.%, which consists of the TV and viewers interacting through a remote control (RC).
 

### 2014-05-29 src/paper.tex
WIP: Rewriting the introduction

--- 2014-05-19 src/paper.tex
@@
 Thus, the interaction between viewers and TV became more elaborated than simply change channels and adjust sound volume.
+As a result, considering the interactions of millions of viewers, the captured data can provide valuable information for content providers (CPs)~\cite{Teixeira2010}.
 
-The interactions between viewers and TV can be classified as implicit and explicit.
-Implicit interactions are part of the natural actions of watching TV (e.g., changing channels).
-In explicit interactions, viewers contribute by actively evaluating, commenting, or sharing their opinion about a given content.
-Both kinds of interaction are important for applications that rely on audience generated data.
+The conventional remote control(RC) is typically used as a standard medium for viewers to interact with their TVs.
+However, the RC presents two notable problems: (1) to automatically identify the viewers who have possession of the RC; and (2) to capture viewers' interactions and the context in which they occur.
+Consequently, the whole group of viewers (e.g., a family) is handled as a single user % , rather than to the viewer in charge of the RC.
+As a result, the tastes that were inferred from the users in charge of the RC can be imposed on the tastes of the others.
 
-Considering the interactions of millions of viewers, the captured data can provide valuable information for content providers (CPs)~\cite{Teixeira2010}.
 For example, the CPs can obtain the user ratings given to TV programs, amount of viewers, and user profiles.
 By using proper machine learning algorithms, CPs can infer viewers' tastes from audience data in order to deliver personalized content~\cite{Kim2012,Shin2009}.
-However, identifying viewers and capturing their individual interactions are still a challenging activity in the traditional TV environment, which consists of the TV and viewers interacting through a remote control (RC).
+However, identifying viewers and capturing their individual interactions are still a challenging activity in the traditional TV environment.
 

### 2014-05-30 src/paper.tex
WIP: revising introduction

@@
 The conventional remote control(RC) is typically used as a standard medium for viewers to interact with their TVs.
-However, the RC presents two notable problems: (1) to automatically identify the viewers who have possession of the RC; and (2) to capture viewers' interactions and the context in which they occur.
-Consequently, the whole group of viewers (e.g., a family) is handled as a single user % , rather than to the viewer in charge of the RC.
+However, the interaction by using the RC presents two notable problems: (1) to automatically identify the viewers who have possession of it; and (2) to capture viewers' interactions and the context in which they occur.
+As a consequence, the entire group of viewers (e.g., a family) is treated as a single viewer.
 As a result, the tastes that were inferred from the users in charge of the RC can be imposed on the tastes of the others.

### 2014-06-02 src/paper.tex
WIP: Editing the introduction

@@
 It is also difficult to identify the context in which the interactions occur, e.g., to manage uniquely who is present in the environment and what is being watched on TV.
-In this paper, we present a novel architecture that facilitates viewers' individual interactions and contextual information in shared TV environments.
+In this paper, it is presented a novel architecture that facilitates viewers' individual interactions and contextual information in shared TV environments.
 One important application for this data consists of generating personalized content for single viewers and groups of viewers.
@@
 Thus, the interaction between viewers and TV became more elaborated than simply change channels and adjust sound volume.
-As a result, considering the interactions of millions of viewers, the captured data can provide valuable information for content providers (CPs)~\cite{Teixeira2010}.
+Considering the interactions of millions of viewers, the generated data can provide valuable information for content providers (CPs)~\cite{Teixeira2010}.
 
-The conventional remote control(RC) is typically used as a standard medium for viewers to interact with their TVs.
-However, the interaction by using the RC presents two notable problems: (1) to automatically identify the viewers who have possession of it; and (2) to capture viewers' interactions and the context in which they occur.
-As a consequence, the entire group of viewers (e.g., a family) is treated as a single viewer.
-As a result, the tastes that were inferred from the users in charge of the RC can be imposed on the tastes of the others.
+The conventional remote control (RC) is typically used as a standard medium for viewers to interact with their TVs.
+However, this medium presents two notable problems: (1) to automatically identify the viewers who have possession of the RC; and (2) to capture contextualized viewers' interactions.
+As a result, the entire group of viewers (e.g., a family) is treated as a single viewer and the tastes that were inferred from the users in charge of the RC can be imposed on the tastes of the others.
 
-For example, the CPs can obtain the user ratings given to TV programs, amount of viewers, and user profiles.
+Given the limitations of the conventional RC, a novel secondary screen architecture to identify viewers who are interacting with interactive TV (iTV) services is presented in this paper.
+Personal devices (e.g., smartphones and tablets) are used as secondary screens with the aim of capturing personal and contextualized interactions for each individual viewer in the iTV environment.
 By using proper machine learning algorithms, CPs can infer viewers' tastes from audience data in order to deliver personalized content~\cite{Kim2012,Shin2009}.
-However, identifying viewers and capturing their individual interactions are still a challenging activity in the traditional TV environment.
 
-The conventional RC presents two notable problems for capturing interactions from viewers: (1) to automatically identify the viewers that are in the TV environment in order to contextualize the interaction; and (2) to individually capture interactions from the viewer in charge of the RC.
-Thus, considering a scenario with the traditional TV environment and a conventional TV, all interactions are assigned to the whole group of viewers (e.g., a family), rather than to the viewer in charge of the RC.
-As a result, the tastes that were inferred from the users in charge of the RC can be imposed on the tastes of the others.
+% The approach described in this paper is based on the development of web services that provide means to access content from the CP and
+
+% Since, each viewer has its own personal device as secondary screen
+
+% A range of Applications and iTV services can be developed to obtain the user ratings given to TV programs, amount of viewers, and user profiles. .
+
+Data about viewers' interactions are captured by personal devices (e.g., smartphones and tablets) that are used to mediate the interaction between viewers and the iTV device.
+
+Personal devices (e.g., smartphones and tablets) are almost ubiquitous and present rich interactive capabilities.
 
 As a potential alternative to handle the limitations of the RC, the proliferation of personal mobile devices has changed the people behavior towards audiovisual content consumption in interactive TV environments.
-Nowadays, viewers rarely just sit passively and watch TV.
-In many situations, they use their personal devices as a secondary screen to perform activities related to TV watching, which end up enriching and immersing the user into the TV experience.
 
@@
 With respect to interaction capture, an approach described in~\cite{Teixeira2010} has shown an architecture to capture contextualized viewers interactions through a conventional RC.
-In this paper, we present a novel architecture that facilitates the identification of viewers, which is useful to contextualize the interactions.
-In addition, the identification of viewers facilitates the capture of their individual interactions.
 
@@
 
-\section{Architecture}
+\section{Secondary Screen Architecture and Implementation}
 
@@
 
+Nowadays, viewers rarely just sit passively and watch TV.
+In many situations, they use their personal devices as a secondary screen to perform activities related to TV watching, which end up enriching and immersing the user into the TV experience.
+
 \blurb{Seria bom incluir algum resultado de experimento na conclusao}

### 2014-06-02 src/paper.tex
WIP: Introduction

@@
 Given the limitations of the conventional RC, a novel secondary screen architecture to identify viewers who are interacting with interactive TV (iTV) services is presented in this paper.
-Personal devices (e.g., smartphones and tablets) are used as secondary screens with the aim of capturing personal and contextualized interactions for each individual viewer in the iTV environment.
-By using proper machine learning algorithms, CPs can infer viewers' tastes from audience data in order to deliver personalized content~\cite{Kim2012,Shin2009}.
+Personal devices (e.g., smartphones and tablets) are used as secondary screens with the aim of capturing personal and contextualized interactions from each individual viewer in iTV environments.
+Thus, iTV applications and services can be developed to obtain viewers' interaction data.
+Examples of these applications include rating TV programs, amount of viewers, and user profiles.
+
+By using proper machine learning algorithms, the captured data can reveal valuable information about viewers' tastes and behaviors~\cite{Kim2012,Shin2009}.
 
@@
 
-% A range of Applications and iTV services can be developed to obtain the user ratings given to TV programs, amount of viewers, and user profiles. .
-
 Data about viewers' interactions are captured by personal devices (e.g., smartphones and tablets) that are used to mediate the interaction between viewers and the iTV device.
@@
 Personal devices (e.g., smartphones and tablets) are almost ubiquitous and present rich interactive capabilities.
-
-As a potential alternative to handle the limitations of the RC, the proliferation of personal mobile devices has changed the people behavior towards audiovisual content consumption in interactive TV environments.
 

### 2014-06-02 src/paper.tex
WIP: some changes in text

@@
 Thus, the interaction between viewers and TV became more elaborated than simply change channels and adjust sound volume.
-Considering the interactions of millions of viewers, the generated data can provide valuable information for content providers (CPs)~\cite{Teixeira2010}.
+In this scenario, the data generated by the interactions of millions of viewers can provide valuable information for content providers (CPs) to improve their services and improve audience~\cite{Teixeira2010}.
 
-The conventional remote control (RC) is typically used as a standard medium for viewers to interact with their TVs.
+The conventional remote control (RC) is typically used as the standard medium for viewers to interact with their TVs.
 However, this medium presents two notable problems: (1) to automatically identify the viewers who have possession of the RC; and (2) to capture contextualized viewers' interactions.
@@
 Examples of these applications include rating TV programs, amount of viewers, and user profiles.
+
+% In order to support millions of viewers, this architecture was implemented under a cloud based platform
 

### 2014-06-03 src/paper.tex
Introduction finished

@@
 Thus, the interaction between viewers and TV became more elaborated than simply change channels and adjust sound volume.
-In this scenario, the data generated by the interactions of millions of viewers can provide valuable information for content providers (CPs) to improve their services and improve audience~\cite{Teixeira2010}.
+In this scenario, the data generated by the interactions of millions of viewers can provide valuable information for content providers (CPs) to improve their services~\cite{Teixeira2010}.
 
@@
 Given the limitations of the conventional RC, a novel secondary screen architecture to identify viewers who are interacting with interactive TV (iTV) services is presented in this paper.
-Personal devices (e.g., smartphones and tablets) are used as secondary screens with the aim of capturing personal and contextualized interactions from each individual viewer in iTV environments.
-Thus, iTV applications and services can be developed to obtain viewers' interaction data.
-Examples of these applications include rating TV programs, amount of viewers, and user profiles.
+Personal devices (e.g., smartphones and tablets) are used as secondary screens with the aim of capturing personal and contextualized interactions from each active viewer in iTV environments.
+By using proper machine learning algorithms, the personal data can reveal important information about viewers' tastes~\cite{Kim2012,Shin2009}.
 
-% In order to support millions of viewers, this architecture was implemented under a cloud based platform
+Since personal devices are almost ubiquitous, they stand out as a powerful mechanism to identify viewers by means of wireless communication technologies~\cite{Cabarcos2011}.
+In the architecture described in this paper, personal devices play an additional role as a medium of interaction since they are easy and intuitive to use in iTV environments, as pointed out in~\cite{Courtois2012}.
+An approach described in~\cite{Teixeira2010} has shown an architecture to capture viewers' interactions through a conventional RC.
+The architecture described in this paper goes further and captures individual and contextualized data.
+Since the TV environment is shared, the data about nearby viewers is captured to contextualize the interaction with the purpose of helping to infer the behaviors of one viewer toward others ones.
 
-By using proper machine learning algorithms, the captured data can reveal valuable information about viewers' tastes and behaviors~\cite{Kim2012,Shin2009}.
+% In order to support millions of viewers, this architecture was implemented under a cloud computing based platform.
 
-% The approach described in this paper is based on the development of web services that provide means to access content from the CP and
+% Data about viewers' interactions are captured by personal devices (e.g., smartphones and tablets) that are used to mediate the interaction between viewers and the iTV device.
 
-% Since, each viewer has its own personal device as secondary screen
-
-Data about viewers' interactions are captured by personal devices (e.g., smartphones and tablets) that are used to mediate the interaction between viewers and the iTV device.
-
-Personal devices (e.g., smartphones and tablets) are almost ubiquitous and present rich interactive capabilities.
-
-Secondary screen devices arise as a powerful mechanism to accurately identify viewers and  capture their interactions.
-In this research field, a typical approach is to use personal devices (e.g., smartphones and tablets) as secondary screens~\cite{Courtois2012}, since they are almost ubiquitous and present good graphical capabilities.
-The connectivity capabilities of smartphones were used in~\cite{Cabarcos2011}, where the Bluetooth technology was used to detect nearby users.
-With respect to interaction capture, an approach described in~\cite{Teixeira2010} has shown an architecture to capture contextualized viewers interactions through a conventional RC.
-
-\blurb{Achei que o inicio da introducao esta com muito bla bla bla.
+% \blurb{Achei que o inicio da introducao esta com muito bla bla bla.
 Tente ir direto ao ponto, dando uma visao geral da sua abordagem, as vantagens e desvantages com as outras abordagens existentes na literatura e as principais contribuicoes do seu trabalho.
@@
 
-\section{Secondary Screen Architecture and Implementation}
-
-\blurb{Tente colocar um nome mais sugestivo para esta secao, em vez ``Architecture'' somente.}
+\section{Secondary Screen Architecture}
 
@@
 
-\blurb{Tens algum resumo dos resultados experimentais?}
+% \blurb{Tens algum resumo dos resultados experimentais?}
 
@@
 
-\blurb{Seria bom incluir algum resultado de experimento na conclusao}
+% \blurb{Seria bom incluir algum resultado de experimento na conclusao}
 

### 2014-06-04 src/paper.tex
Review finished

@@
 This architecture consists of secondary screen devices, which might be implemented in personal devices such as smartphones and tablets, an iTV device, and the CP infrastructure, which includes an application hosting infrastructure.
+Since applications in the TV domain have the potential to reach out millions of users at the same time, it is interesting that the CP infrastructure has the ability to deal with a large amount of incoming data~\cite{Lee2010}.
 The dashed green arrows represent the data flowing from the viewer to the CP infrastructure, while dotted blue arrows denote the flow from CP infrastructure to the viewers.
@@
 
+The Data Writer components consists basically of web services that provide interoperability among secondary screen devices, iTV devices and the CP infrastructure.
+The main advantage of this approach is the independence of operating system and programming languages used in the personal devices.
+
 Once the audience data is stored, the CP can use this data to derive valuable information about users' tastes.
@@
 
-Since applications in the TV domain have the potential to reach out millions of users at the same time, it is interesting that the CP infrastructure has the ability to deal with a large amount of incoming data~\cite{Lee2010}.
 The captured data relates to a single viewer in a given context.
@@
 
+An implementation of the architecture described in this paper was made using a commercial cloud computing environment, which plays the role of the CP infrastructure.
+A web application that consists of a content rating system was developed for both storing and retrieving viewers' tastes toward audio visual content provided by the CPs.
+Some RESTful web services were developed to store and retrieve viewers' interactions.
+The secondary screen applications were developed in Java and the viewer's interaction data were saved by means of HTTP requests.
+
 The architecture described in this paper is also technically feasible to implement in Digital TV (DTV) systems and can be particularly interesting for CPs in DTV standards like the Brazilian (ISDB-TB) and the Japanese(ISDB-T).
@@
 Since the mobile devices support the implementation of the DTV middleware, a cooperative exhibition of the DTV application can be managed with the iTV device, thus enabling the capture of viewers' interactions and contextual data.
-
-Nowadays, viewers rarely just sit passively and watch TV.
-In many situations, they use their personal devices as a secondary screen to perform activities related to TV watching, which end up enriching and immersing the user into the TV experience.
-
-% \blurb{Seria bom incluir algum resultado de experimento na conclusao}
 

### 2014-06-04 src/paper.tex
Abstract changed

@@
 
-\title{A secondary screen architecture to accurately capture viewers' interactions in an iTV environment}
+\title{A Secondary Screen Architecture to Accurately Capture Viewers' Interactions in an iTV Environment}
 
@@
 \begin{abstract}
-TV watching is frequently seen as a social activity.
-People gather around the television for entertainment or information.
-Advances in TV technology have enabled the viewer to actively interact with the TV through interactive applications instead of just passively watching TV.
-Since the TV set is shared by many people and the interaction occurs by using a single remote control, it is difficult to accurately capture the individual needs from each viewer.
-It is also difficult to identify the context in which the interactions occur, e.g., to manage uniquely who is present in the environment and what is being watched on TV.
-In this paper, it is presented a novel architecture that facilitates viewers' individual interactions and contextual information in shared TV environments.
+Advances in TV technology have enabled viewers to actively interact with the TV through interactive applications instead of just passively watching TV.
+Typically, a conventional remote control is shared by many viewers and it is difficult to accurately capture the individual interaction data from each viewer.
+It is also difficult to identify the context in which the interactions occur, e.g., to manage precisely who is present in the environment and what is being watched on TV.
+In this paper, it is presented a novel architecture that facilitates the capture of viewers' individual interactions and contextual data in shared TV environments.
+This architecture uses personal devices as secondary screen devices with the aim of identifying viewers and capturing their interactions with interactive TV devices.
 One important application for this data consists of generating personalized content for single viewers and groups of viewers.

### 2014-06-04 src/paper.tex
WIP: Changes in introduction and architecture

@@
 This architecture uses personal devices as secondary screen devices with the aim of identifying viewers and capturing their interactions with interactive TV devices.
-One important application for this data consists of generating personalized content for single viewers and groups of viewers.
+A work in progress implementation of this architecture is briefly reported as an audio-visual content rating system for interactive TV.
 \end{abstract}
@@
 
-TV watching is essentially a social activity, where family, friends, or people with a common interest share the same space and TV set for entertainment or information.
+TV watching is essentially a social activity, where groups of people with a common interest share the same space and TV set for entertainment or information.
 After many advances in technologies, for interactive and digital TV, it is possible to develop high-level applications to enrich the watching experience.
@@
 
-The conventional remote control (RC) is typically used as the standard medium for viewers to interact with their TVs.
-However, this medium presents two notable problems: (1) to automatically identify the viewers who have possession of the RC; and (2) to capture contextualized viewers' interactions.
+The conventional remote control (RC) is typically seen as the standard medium of interaction among the viewers and their TVs.
+However, this medium presents two notable problems: (1) to precisely and automatically identify the viewers who have possession of the RC; and (2) to capture contextualized viewers' interactions.
 As a result, the entire group of viewers (e.g., a family) is treated as a single viewer and the tastes that were inferred from the users in charge of the RC can be imposed on the tastes of the others.
@@
 Personal devices (e.g., smartphones and tablets) are used as secondary screens with the aim of capturing personal and contextualized interactions from each active viewer in iTV environments.
-By using proper machine learning algorithms, the personal data can reveal important information about viewers' tastes~\cite{Kim2012,Shin2009}.
+By using proper machine learning algorithms, the personal and contextualized data can reveal important information about viewers' tastes~\cite{Kim2012,Shin2009}.
 
@@
 
-% In order to support millions of viewers, this architecture was implemented under a cloud computing based platform.
-
-% Data about viewers' interactions are captured by personal devices (e.g., smartphones and tablets) that are used to mediate the interaction between viewers and the iTV device.
-
-% \blurb{Achei que o inicio da introducao esta com muito bla bla bla.
-Tente ir direto ao ponto, dando uma visao geral da sua abordagem, as vantagens e desvantages com as outras abordagens existentes na literatura e as principais contribuicoes do seu trabalho.
-Por que voce acredita que os revisores devem aceitar o seu trabalho?
-Qual eh a inovacao?
-Quais sao as contribuicoes?
-Quais sao as diferencas para os trabalhos que ja existem.
-Diminua o bla bla bla do inicio e tente responder estas perguntas na introducao do seu trabalho.}
-
 \section{Secondary Screen Architecture}
@@
 
-The architecture to accurately capture viewers interactions is shown in~\figurename~\ref{fig_architecture}.
-This architecture consists of secondary screen devices, which might be implemented in personal devices such as smartphones and tablets, an iTV device, and the CP infrastructure, which includes an application hosting infrastructure.
-Since applications in the TV domain have the potential to reach out millions of users at the same time, it is interesting that the CP infrastructure has the ability to deal with a large amount of incoming data~\cite{Lee2010}.
-The dashed green arrows represent the data flowing from the viewer to the CP infrastructure, while dotted blue arrows denote the flow from CP infrastructure to the viewers.
-The continuous black arrows represent the data flow inside the CP infrastructure.
+The architecture to precisely capture viewers interactions and the context is shown in~\figurename~\ref{fig_architecture}.
+This architecture consists of tree subsystems that interact with each other: (1) secondary screen devices, which might be implemented in personal devices; an (2) iTV device; and (3) the CP infrastructure, which includes an application hosting infrastructure.
+The dashed arrows represent the data flowing from the viewer to the CP infrastructure, while dotted arrows denote the flow from CP infrastructure to the viewers.
+The continuous arrows represent the data flow inside the CP infrastructure.
+Since applications in the TV domain have the potential to reach out millions of users at the same time, it is interesting to consider the use of a cloud based platform in the CP infrastructure~\cite{Lee2010}.
 

### 2014-06-05 src/paper.tex

@@
 Modern commercial operating systems for personal devices enable service discovery by using peer-to-peer or WiFi communication.
-This feature is particularly appropriate for the TV environment, once the secondary screen device must be wirelessly linked up with the iTV device.
+This feature is particularly appropriated for the TV environment, once the secondary screen device must be wirelessly linked up with the iTV device.
 The devices are synchronized in order to update the contextual data about the viewer profile and the amount of people using a secondary device in the room.
 This contextual data can be used to learn how a viewer behaves towards other viewers.
-For example, a viewer behaves in a particular way while watching TV alone and the same viewer can behave in a different way while watching in group.
+For example, a viewer behaves in a particular way while watching TV alone and the same viewer behaves in a different way while watching in group.
 
@@
 
-The Data Writer components consists basically of web services that provide interoperability among secondary screen devices, iTV devices and the CP infrastructure.
+The Data Writer service in the application hosting infrastructure consists basically of web services that provide interoperability among secondary screen devices, iTV devices and the CP infrastructure.
 The main advantage of this approach is the independence of operating system and programming languages used in the personal devices.
@@
 This information can be used to provide personalized content to viewers based on the current context.
-Thus, the personalized content can be delivered via Internet to the secondary screen device or even to the TV depending on the context.
-
-% \blurb{Tens algum resumo dos resultados experimentais?}
+Thus, the personalized content can be delivered via Internet to the secondary screen device or even to the TV, depending on the context.
 
@@
 The use of personal devices as secondary screens arises as a new trend in TV environments.
-This is paper presents an architecture to accurately capture audience data by using secondary screen devices and iTV devices.
-The implementation of this architecture is a work in progress, and take advantage of the ubiquity and interactive capabilities of secondary screen devices to identify viewers and provide feedback for CPs through viewers' interactions and contextual information.
-
+This paper presents an architecture to accurately capture audience data by using secondary screen devices and iTV devices.
+This architecture take advantage of the ubiquity and interactive capabilities of secondary screen devices to identify viewers and provide feedback for CPs through viewers' interactions and contextual information.
 The captured data relates to a single viewer in a given context.
 By using proper machine learning algorithms, this data can provide valuable information about viewers.
-Knowing this information can be very useful for CPs to improve their services aiming to increase the audience and enrich user experience.
+Knowing this information can be very useful for CPs to improve their services aiming to increase the audience and to enrich user experience.
 As an example, CPs can deliver personalized content for a single person as well as for groups of people by using content recommendation algorithms.
 
-An implementation of the architecture described in this paper was made using a commercial cloud computing environment, which plays the role of the CP infrastructure.
-A web application that consists of a content rating system was developed for both storing and retrieving viewers' tastes toward audio visual content provided by the CPs.
-Some RESTful web services were developed to store and retrieve viewers' interactions.
-The secondary screen applications were developed in Java and the viewer's interaction data were saved by means of HTTP requests.
+An implementation of the architecture that is described in this paper is under development in a commercial cloud computing platform, which plays the role of the application hosting infrastructure.
+A web application that consists of a content rating system was developed to store  viewers' evaluations toward audio visual content, and a secondary screen client is under development.
+Some RESTful web services were developed to provide interoperability among the secondary screen devices and the web application.
+Thus, the interactions that are captured during the use of the secondary screen application are sent to the CP by means of HTTP requests to RESTful web services.
+Although the secondary screen application is not fully finished yet, some specific features have been tested to capture viewers' interactions as they occur.
 
-The architecture described in this paper is also technically feasible to implement in Digital TV (DTV) systems and can be particularly interesting for CPs in DTV standards like the Brazilian (ISDB-TB) and the Japanese(ISDB-T).
+The architecture described in this paper is also technically feasible to implement in Digital TV (DTV) systems and can be particularly interesting for CPs in DTV standards, e.g., the Brazilian (ISDB-TB) and the Japanese(ISDB-T).
 These standards support the use of mobile devices as part of the system.

### 2014-06-06 src/paper.tex
Email and README.md added to the project

@@
 Manaus, AM, Brazil\\
-Email: vicente@ufam.edu.br}
+Email: \{lucascordeiro,vicente\}@ufam.edu.br}
 }

### 2014-06-16 src/paper.tex
WIP: new paper to TCE

--- 2014-06-05 src/paper.tex
@@
-\documentclass[conference,a4paper]{IEEEtran}
+\documentclass[journal]{IEEEtran}
 
@@
 
-\title{A Secondary Screen Architecture to Accurately Capture Viewers' Interactions in an iTV Environment}
+\title{Capturing Explicit Viewers' Interactions Using Second Screen Devices in iTV Environments}
+\author{Ricardo~Erikson~V.~de~S.~Rosa, 
+	Vicente~Ferreira~de~Lucena~Jr.,\IEEEmembership{Member,IEEE,}
+	and~Lucas~Carvalho~Cordeiro
+	%
+\thanks{Ricardo Erikson V.
+de S.
+Rosa is with the Graduate Program in Electrical Engineering, Federal University of Minas Gerais, Av.
+Antônio Carlos 6627, 31270-901, Belo Horizonte, MG, Brazil(e-mail: ricardoerikson@ufmg.br)}%
+\thanks{Vicente Ferreira de Lucena Jr. is with the PPGEE, PPGI, and CETELI - Electronics and Information Technology R\&D Center at UFAM, Manaus, Amazonas, Brazil (e-mail: vicente@ufam.edu.br)}%
+\thanks{Lucas Carvalho Cordeiro is with the PPGEE-UFAM, and CETELI - Electronics and Information Technology R\&D Center, Manaus, Amazonas, Brazil (e-mail:lucascordeiro@ufam.edu.br)}}
 
-\author{\IEEEauthorblockN{Ricardo E. V. de S. Rosa}
-\IEEEauthorblockA{Graduate Program in Electrical Engineering \\ Federal University of Minas Gerais \\ Av. Antônio Carlos 6627, 31270-901, \\Belo Horizonte, MG, Brazil\\
-Email: ricardoerikson@ufmg.br}
-\and
-\IEEEauthorblockN{Lucas C. Cordeiro and Vicente F. de Lucena Junior}
-\IEEEauthorblockA{Graduate Program in Electrical Engineering \\
-Federal University of Amazonas \\
-Manaus, AM, Brazil\\
-Email: vicente@ufam.edu.br}
-}
+% \author{\IEEEauthorblockN{Ricardo E. V. de S. Rosa}
+% \IEEEauthorblockA{\\
+% Email: ricardoerikson@ufmg.br}
+% \and
+% \IEEEauthorblockN{Lucas C. Cordeiro and Vicente F. de Lucena Junior}
+% \IEEEauthorblockA{Graduate Program in Electrical Engineering \\
+% Federal University of Amazonas \\
+% Manaus, AM, Brazil\\
+% Email: \{lucascordeiro,vicente\}@ufam.edu.br}
+% }
 \maketitle
@@
 \end{abstract}
-
+\begin{IEEEkeywords}
+Second screen devices, Interactive Television, Viewers' Interactions
+\end{IEEEkeywords}
 \IEEEpeerreviewmaketitle
@@
 As a result, the entire group of viewers (e.g., a family) is treated as a single viewer and the tastes that were inferred from the users in charge of the RC can be imposed on the tastes of the others.
+
+The interactions between viewers and TV can be classified as implicit and explicit.
+Implicit interactions are part of the natural actions of watching TV e.g., changing channels.
+In explicit interactions, viewers contribute by actively evaluating, commenting or sharing their opinion about a given content.
+Both kinds of interaction are important for applications that rely on audience generated data.
 

### 2014-06-17 src/paper.tex
Changes in abstract

@@
 Advances in TV technology have enabled viewers to actively interact with the TV through interactive applications instead of just passively watching TV.
-Typically, a conventional remote control is shared by many viewers and it is difficult to accurately capture the individual interaction data from each viewer.
+Typically, the interactions occur by using a conventional remote control, which is shared by many viewers and hinders the capture of viewers' individual interactions.
 It is also difficult to identify the context in which the interactions occur, e.g., to manage precisely who is present in the environment and what is being watched on TV.
 In this paper, it is presented a novel architecture that facilitates the capture of viewers' individual interactions and contextual data in shared TV environments.
-This architecture uses personal devices as secondary screen devices with the aim of identifying viewers and capturing their interactions with interactive TV devices.
-A work in progress implementation of this architecture is briefly reported as an audio-visual content rating system for interactive TV.
+This architecture uses personal devices as second screen devices with the aim of identifying viewers and capturing their interactions with interactive TV devices.
+An implementation of this architecture is reported as an audio-visual content rating system for interactive TV.
 \end{abstract}

### 2014-06-17 src/paper.tex
Some changes in introduction

@@
 
-% \author{\IEEEauthorblockN{Ricardo E. V. de S. Rosa}
-% \IEEEauthorblockA{\\
-% Email: ricardoerikson@ufmg.br}
-% \and
-% \IEEEauthorblockN{Lucas C. Cordeiro and Vicente F. de Lucena Junior}
-% \IEEEauthorblockA{Graduate Program in Electrical Engineering \\
-% Federal University of Amazonas \\
-% Manaus, AM, Brazil\\
-% Email: \{lucascordeiro,vicente\}@ufam.edu.br}
-% }
 \maketitle
@@
 
-TV watching is essentially a social activity, where groups of people with a common interest share the same space and TV set for entertainment or information.
-After many advances in technologies, for interactive and digital TV, it is possible to develop high-level applications to enrich the watching experience.
+TV watching is essentially a social activity, where groups of people with a common interest share the same space and TV set for entertainment or information~\cite{Masthoff2004}.
+After many advances in technologies for interactive and digital TV, it is possible to develop high-level applications to enrich the watching experience.
 Thus, the interaction between viewers and TV became more elaborated than simply change channels and adjust sound volume.
-In this scenario, the data generated by the interactions of millions of viewers can provide valuable information for content providers (CPs) to improve their services~\cite{Teixeira2010}.
+In this scenario, the data generated by viewers' interactions can provide valuable information to improve TV services, especially when data comes from millions of viewers~\cite{Teixeira2010}.
 
@@
 Both kinds of interaction are important for applications that rely on audience generated data.
+
+As a potential alternative to handle the limitations of the RC for capturing interactions from viewers, the proliferation of personal mobile devices has changed the people behavior towards audiovisual content consumption in interactive TV environments.
+Nowadays, viewers rarely just sit passively and watch TV.
+In many situations, they use their personal devices as a secondary screen to perform activities related to TV watching, which end up enriching and immersing the user into the TV experience.
 

### 2014-06-20 src/paper.tex
Paper sections structure
WIP: Second screen usage in itv

@@
 
-\title{Capturing Explicit Viewers' Interactions Using Second Screen Devices in iTV Environments}
+\title{A Second Screen Architecture for Capturing Explicit Viewers' Interactions in TV Environments}
 \author{Ricardo~Erikson~V.~de~S.~Rosa, 
@@
 \end{abstract}
+
 \begin{IEEEkeywords}
@@
 \end{IEEEkeywords}
+
 \IEEEpeerreviewmaketitle
@@
 
-As a potential alternative to handle the limitations of the RC for capturing interactions from viewers, the proliferation of personal mobile devices has changed the people behavior towards audiovisual content consumption in interactive TV environments.
-Nowadays, viewers rarely just sit passively and watch TV.
-In many situations, they use their personal devices as a secondary screen to perform activities related to TV watching, which end up enriching and immersing the user into the TV experience.
-
 Given the limitations of the conventional RC, a novel secondary screen architecture to identify viewers who are interacting with interactive TV (iTV) services is presented in this paper.
@@
 Since the TV environment is shared, the data about nearby viewers is captured to contextualize the interaction with the purpose of helping to infer the behaviors of one viewer toward others ones.
+
+\section{Background}
+
+The key concepts and some relevant related works that underpin the proposed architecture are described in this section.
+Section~\ref{ssec_ss_usage} presents the usage of second screen devices in iTV environments.
+Section~\ref{ssec_capture_viewers_int_data} describes the current status of research on the capture of viewers' interactions.
+The identification of viewers in iTV environments is described in Section~\ref{ssec_ident_viewers_itv}.
+
+\subsection{Second Screen Usage in iTV}
+\label{ssec_ss_usage}
+
+Nowadays, TV viewers rarely just sit passively and watch TV.
+In many situations, they use second screen devices to perform activities related to TV watching.
+In the context of TV, second screen refers to the use of a computing device, which is usually a smartphone or a tablet PC, to enhance the viewing experience.
+
+Research studies on second screen have been conducted in TV environments since late 1990s.
+In 1996, Robertson~\emph{et~al}.~\cite{Robertson1996} developed a prototype where handheld devices were used to control TV and to provide additional content (e.g., textual information) related to the audio-visual content presented in the TV screen.
+More recent studies show that viewers evaluated positively the use of second screen devices to control the TV and to enrich the TV content~\cite{Cesar2008,Cesar2011,Tsekleves2011}.
+Cesar~\emph{et~al.}~\cite{Cesar2009} contributed identifying and modeling the requirements to implement an interactive TV architecture based on second screen devices.
+
+A review of the related works shown that the usage of personal devices as second screen devices has some advantages. Firstly, the visual elements
+
+\subsection{Capture of Viewers' Interaction Data}
+\label{ssec_capture_viewers_int_data}
+
+With respect to interaction capture, an approach described in~\cite{Teixeira2010} shown an architecture to capture contextualized viewers interactions through a conventional RC.
+In this paper, we present an architecture that facilitates the identification of viewers, which is useful to contextualize the interactions.
+In addition, the identification of viewers facilitates the capture of their individual interactions.
+
+\subsection{Identification of Viewers in iTV}
+\label{ssec_ident_viewers_itv}
+
+The connectivity capabilities of smartphones were used in~\cite{Cabarcos2011}, where the Bluetooth technology was used to detect nearby users.
+
+\section{Overview of the Secondary Screen Architecture}
 
@@
 
+\subsection{Web Services Design}
+
+- RESTful API
+
+- Resources
+
+\subsection{iTV Environment}
+
+- set-top box
+
+- Service discovery
+
+- context synchronization
+
+\subsection{Second Screen Devices}
+
+\section{Secondary Screen Implementation}
+
+\subsection{RESTful API}
+
+\subsection{Web Service Consumption}
+
+\subsection{iTV and Secondary Screen Devices Communication}
+
+\section{Experiments}
+
 \section{Conclusions}

### 2014-06-22 src/paper.tex
Section Second Screen usage in iTV finished

@@
 
-\title{A Second Screen Architecture for Capturing Explicit Viewers' Interactions in TV Environments}
+\title{A Second Screen Architecture for Capturing Viewers' Interaction Data in TV Environments}
 \author{Ricardo~Erikson~V.~de~S.~Rosa, 
@@
 
-The interactions between viewers and TV can be classified as implicit and explicit.
-Implicit interactions are part of the natural actions of watching TV e.g., changing channels.
-In explicit interactions, viewers contribute by actively evaluating, commenting or sharing their opinion about a given content.
-Both kinds of interaction are important for applications that rely on audience generated data.
-
 Given the limitations of the conventional RC, a novel secondary screen architecture to identify viewers who are interacting with interactive TV (iTV) services is presented in this paper.
@@
 
-A review of the related works shown that the usage of personal devices as second screen devices has some advantages. Firstly, the visual elements
+A review of the related works shown that the usage of personal devices as second screen devices has two main advantages.
+Firstly, the visual elements (e.g., an EPG guide) that are commonly shown on the TV screen can be presented on the second screen device in such a way that the viewing experience of other viewers is not affected.
+Secondly, the second screen device can enhance the capabilities of a conventional remote control by providing more sophisticated interactions, e.g., gesture and voice commands.
 
@@
 \label{ssec_capture_viewers_int_data}
+
+The interactions between viewers and TV can be classified as implicit and explicit.
+Implicit interactions are part of the natural actions of watching TV e.g., changing channels and adjust sound volume.
+In explicit interactions, viewers contribute by actively performing actions that are not usual in the their experience from conventional TV.
+Such actions include among other things, evaluating audio-visual content, commenting and sharing opinion about a given content through iTV applications.
+Both kinds of interaction are important for services that rely on audience generated data, e.g., content personalization and targeted advertising.
 

### 2014-06-23 src/paper.tex
'Capture of Viewer's Interaction Data' section finished

@@
 Thus, the interaction between viewers and TV became more elaborated than simply change channels and adjust sound volume.
-In this scenario, the data generated by viewers' interactions can provide valuable information to improve TV services, especially when data comes from millions of viewers~\cite{Teixeira2010}.
+In this scenario, .
 
@@
 Such actions include among other things, evaluating audio-visual content, commenting and sharing opinion about a given content through iTV applications.
-Both kinds of interaction are important for services that rely on audience generated data, e.g., content personalization and targeted advertising.
+Both kinds of interaction are important for services that rely on audience generated data, e.g., content personalization, targeted advertising or content improvement.
 
-With respect to interaction capture, an approach described in~\cite{Teixeira2010} shown an architecture to capture contextualized viewers interactions through a conventional RC.
-In this paper, we present an architecture that facilitates the identification of viewers, which is useful to contextualize the interactions.
-In addition, the identification of viewers facilitates the capture of their individual interactions.
+ Some researchers proposed mechanisms to capture viewers' interactions in TV environments~\cite{Teixeira2010,Basilio2013,DosSantos2013}.
+Teixeira~\emph{et~al.}~\cite{Teixeira2010} described an infrastructure to capture both implicit and explicit interactions that are performed by means of a conventional remote control.
+The remote control events are recorded in XML format using elements to describe contextual information, implicit and explicit interactions.
+Basílio~\emph{et~al.}~\cite{Basilio2013} described two methods for capturing viewers interactions: (1) one using a middleware extension, which adds new functionalities to an existing standard; and (2) another one using applications that are added to the broadcaster data stream, which requires no changes in the middleware.
+
+ Another research study conduced by Santos and Bianchini~\cite{DosSantos2013}, described a advertising system for interactive TV, which uses viewers interactions in order to generate targeted advertising.
+Basically, the described system implicitly captures metadata (e.g., the category of a TV program) information when interactions take place.
+
+ The data generated by viewers' interactions can provide valuable information to improve TV personalization, especially when data comes from millions of viewers.
+For instance, viewers can change quickly the channel when a TV program loses their interest.
+Knowing when the channel change occurs can help producers to improve the content in order to increase viewers interest.
 
@@
 The connectivity capabilities of smartphones were used in~\cite{Cabarcos2011}, where the Bluetooth technology was used to detect nearby users.
+
+\cite{Hwang2007}
+
+With respect to interaction capture, an approach described in~\cite{Teixeira2010} shown an architecture to capture contextualized viewers interactions through a conventional RC.
+In this paper, we present an architecture that facilitates the identification of viewers, which is useful to contextualize the interactions.
+In addition, the identification of viewers facilitates the capture of their individual interactions.
 

### 2014-06-24 src/paper.tex
'Viewers' Identification in iTV' section finished

@@
 
-\subsection{Identification of Viewers in iTV}
+\subsection{Viewers' Identification in iTV}
 \label{ssec_ident_viewers_itv}
 
-The connectivity capabilities of smartphones were used in~\cite{Cabarcos2011}, where the Bluetooth technology was used to detect nearby users.
+Since the TV environment can be shared by many viewers (e.g., a family or group of friends), audience identification becomes an important feature for interactive TV systems.
+Many modern computational systems consist of identifying viewers in order to provide personalized services.
+Unfortunately, the traditional model of interaction, which uses a conventional remote control, hinders the identification of viewers.
+The only existing remote control is shared by all viewers and it is difficult to manage who is in possession of the remote control at a given moment.
 
-\cite{Hwang2007}
+Some research studies have addressed automated technologies for viewers' identification in interactive TV environments.
+Cabarcos~\emph{et~al.}~\cite{Cabarcos2011} described an architecture that uses the connectivity capabilities of smartphones to detect viewers nearby the TV terminal at every moment.
+Each smartphone has a hardware identifier(e.g., the Bluetooth MAC address), which is associated to a viewer identifier according to an association table.
+Hwang~\emph{et~al.}~\cite{Hwang2007} presented a method based on face recognition to identify viewers in interactive TV environments.
+The method uses sensors to capture images of the audience and then employs images processing techniques and machine learning algorithm for both detecting and recognizing the faces of the viewers.
 

### 2014-06-24 src/paper.tex
WIP: architecture overview

@@
 
-Since the TV environment can be shared by many viewers (e.g., a family or group of friends), audience identification becomes an important feature for interactive TV systems.
+Since the TV environment can be shared by many viewers (e.g., a family or a group of friends), audience identification becomes an important feature for interactive TV systems.
 Many modern computational systems consist of identifying viewers in order to provide personalized services.
 Unfortunately, the traditional model of interaction, which uses a conventional remote control, hinders the identification of viewers.
-The only existing remote control is shared by all viewers and it is difficult to manage who is in possession of the remote control at a given moment.
+The only existing remote control is shared by many viewers and it is difficult to manage who is in possession of the remote control at a given moment.
 
@@
 
-With respect to interaction capture, an approach described in~\cite{Teixeira2010} shown an architecture to capture contextualized viewers interactions through a conventional RC.
-In this paper, we present an architecture that facilitates the identification of viewers, which is useful to contextualize the interactions.
-In addition, the identification of viewers facilitates the capture of their individual interactions.
+\section{Overview of the Second Screen Architecture}
 
-\section{Overview of the Secondary Screen Architecture}
+The architecture presented in this paper applies to the capture of audience interactions by means of the use of second screen devices in interactive TV environments.
+Personal devices are used as second screen devices to enrich the interactivity between viewers and TV at the same time that they allow the capture of viewers' interactions.
 
-\section{Secondary Screen Architecture}
+\begin{figure}[!t]
+	\centering
+	\includegraphics[width=3.5in]{img/architecture.pdf}
+	\caption{Architecture to capture audience interactions. The viewers' interactions and the contextual information are captured and stored in the CP infrastructure.}
+	\label{fig_architecture}
+\end{figure}
 
-Initially, the CP infrastructure must provide an authentication mechanism so that viewers and their interactive TV (iTV) devices (e.g., set-top boxes or smart TVs) can be authenticated by the CP in order to use the secondary screen services.
-Once one iTV device is authenticated, the viewers in the TV environment must link their user accounts up with the iTV device.
-As a result, the viewer can use their personal device as a secondary screen, for that iTV device, in order to capture the viewers' interactions.
-
-The architecture to precisely capture viewers interactions and the context is shown in~\figurename~\ref{fig_architecture}.
+An overview of the architecture described in this paper is presented in \figurename~\ref{fig_architecture}.
 This architecture consists of tree subsystems that interact with each other: (1) secondary screen devices, which might be implemented in personal devices; an (2) iTV device; and (3) the CP infrastructure, which includes an application hosting infrastructure.
@@
 
-\begin{figure}[!t]
-	\centering
-	\includegraphics[width=3.5in]{img/architecture.pdf}
-	\caption{Architecture to capture audience data. The viewers' interactions and the contextual information are captured and stored in the CP infrastructure.}
-	\label{fig_architecture}
-\end{figure}
+\section{Second Screen Architecture}
+
+Initially, the CP infrastructure must provide an authentication mechanism so that viewers and their interactive TV (iTV) devices (e.g., set-top boxes or smart TVs) can be authenticated by the CP in order to use the secondary screen services.
+Once one iTV device is authenticated, the viewers in the TV environment must link their user accounts up with the iTV device.
+As a result, the viewer can use their personal device as a secondary screen, for that iTV device, in order to capture the viewers' interactions.
 

### 2014-06-25 src/paper.tex
Overview figure changed

@@
 The architecture presented in this paper applies to the capture of audience interactions by means of the use of second screen devices in interactive TV environments.
-Personal devices are used as second screen devices to enrich the interactivity between viewers and TV at the same time that they allow the capture of viewers' interactions.
+Personal devices are used as second screen devices to enrich the interactivity between viewers and TV at the same time that they facilitate the capture of viewers' interactions.
 
@@
 	\centering
-	\includegraphics[width=3.5in]{img/architecture.pdf}
+	\includegraphics[width=3.5in]{img/architecture-overview.pdf}
 	\caption{Architecture to capture audience interactions. The viewers' interactions and the contextual information are captured and stored in the CP infrastructure.}

### 2014-06-25 src/paper.tex
Overview finished

@@
 An overview of the architecture described in this paper is presented in \figurename~\ref{fig_architecture}.
-This architecture consists of tree subsystems that interact with each other: (1) secondary screen devices, which might be implemented in personal devices; an (2) iTV device; and (3) the CP infrastructure, which includes an application hosting infrastructure.
+This architecture consists of tree subsystems that interact with each other: (1) second screen devices, which are personal devices; an (2) iTV terminal; and (3) the CP infrastructure.
+The CP infrastructure includes an application hosting infrastructure, which comprises the computing resources (e.g., application server, database services and storage services) that are necessary to host web applications.
 The dashed arrows represent the data flowing from the viewer to the CP infrastructure, while dotted arrows denote the flow from CP infrastructure to the viewers.
 The continuous arrows represent the data flow inside the CP infrastructure.
+
+A web application is hosted in the application hosting infrastructure.
+This application contains web services that are designed for interoperability between the second screen devices, the iTV terminal and the CP infrastructure.
+A authentication mechanism must be provided so that the viewers can connect their second screen devices and interactive TV terminals (e.g., set-top boxes or smart TVs) with an user account.
+Once the authentication is completed, the second screen device can be used to interact with the iTV terminal.
+
+A presence service detects physically nearby second screen devices that have the services necessary to establish a communication with an iTV terminal.
+This feature is particularly appropriated for the TV environment, once second screen devices are usually connected with an iTV terminal via WiFi.
+The devices are synchronized in order to update the contextual data about the viewer profile, amount of people using a second screen devices in the room and content being watched.
+This contextual data can be used to learn how a viewer behaves towards other viewers.
+For example, a viewer behaves in a particular way while watching TV alone and the same viewer behaves in a different way while watching in group.
+
+The second screen devices and iTV terminals have client applications that consume the web services in order to present the content to the viewers.
+These client applications also use the web services as an interface to send interaction data to the CP.
 Since applications in the TV domain have the potential to reach out millions of users at the same time, it is interesting to consider the use of a cloud based platform in the CP infrastructure~\cite{Lee2010}.
 
+% descrever a possibilidade de autenticação em outra arte do texto: A possible way to do so is to generate a registration code for each equipment
+
+% commentar sobre a sincronização de contexto entre iTV terminals e second screen devices.
+
 \section{Second Screen Architecture}
-
-Initially, the CP infrastructure must provide an authentication mechanism so that viewers and their interactive TV (iTV) devices (e.g., set-top boxes or smart TVs) can be authenticated by the CP in order to use the secondary screen services.
-Once one iTV device is authenticated, the viewers in the TV environment must link their user accounts up with the iTV device.
-As a result, the viewer can use their personal device as a secondary screen, for that iTV device, in order to capture the viewers' interactions.
-
-A presence service detects physically nearby secondary screen devices that have the services necessary to establish a communication with an iTV device.
-Modern commercial operating systems for personal devices enable service discovery by using peer-to-peer or WiFi communication.
-This feature is particularly appropriated for the TV environment, once the secondary screen device must be wirelessly linked up with the iTV device.
-The devices are synchronized in order to update the contextual data about the viewer profile and the amount of people using a secondary device in the room.
-This contextual data can be used to learn how a viewer behaves towards other viewers.
-For example, a viewer behaves in a particular way while watching TV alone and the same viewer behaves in a different way while watching in group.
 

### 2014-07-02 src/paper.tex
Changes in structure
WIP: Application hosting infrastructure

@@
 Since the TV environment is shared, the data about nearby viewers is captured to contextualize the interaction with the purpose of helping to infer the behaviors of one viewer toward others ones.
+
+The sections of this paper are arranged as follows: \ldots
 
@@
 These client applications also use the web services as an interface to send interaction data to the CP.
-Since applications in the TV domain have the potential to reach out millions of users at the same time, it is interesting to consider the use of a cloud based platform in the CP infrastructure~\cite{Lee2010}.
 
-% descrever a possibilidade de autenticação em outra arte do texto: A possible way to do so is to generate a registration code for each equipment
+% descrever a possibilidade de autenticação em outra parte do texto: A possible way to do so is to generate a registration code for each equipment
 
@@
 
-\subsection{Web Services Design}
+\subsection{Application Hosting Infrastructure}
 
-- RESTful API
+Since applications in the TV domain have the potential to reach out millions of users at the same time, it is interesting to consider the use of a cloud based platform for storing and processing data~\cite{Lee2010}.
 
-- Resources
+The front controller architectural design pattern~\cite{Buschmann2007} is employed to centralize requests processing in a single component.
+Each second screen client sends requests to and receives responses from a single URL, what simplifies application development.
+
+- Web application
 
@@
 
+- presence detector
+
 \subsection{Second Screen Devices}
+
+- web services Consumption
+
+- web services usage
 
@@
 
-\subsection{RESTful API}
+\subsection{RESTful API Design and Implementation}
 

### 2014-07-07 src/paper.tex
WIP: Application Hosting Infrastructure

@@
 
+The application hosting infrastructure (AHI) consists of a server environment that provides the necessary computing resources to run a web application and make it available to client applications.
+This infrastructure can be owned by the CP or it can be hired from a third party service provider.
+As a matter of fact, the web applications hosted in the AHI are responsible for providing the interactive content that can be accessed by the second screen clients.
+In addition, the web applications offer features for storing and processing data.
+
+Cloud computing can be used to quickly scale up and scale down according to workload demands.
 Since applications in the TV domain have the potential to reach out millions of users at the same time, it is interesting to consider the use of a cloud based platform for storing and processing data~\cite{Lee2010}.

### 2014-07-10 src/paper.tex
WIP: Front controller pattern

@@
 As a matter of fact, the web applications hosted in the AHI are responsible for providing the interactive content that can be accessed by the second screen clients.
-In addition, the web applications offer features for storing and processing data.
+In addition, the web applications offer features to store and to process interaction data obtained from the viewers.
 
-Cloud computing can be used to quickly scale up and scale down according to workload demands.
-Since applications in the TV domain have the potential to reach out millions of users at the same time, it is interesting to consider the use of a cloud based platform for storing and processing data~\cite{Lee2010}.
+The TV domain has the potential of facing very different audience ratings and the amount of viewers using a particular iTV service can drastically change during the day.
+So, an important requirement for shared iTV services is to dynamically scale up and down the service according to workload demands.
+Researchers have proposed novel ideas based on cloud-computing environments to tackle this requirement~\cite{Lee2010,Lai2011}.
+Since the architecture presented this paper is intended for a potentially large amount of viewers, the use of a cloud-computing infrastructure is considered to handle the requests from client applications.
 

### 2014-07-10 src/paper.tex
Adjusts in cloud-computing paragraph

@@
 
-The TV domain has the potential of facing very different audience ratings and the amount of viewers using a particular iTV service can drastically change during the day.
-So, an important requirement for shared iTV services is to dynamically scale up and down the service according to workload demands.
+The TV domain has the potential of facing very different audience ratings.
+As a result, the amount of viewers using a particular iTV service can drastically change during the day.
+So, an important requirement for shared iTV services is to dynamically scale up and down the computing resources according to workload demands.
 Researchers have proposed novel ideas based on cloud-computing environments to tackle this requirement~\cite{Lee2010,Lai2011}.
-Since the architecture presented this paper is intended for a potentially large amount of viewers, the use of a cloud-computing infrastructure is considered to handle the requests from client applications.
+Since the architecture presented in this paper is intended to support a potentially large amount of viewers, the use of a cloud-computing infrastructure is considered in order to handle the requests from client applications.
 

### 2014-07-10 src/paper.tex
WIP:Web environment

@@
 
-The front controller architectural design pattern~\cite{Buschmann2007} is employed to centralize requests processing in a single component.
+\subsection{Web Environment}
+\label{ssub_web_environment}
+
+Request handling is carried out by using the front controller architectural design pattern~\cite{Buschmann2007}.
+It is employed to centralize requests processing in a single component.
 Each second screen client sends requests to and receives responses from a single URL, what simplifies application development.

### 2014-07-11 src/paper.tex
WIP: Web application

--- 2014-07-10 src/paper.tex
@@
 
-The front controller architectural design pattern~\cite{Buschmann2007} is employed to centralize requests processing in a single component.
-Each second screen client sends requests to and receives responses from a single URL, what simplifies application development.
+\subsection{Web Application}
+\label{ssub_web_application}
 
-- Web application
+The web applications can be hosted in the AHI
+
+The requests handling is carried out by using the front controller architectural design pattern~\cite{Buschmann2007}.
+It is employed to centralize requests processing in a single component, which facilitates the design and implementation of services for different clients.
+As a result, client applications developed for different operating systems and send requests to and receive response from a single URL.
 

### 2014-07-14 src/paper.tex
Web application section

--- 2014-07-10 src/paper.tex
@@
 
-The front controller architectural design pattern~\cite{Buschmann2007} is employed to centralize requests processing in a single component.
-Each second screen client sends requests to and receives responses from a single URL, what simplifies application development.
+\subsection{Web Application}
+\label{ssub_web_application}
 
-- Web application
+The web applications, which can be hosted in the AHI, comprises the business logic of the second screen application.
+The content that is presented in the second screen client can be requested through HTTP requests to web services designed for this purpose.
+In addition, the data captured by second screen clients in the iTV environment can also be sent to the CP by using HTTP requests.
+
+The HTTP requests handling is carried out by using the front controller architectural design pattern~\cite{Buschmann2007}.
+It is employed to centralize requests processing in a single component, which facilitates the design and implementation of services for different second screen clients.
+As a result, client applications can access the same service even when they are developed for different operating systems in different programming languages.
 
@@
 \subsection{Second Screen Devices}
+
+The second screen devices are used to interact with the content being watched in the TV terminal.
+
+- Presentation of the content
+
+- Capture of viewers interactions (using the WEB service)
 

### 2014-07-14 src/paper.tex
Adjustments in Web application section

@@
 
-The web applications, which can be hosted in the AHI, comprises the business logic of the second screen application.
-The content that is presented in the second screen client can be requested through HTTP requests to web services designed for this purpose.
+The web applications include the business logic, models and controllers that are necessary for the proper functioning of the second screen client application.
+The content that is presented in the client can be requested through HTTP requests to web services designed for this purpose.
 In addition, the data captured by second screen clients in the iTV environment can also be sent to the CP by using HTTP requests.

### 2014-07-15 src/paper.tex
iTV Environment section

@@
 
-- set-top box
-
-- Service discovery
-
-- context synchronization
-
-- presence detector
+The iTV environment consists of an iTV terminal, which can be an set-top box or some iTV terminal with some sort of communication technology e.g., WiFi or Bluetooth.
+This environment has some services to facilitate the communication with second screen devices.
+The first one is the Presence Detector Service (PDS), which is in charge of discovering nearby second screen devices.
+The PDS is connected with another service known as Profile Handler Service (PHS), which aims to perform profile management at each moment when new second screen devices are found in the iTV environment.
 

### 2014-07-15 src/paper.tex
Second screen devices

@@
 
-The second screen devices are used to interact with the content being watched in the TV terminal.
+The second screen devices are clients that communicate with the web application in order to provide interactive capabilities with the content being watched in the TV terminal.
+The communication between second screen clients and the web application is made through HTTP requests to the web services.
+Basically, there are two types of communication between clients and the web application: retrieval the content from the web server and transmission of the captured interactions to the web server.
+
+The content presented on the second screen client, which is related to the TV content, is retrieved from the web server and it is formated to be shown on a specially designed native application.
 

### 2014-07-16 src/paper.tex
Second Screen Devices

@@
 Basically, there are two types of communication between clients and the web application: retrieval the content from the web server and transmission of the captured interactions to the web server.
+Since these types communication occurs between client and a web server, Internet connection is required to accomplish the requests.
 
-The content presented on the second screen client, which is related to the TV content, is retrieved from the web server and it is formated to be shown on a specially designed native application.
+The retrieval of content from the web server happens when the client application performs a specially designed HTTP request aiming to obtain the content from a specified resource.
+The content obtained from this request is then shown by the client application on the second screen device.
 
-- Presentation of the content
-
-- Capture of viewers interactions (using the WEB service)
-
-- web services Consumption
-
-- web services usage
+The interactions captured from the viewers are transmitted to the web server by using the same principle of the retrieval of content.
+A HTTP request is sent to the web service.
+However, in this case, the request is designed to submit the data to a specified resource in order to be processed by the web application.
 

### 2014-07-21 src/paper.tex
Second screen implementation
Web service implementation

@@
 A presence service detects physically nearby second screen devices that have the services necessary to establish a communication with an iTV terminal.
-This feature is particularly appropriated for the TV environment, once second screen devices are usually connected with an iTV terminal via WiFi.
+This feature is particularly appropriate for the TV environment, once second screen devices are usually connected with an iTV terminal via WiFi.
 The devices are synchronized in order to update the contextual data about the viewer profile, amount of people using a second screen devices in the room and content being watched.
@@
 
+% describe a device to device communication
+
 \subsection{Second Screen Devices}
@@
 
-\subsection{RESTful API Design and Implementation}
+The architecture described in this paper combines several concepts and technologies, e.g.
+second screen, interactive television, mobile applications, device to device communication and web technologies.
+An implementation of this architecture was developed using the Java programming language in both server and client.
+However, the proposed approach does not require uniquely the Java programming language and the its use does not subtract the generality of the architecture.
+Indeed, the architecture can be implemented in other languages provided that they are compatible with the development platforms.
+
+\subsection{Web Services API Design and Implementation}
+
+A open source web framework~\cite{Walls2011} was used to develop the server side of the architecture, i.e., the web application.
+This framework facilitates the development of the front controller pattern in web-based application.
+Thus, all the requests are tunneled through a single entry point, which is an efficient way to implement a command-based mechanism.
+This mechanism not only routes and dispatches commands to appropriate handlers, but also exposes a flexible and scalable API structure.
 
@@
 
-\section{Experiments}
+\subsection{Experiment}
+
+ - Content rating systems
 

### 2014-07-21 src/paper.tex
Table with URI resources

@@
 \usepackage{graphicx}
+\usepackage{array}
 
@@
 
+\begin{table}[!t]
+	\label{tab_web_services_design}
+	\renewcommand{\arraystretch}{1.3}
+	\caption{Description of the strategies to handle CRUD by mapping HTTP methods onto URIs.}
+	\centering
+	\begin{tabular}{|l|c|p{3.5cm}|}
+		\hline
+		\bfseries Resource & \bfseries HTTP Method & \bfseries
+		\tabularnewline\hline
+		/resource & GET& Retrieves a list of the specified resource 
+		\tabularnewline\hline
+		/resource/\emph{\{id\}} & GET & Retrieves a specific resource identified by \emph{\{id\}}
+		\tabularnewline\hline
+		/resource & POST & Creates a new resource 
+		\tabularnewline
+		\hline
+		/resource/\emph{\{id\}} & PUT & Updates a resource identified by \{id\}
+		\tabularnewline
+		\hline
+		/resource/\emph{\{id\}} & DELETE & Deletes a resource identified by \emph{\{id\}}
+		\tabularnewline
+		\hline
+	\end{tabular}
+\end{table}
+
 \subsection{Web Service Consumption}

### 2014-07-21 src/paper.tex
Versão quase finalizada

@@
 
-\title{A Second Screen Architecture for Capturing Viewers' Interaction Data in TV Environments}
+\title{Capture of Individualized Viewers' Interactions in Interactive TV Environments}
 \author{Ricardo~Erikson~V.~de~S.~Rosa, 
@@
 
-TV watching is essentially a social activity, where groups of people with a common interest share the same space and TV set for entertainment or information~\cite{Masthoff2004}.
-After many advances in technologies for interactive and digital TV, it is possible to develop high-level applications to enrich the watching experience.
+TV watching is essentially a social activity, where groups of people with a common interest share the same space and TV set for entertainment or information.
+After many advances in technologies for interactive TV (iTV), it is possible to develop high-level applications to enrich the watching experience.
 Thus, the interaction between viewers and TV became more elaborated than simply change channels and adjust sound volume.
-In this scenario, .
+Many iTV products, such as Google TV, Apple TV and smart TVs from several manufactures have contributed significantly to this scenario, since they carry a wide variety of services, e.g., news, games and shopping.
 
-The conventional remote control (RC) is typically seen as the standard medium of interaction among the viewers and their TVs.
-However, this medium presents two notable problems: (1) to precisely and automatically identify the viewers who have possession of the RC; and (2) to capture contextualized viewers' interactions.
-As a result, the entire group of viewers (e.g., a family) is treated as a single viewer and the tastes that were inferred from the users in charge of the RC can be imposed on the tastes of the others.
+The data that is collected from viewers during their interactions with iTV services can provide valuable information for iTV personalization, specially when it is individualized and comes from millions of viewers.
+Unfortunately, the traditional model of interaction between viewer and TV, which uses a conventional remote control (RC), hinders to obtain this individualized data~\cite{Cesar2008}.
+The only existing remote control is shared by many viewers and it is difficult to manage who is in possession of the remote control at a given moment.
+Consequently, it is also difficult to provide high quality personalization, since the tastes that were inferred from the users in charge of the RC can be imposed on the tastes of the others.
 
-Given the limitations of the conventional RC, a novel secondary screen architecture to identify viewers who are interacting with interactive TV (iTV) services is presented in this paper.
-Personal devices (e.g., smartphones and tablets) are used as secondary screens with the aim of capturing personal and contextualized interactions from each active viewer in iTV environments.
-By using proper machine learning algorithms, the personal and contextualized data can reveal important information about viewers' tastes~\cite{Kim2012,Shin2009}.
-
-Since personal devices are almost ubiquitous, they stand out as a powerful mechanism to identify viewers by means of wireless communication technologies~\cite{Cabarcos2011}.
-In the architecture described in this paper, personal devices play an additional role as a medium of interaction since they are easy and intuitive to use in iTV environments, as pointed out in~\cite{Courtois2012}.
-An approach described in~\cite{Teixeira2010} has shown an architecture to capture viewers' interactions through a conventional RC.
-The architecture described in this paper goes further and captures individual and contextualized data.
-Since the TV environment is shared, the data about nearby viewers is captured to contextualize the interaction with the purpose of helping to infer the behaviors of one viewer toward others ones.
-
-The sections of this paper are arranged as follows: \ldots
-
-\section{Background}
-
-The key concepts and some relevant related works that underpin the proposed architecture are described in this section.
-Section~\ref{ssec_ss_usage} presents the usage of second screen devices in iTV environments.
-Section~\ref{ssec_capture_viewers_int_data} describes the current status of research on the capture of viewers' interactions.
-The identification of viewers in iTV environments is described in Section~\ref{ssec_ident_viewers_itv}.
-
-\subsection{Second Screen Usage in iTV}
-\label{ssec_ss_usage}
-
-Nowadays, TV viewers rarely just sit passively and watch TV.
-In many situations, they use second screen devices to perform activities related to TV watching.
-In the context of TV, second screen refers to the use of a computing device, which is usually a smartphone or a tablet PC, to enhance the viewing experience.
-
-Research studies on second screen have been conducted in TV environments since late 1990s.
-In 1996, Robertson~\emph{et~al}.~\cite{Robertson1996} developed a prototype where handheld devices were used to control TV and to provide additional content (e.g., textual information) related to the audio-visual content presented in the TV screen.
-More recent studies show that viewers evaluated positively the use of second screen devices to control the TV and to enrich the TV content~\cite{Cesar2008,Cesar2011,Tsekleves2011}.
-Cesar~\emph{et~al.}~\cite{Cesar2009} contributed identifying and modeling the requirements to implement an interactive TV architecture based on second screen devices.
-
-A review of the related works shown that the usage of personal devices as second screen devices has two main advantages.
-Firstly, the visual elements (e.g., an EPG guide) that are commonly shown on the TV screen can be presented on the second screen device in such a way that the viewing experience of other viewers is not affected.
-Secondly, the second screen device can enhance the capabilities of a conventional remote control by providing more sophisticated interactions, e.g., gesture and voice commands.
-
-\subsection{Capture of Viewers' Interaction Data}
-\label{ssec_capture_viewers_int_data}
-
-The interactions between viewers and TV can be classified as implicit and explicit.
-Implicit interactions are part of the natural actions of watching TV e.g., changing channels and adjust sound volume.
-In explicit interactions, viewers contribute by actively performing actions that are not usual in the their experience from conventional TV.
-Such actions include among other things, evaluating audio-visual content, commenting and sharing opinion about a given content through iTV applications.
-Both kinds of interaction are important for services that rely on audience generated data, e.g., content personalization, targeted advertising or content improvement.
-
-Some researchers proposed mechanisms to capture viewers' interactions in TV environments~\cite{Teixeira2010,Basilio2013,DosSantos2013}.
-Teixeira~\emph{et~al.}~\cite{Teixeira2010} described an infrastructure to capture both implicit and explicit interactions that are performed by means of a conventional remote control.
-The remote control events are recorded in XML format using elements to describe contextual information, implicit and explicit interactions.
-Basílio~\emph{et~al.}~\cite{Basilio2013} described two methods for capturing viewers interactions: (1) one using a middleware extension, which adds new functionalities to an existing standard; and (2) another one using applications that are added to the broadcaster data stream, which requires no changes in the middleware.
-
-Another research study conduced by Santos and Bianchini~\cite{DosSantos2013}, described a advertising system for interactive TV, which uses viewers interactions in order to generate targeted advertising.
-Basically, the described system implicitly captures metadata (e.g., the category of a TV program) information when interactions take place.
-
-The data generated by viewers' interactions can provide valuable information to improve TV personalization, especially when data comes from millions of viewers.
-For instance, viewers can change quickly the channel when a TV program loses their interest.
-Knowing when the channel change occurs can help producers to improve the content in order to increase viewers interest.
-
-\subsection{Viewers' Identification in iTV}
-\label{ssec_ident_viewers_itv}
-
-Since the TV environment can be shared by many viewers (e.g., a family or a group of friends), audience identification becomes an important feature for interactive TV systems.
-Many modern computational systems consist of identifying viewers in order to provide personalized services.
-Unfortunately, the traditional model of interaction, which uses a conventional remote control, hinders the identification of viewers.
-The only existing remote control is shared by many viewers and it is difficult to manage who is in possession of the remote control at a given moment.
-
-Some research studies have addressed automated technologies for viewers' identification in interactive TV environments.
-Cabarcos~\emph{et~al.}~\cite{Cabarcos2011} described an architecture that uses the connectivity capabilities of smartphones to detect viewers nearby the TV terminal at every moment.
-Each smartphone has a hardware identifier(e.g., the Bluetooth MAC address), which is associated to a viewer identifier according to an association table.
-Hwang~\emph{et~al.}~\cite{Hwang2007} presented a method based on face recognition to identify viewers in interactive TV environments.
-The method uses sensors to capture images of the audience and then employs images processing techniques and machine learning algorithm for both detecting and recognizing the faces of the viewers.
+In this paper it is presented a novel secondary screen architecture to capture individualized viewers' interactions with iTV services.
+Personal devices (e.g., smartphones and tablets) are used as secondary screens with the aim of capturing personal interactions from each viewer that is actively interacting with the iTV terminal.In addition, the interactive capabilities of personal devices can provide a powerful interface for users~\cite{Lee2013}.
 
@@
 
-The architecture presented in this paper applies to the capture of audience interactions by means of the use of second screen devices in interactive TV environments.
-Personal devices are used as second screen devices to enrich the interactivity between viewers and TV at the same time that they facilitate the capture of viewers' interactions.
+The overall architecture of the system is presented in \figurename~\ref{fig_architecture}.
+This architecture consists of tree parts that interact with each other: (1) second screen devices, which are the personal devices; an (2) iTV terminal; and (3) the content provider infrastructure.
+The CP infrastructure includes an application hosting infrastructure, which comprises the computing resources (e.g., application server, database services and storage services) that are necessary to host web applications.
+%The dashed arrows represent the data flowing from the viewer to the CP infrastructure, while dotted arrows denote the flow from CP infrastructure to the viewers.
+The continuous arrows represent the data flow inside the CP infrastructure.
 
+% Trocar essa figura
 \begin{figure}[!t]
@@
 	\includegraphics[width=3.5in]{img/architecture-overview.pdf}
-	\caption{Architecture to capture audience interactions. The viewers' interactions and the contextual information are captured and stored in the CP infrastructure.}
+	\caption{Architecture to capture audience interactions. The data is captured and stored in the CP infrastructure.}
 	\label{fig_architecture}
 \end{figure}
-
-An overview of the architecture described in this paper is presented in \figurename~\ref{fig_architecture}.
-This architecture consists of tree subsystems that interact with each other: (1) second screen devices, which are personal devices; an (2) iTV terminal; and (3) the CP infrastructure.
-The CP infrastructure includes an application hosting infrastructure, which comprises the computing resources (e.g., application server, database services and storage services) that are necessary to host web applications.
-The dashed arrows represent the data flowing from the viewer to the CP infrastructure, while dotted arrows denote the flow from CP infrastructure to the viewers.
-The continuous arrows represent the data flow inside the CP infrastructure.
 
@@
 Once the authentication is completed, the second screen device can be used to interact with the iTV terminal.
+By using the authentication mechanism and considering the second screen client as being a personal device, every user interacting with the iTV terminal can be properly identified.
 
@@
 This feature is particularly appropriate for the TV environment, once second screen devices are usually connected with an iTV terminal via WiFi.
-The devices are synchronized in order to update the contextual data about the viewer profile, amount of people using a second screen devices in the room and content being watched.
-This contextual data can be used to learn how a viewer behaves towards other viewers.
-For example, a viewer behaves in a particular way while watching TV alone and the same viewer behaves in a different way while watching in group.
 
@@
 
-% descrever a possibilidade de autenticação em outra parte do texto: A possible way to do so is to generate a registration code for each equipment
+% \section{Second Screen Architecture}
 
-% commentar sobre a sincronização de contexto entre iTV terminals e second screen devices.
-
-\section{Second Screen Architecture}
-
-Viewers can use their secondary screen device to interact with TV by evaluating, commenting, sharing, or even recommending TV content to other viewers.
+% Viewers can use their secondary screen device to interact with TV by evaluating, commenting, sharing, or even recommending TV content to other viewers.
 The data related to those actions are naturally provided by the viewers, as part of the interaction, when using interactive applications (developed by the CP) installed in their secondary screen devices.
@@
 
-The Data Writer service in the application hosting infrastructure consists basically of web services that provide interoperability among secondary screen devices, iTV devices and the CP infrastructure.
+% The Data Writer service in the application hosting infrastructure consists basically of web services that provide interoperability among secondary screen devices, iTV devices and the CP infrastructure.
 The main advantage of this approach is the independence of operating system and programming languages used in the personal devices.
 
-Once the audience data is stored, the CP can use this data to derive valuable information about users' tastes.
+% Once the audience data is stored, the CP can use this data to derive valuable information about users' tastes.
 This information can be used to provide personalized content to viewers based on the current context.
@@
 
-\subsection{Application Hosting Infrastructure}
+% \subsection{Application Hosting Infrastructure}
 
-The application hosting infrastructure (AHI) consists of a server environment that provides the necessary computing resources to run a web application and make it available to client applications.
+% The application hosting infrastructure (AHI) consists of a server environment that provides the necessary computing resources to run a web application and make it available to client applications.
 This infrastructure can be owned by the CP or it can be hired from a third party service provider.
@@
 
-The TV domain has the potential of facing very different audience ratings.
+% The TV domain has the potential of facing very different audience ratings.
 As a result, the amount of viewers using a particular iTV service can drastically change during the day.
@@
 
-\subsection{Web Application}
-\label{ssub_web_application}
+% \subsection{Web Application}
+% \label{ssub_web_application}
 
-The web applications include the business logic, models and controllers that are necessary for the proper functioning of the second screen client application.
+% The web applications include the business logic, models and controllers that are necessary for the proper functioning of the second screen client application.
 The content that is presented in the client can be requested through HTTP requests to web services designed for this purpose.
@@
 
-The HTTP requests handling is carried out by using the front controller architectural design pattern~\cite{Buschmann2007}.
+% The HTTP requests handling is carried out by using the front controller architectural design pattern~\cite{Buschmann2007}.
 It is employed to centralize requests processing in a single component, which facilitates the design and implementation of services for different second screen clients.
@@
 
-\subsection{iTV Environment}
+% \subsection{iTV Environment}
 
-The iTV environment consists of an iTV terminal, which can be an set-top box or some iTV terminal with some sort of communication technology e.g., WiFi or Bluetooth.
+% The iTV environment consists of an iTV terminal, which can be an set-top box or some iTV terminal with some sort of communication technology e.g., WiFi or Bluetooth.
 This environment has some services to facilitate the communication with second screen devices.
@@
 
-% describe a device to device communication
+% % describe a device to device communication
 
-\subsection{Second Screen Devices}
+% \subsection{Second Screen Devices}
 
-The second screen devices are clients that communicate with the web application in order to provide interactive capabilities with the content being watched in the TV terminal.
+% The second screen devices are clients that communicate with the web application in order to provide interactive capabilities with the content being watched in the TV terminal.
 The communication between second screen clients and the web application is made through HTTP requests to the web services.
@@
 
-The retrieval of content from the web server happens when the client application performs a specially designed HTTP request aiming to obtain the content from a specified resource.
+% The retrieval of content from the web server happens when the client application performs a specially designed HTTP request aiming to obtain the content from a specified resource.
 The content obtained from this request is then shown by the client application on the second screen device.
 
-The interactions captured from the viewers are transmitted to the web server by using the same principle of the retrieval of content.
+% The interactions captured from the viewers are transmitted to the web server by using the same principle of the retrieval of content.
 A HTTP request is sent to the web service.
@@
 
-\section{Secondary Screen Implementation}
+\section{Implementation of the iTV Environment using Second Screen}
 
@@
 An implementation of this architecture was developed using the Java programming language in both server and client.
-However, the proposed approach does not require uniquely the Java programming language and the its use does not subtract the generality of the architecture.
+However, the proposed approach does not require uniquely the Java programming language and its use does not subtract the generality of the architecture.
 Indeed, the architecture can be implemented in other languages provided that they are compatible with the development platforms.
 
-\subsection{Web Services API Design and Implementation}
-
 A open source web framework~\cite{Walls2011} was used to develop the server side of the architecture, i.e., the web application.
-This framework facilitates the development of the front controller pattern in web-based application.
+This framework facilitates the development of the front controller pattern in web-based applications.
 Thus, all the requests are tunneled through a single entry point, which is an efficient way to implement a command-based mechanism.
 This mechanism not only routes and dispatches commands to appropriate handlers, but also exposes a flexible and scalable API structure.
+
+Table~\ref{tab_web_services_design} describes the design of the Unified Resource Identifiers (URIs) to provide a mechanism of interaction between second screen devices and the web application.
+The CRUD\footnote{Create, Read, Update and Delete} operations can be requested by requested by the second screen client by using a proper resource and selecting a proper HTTP method.
+Thus, the second screen clients can retrieve the content from the server in order to present personalized content to the client.
 
@@
 	\renewcommand{\arraystretch}{1.3}
-	\caption{Description of the strategies to handle CRUD by mapping HTTP methods onto URIs.}
+	\caption{Description of the strategies to handle CRUD by combining HTTP methods and URIs.}
 	\centering
@@
 		\hline
-		\bfseries Resource & \bfseries HTTP Method & \bfseries
+		\bfseries Resource & \bfseries HTTP Method & \bfseries Description of the operation
 		\tabularnewline\hline
@@
 
-\subsection{Web Service Consumption}
-
-\subsection{iTV and Secondary Screen Devices Communication}
-
-\subsection{Experiment}
-
- - Content rating systems
+Second screen clients can also perform requests to send content to the web-application. The client application can be designed to track user interactions and
 

### 2014-07-22 src/paper.tex
Version submitted to ICCE

@@
 \title{Capture of Individualized Viewers' Interactions in Interactive TV Environments}
-\author{Ricardo~Erikson~V.~de~S.~Rosa, 
+\author{Ricardo~Erikson~V.~de~S.~Rosa~and~ 
 	Vicente~Ferreira~de~Lucena~Jr.,\IEEEmembership{Member,IEEE,}
-	and~Lucas~Carvalho~Cordeiro
 	%
@@
 \thanks{Vicente Ferreira de Lucena Jr. is with the PPGEE, PPGI, and CETELI - Electronics and Information Technology R\&D Center at UFAM, Manaus, Amazonas, Brazil (e-mail: vicente@ufam.edu.br)}%
-\thanks{Lucas Carvalho Cordeiro is with the PPGEE-UFAM, and CETELI - Electronics and Information Technology R\&D Center, Manaus, Amazonas, Brazil (e-mail:lucascordeiro@ufam.edu.br)}}
+}
 
@@
 This architecture uses personal devices as second screen devices with the aim of identifying viewers and capturing their interactions with interactive TV devices.
-An implementation of this architecture is reported as an audio-visual content rating system for interactive TV.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Second screen devices, Interactive Television, Viewers' Interactions
+Interactive TV; Second Screen; Web Applications
 \end{IEEEkeywords}
@@
 
-The data that is collected from viewers during their interactions with iTV services can provide valuable information for iTV personalization, specially when it is individualized and comes from millions of viewers.
+The data that is collected from viewers during their interactions with iTV services can provide valuable information for iTV personalization, specially when it is individualized and comes from millions of viewers~\cite{Teixeira2010}.
 Unfortunately, the traditional model of interaction between viewer and TV, which uses a conventional remote control (RC), hinders to obtain this individualized data~\cite{Cesar2008}.
@@
 The overall architecture of the system is presented in \figurename~\ref{fig_architecture}.
-This architecture consists of tree parts that interact with each other: (1) second screen devices, which are the personal devices; an (2) iTV terminal; and (3) the content provider infrastructure.
-The CP infrastructure includes an application hosting infrastructure, which comprises the computing resources (e.g., application server, database services and storage services) that are necessary to host web applications.
-%The dashed arrows represent the data flowing from the viewer to the CP infrastructure, while dotted arrows denote the flow from CP infrastructure to the viewers.
-The continuous arrows represent the data flow inside the CP infrastructure.
+This architecture consists of tree parts that interact with each other: (1) second screen devices, which are the personal devices; an (2) iTV terminal; and (3) the web application.
+The web application is hosted in the content provider (CP) infrastructure, which comprises the computing resources (e.g., application server, database services and storage services) that are necessary to host web applications.
 
@@
 	\includegraphics[width=3.5in]{img/architecture-overview.pdf}
-	\caption{Architecture to capture audience interactions. The data is captured and stored in the CP infrastructure.}
+	\caption{Architecture to capture viewers' interactions. The data is captured and stored in the CP infrastructure.}
 	\label{fig_architecture}
@@
 
-A web application is hosted in the application hosting infrastructure.
-This application contains web services that are designed for interoperability between the second screen devices, the iTV terminal and the CP infrastructure.
-A authentication mechanism must be provided so that the viewers can connect their second screen devices and interactive TV terminals (e.g., set-top boxes or smart TVs) with an user account.
+The web application contains web services that are designed for interoperability between the second screen devices, the iTV terminal and the CP infrastructure.
+This web application contains business logic for storing and processing the interaction data obtained from the viewers.
+By using the web services, second screen clients can perform requests to the web application with the purpose of obtaining resources and sending data.
+
+In the web application, an authentication mechanism must be provided so that the viewers can connect their second screen devices and iTV terminals (e.g., set-top boxes or smart TVs) with an user account.
 Once the authentication is completed, the second screen device can be used to interact with the iTV terminal.
@@
 
-A presence service detects physically nearby second screen devices that have the services necessary to establish a communication with an iTV terminal.
-This feature is particularly appropriate for the TV environment, once second screen devices are usually connected with an iTV terminal via WiFi.
-
-The second screen devices and iTV terminals have client applications that consume the web services in order to present the content to the viewers.
-These client applications also use the web services as an interface to send interaction data to the CP.
-
-% \section{Second Screen Architecture}
-
-% Viewers can use their secondary screen device to interact with TV by evaluating, commenting, sharing, or even recommending TV content to other viewers.
-The data related to those actions are naturally provided by the viewers, as part of the interaction, when using interactive applications (developed by the CP) installed in their secondary screen devices.
-The interactions and the contextual data are represented by the audience data, which is sent to CP infrastructure through Internet.
-Then, a Data Writer service stores the audience data related to each particular viewer and the iTV device.
-
-% The Data Writer service in the application hosting infrastructure consists basically of web services that provide interoperability among secondary screen devices, iTV devices and the CP infrastructure.
-The main advantage of this approach is the independence of operating system and programming languages used in the personal devices.
-
-% Once the audience data is stored, the CP can use this data to derive valuable information about users' tastes.
-This information can be used to provide personalized content to viewers based on the current context.
-Thus, the personalized content can be delivered via Internet to the secondary screen device or even to the TV, depending on the context.
-
-% \subsection{Application Hosting Infrastructure}
-
-% The application hosting infrastructure (AHI) consists of a server environment that provides the necessary computing resources to run a web application and make it available to client applications.
-This infrastructure can be owned by the CP or it can be hired from a third party service provider.
-As a matter of fact, the web applications hosted in the AHI are responsible for providing the interactive content that can be accessed by the second screen clients.
-In addition, the web applications offer features to store and to process interaction data obtained from the viewers.
-
-% The TV domain has the potential of facing very different audience ratings.
-As a result, the amount of viewers using a particular iTV service can drastically change during the day.
-So, an important requirement for shared iTV services is to dynamically scale up and down the computing resources according to workload demands.
-Researchers have proposed novel ideas based on cloud-computing environments to tackle this requirement~\cite{Lee2010,Lai2011}.
-Since the architecture presented in this paper is intended to support a potentially large amount of viewers, the use of a cloud-computing infrastructure is considered in order to handle the requests from client applications.
-
-% \subsection{Web Application}
-% \label{ssub_web_application}
-
-% The web applications include the business logic, models and controllers that are necessary for the proper functioning of the second screen client application.
-The content that is presented in the client can be requested through HTTP requests to web services designed for this purpose.
-In addition, the data captured by second screen clients in the iTV environment can also be sent to the CP by using HTTP requests.
-
-% The HTTP requests handling is carried out by using the front controller architectural design pattern~\cite{Buschmann2007}.
-It is employed to centralize requests processing in a single component, which facilitates the design and implementation of services for different second screen clients.
-As a result, client applications can access the same service even when they are developed for different operating systems in different programming languages.
-
-% \subsection{iTV Environment}
-
-% The iTV environment consists of an iTV terminal, which can be an set-top box or some iTV terminal with some sort of communication technology e.g., WiFi or Bluetooth.
-This environment has some services to facilitate the communication with second screen devices.
-The first one is the Presence Detector Service (PDS), which is in charge of discovering nearby second screen devices.
-The PDS is connected with another service known as Profile Handler Service (PHS), which aims to perform profile management at each moment when new second screen devices are found in the iTV environment.
-
-% % describe a device to device communication
-
-% \subsection{Second Screen Devices}
-
-% The second screen devices are clients that communicate with the web application in order to provide interactive capabilities with the content being watched in the TV terminal.
-The communication between second screen clients and the web application is made through HTTP requests to the web services.
-Basically, there are two types of communication between clients and the web application: retrieval the content from the web server and transmission of the captured interactions to the web server.
-Since these types communication occurs between client and a web server, Internet connection is required to accomplish the requests.
-
-% The retrieval of content from the web server happens when the client application performs a specially designed HTTP request aiming to obtain the content from a specified resource.
-The content obtained from this request is then shown by the client application on the second screen device.
-
-% The interactions captured from the viewers are transmitted to the web server by using the same principle of the retrieval of content.
-A HTTP request is sent to the web service.
-However, in this case, the request is designed to submit the data to a specified resource in order to be processed by the web application.
+Once the second screen devices are authenticated, a Presence Detector Service (PDS) in the iTV terminal detects physically nearby devices that have the services that are necessary to establish a communication with an iTV terminal.
+A Profile Handler Service (PHS) performs profile management when new second screen devices are found in the iTV environment.
+Thus, only authorized devices can communicate with the iTV terminals.
+This feature is particularly appropriate for the TV environment, since the second screen devices are usually connected with an iTV terminal via a short range wireless communication technology, e.g., WiFi or Bluetooth.
 
@@
 Table~\ref{tab_web_services_design} describes the design of the Unified Resource Identifiers (URIs) to provide a mechanism of interaction between second screen devices and the web application.
-The CRUD\footnote{Create, Read, Update and Delete} operations can be requested by requested by the second screen client by using a proper resource and selecting a proper HTTP method.
-Thus, the second screen clients can retrieve the content from the server in order to present personalized content to the client.
+The CRUD operations can be requested by requested by the second screen client by using a proper resource and selecting a proper HTTP method.
+The HTTP methods that were used for CRUD were the following: POST, used when it is necessary to create a new resource; GET, used when it is needed a response with the resource specified in the request; PUT, employed to update a resource; and DELETE, used to delete a resource.
 
@@
 
-Second screen clients can also perform requests to send content to the web-application. The client application can be designed to track user interactions and
+The second screen clients can retrieve the content from the server in order to present personalized content to the client.
+This content can be obtained by using a GET request to the specified resource.
+Second screen clients can also perform requests to send content to the web-application.
+A special use of this type of request is to track viewers' interactions.
+Every interaction can be sent to the web application for further processing in order to provide content customization.
+
+An important feature of this implementation is the possibility of deployment under a cloud-computing environment to support requests from millions of iTV viewers~\cite{Lee2010}.
+In a cloud-computing environment, the data obtained from viewers can be stored in NoSQL databases, which can provide flexibility and scalability in the second screen implementation.
+
+The communication between second screen devices and iTV terminals can be accomplished by using some communication protocol.
+Usual communication protocols in this type communication include the Anymote protocol, which provides commands for communication between Android devices and iTV terminal, and the push notification mechanism, which enables the communication by means of an intermediary server on the Internet.
 
@@
 
-The use of personal devices as secondary screens arises as a new trend in TV environments.
-This paper presents an architecture to accurately capture audience data by using secondary screen devices and iTV devices.
-This architecture take advantage of the ubiquity and interactive capabilities of secondary screen devices to identify viewers and provide feedback for CPs through viewers' interactions and contextual information.
-The captured data relates to a single viewer in a given context.
-By using proper machine learning algorithms, this data can provide valuable information about viewers.
-Knowing this information can be very useful for CPs to improve their services aiming to increase the audience and to enrich user experience.
-As an example, CPs can deliver personalized content for a single person as well as for groups of people by using content recommendation algorithms.
-
-An implementation of the architecture that is described in this paper is under development in a commercial cloud computing platform, which plays the role of the application hosting infrastructure.
-A web application that consists of a content rating system was developed to store  viewers' evaluations toward audio visual content, and a secondary screen client is under development.
-Some RESTful web services were developed to provide interoperability among the secondary screen devices and the web application.
-Thus, the interactions that are captured during the use of the secondary screen application are sent to the CP by means of HTTP requests to RESTful web services.
-Although the secondary screen application is not fully finished yet, some specific features have been tested to capture viewers' interactions as they occur.
-
-The architecture described in this paper is also technically feasible to implement in Digital TV (DTV) systems and can be particularly interesting for CPs in DTV standards, e.g., the Brazilian (ISDB-TB) and the Japanese(ISDB-T).
-These standards support the use of mobile devices as part of the system.
-Since the mobile devices support the implementation of the DTV middleware, a cooperative exhibition of the DTV application can be managed with the iTV device, thus enabling the capture of viewers' interactions and contextual data.
+An implementation of a second screen architecture is described in this paper.
+It was developed in a commercial cloud computing platform, in which the web application is deployed.
+An experimental application that consists of a content rating system was developed to store  viewers' interactions.
+This application tracks the viewers' behaviors during their interactions and stores their ratings towards audio visual content.
+Some web services were developed to provide interoperability among the secondary screen devices and the web application.
+Thus, the interactions that are captured during the use of the secondary screen application are sent to the web application by means of HTTP requests to the web services.
 

### 2014-08-20 src/paper.tex
Algumas modificações no artigo

@@
-\documentclass[journal]{IEEEtran}
+\documentclass[conference,a4paper]{IEEEtran}
 
@@
 \usepackage{graphicx}
-\usepackage{array}
 
@@
 
-\title{Capture of Individualized Viewers' Interactions in Interactive TV Environments}
-\author{Ricardo~Erikson~V.~de~S.~Rosa~and~ 
-	Vicente~Ferreira~de~Lucena~Jr.,\IEEEmembership{Member,IEEE,}
-	%
-\thanks{Ricardo Erikson V.
-de S.
-Rosa is with the Graduate Program in Electrical Engineering, Federal University of Minas Gerais, Av.
-Antônio Carlos 6627, 31270-901, Belo Horizonte, MG, Brazil(e-mail: ricardoerikson@ufmg.br)}%
-\thanks{Vicente Ferreira de Lucena Jr. is with the PPGEE, PPGI, and CETELI - Electronics and Information Technology R\&D Center at UFAM, Manaus, Amazonas, Brazil (e-mail: vicente@ufam.edu.br)}%
+\title{A Secondary Screen Architecture to Accurately Capture Viewers' Interactions in an iTV Environment}
+
+\author{\IEEEauthorblockN{Ricardo E. V. de S. Rosa}
+\IEEEauthorblockA{Graduate Program in Electrical Engineering \\ Federal University of Minas Gerais \\ Av. Antônio Carlos 6627, 31270-901, \\Belo Horizonte, MG, Brazil\\
+Email: ricardoerikson@ufmg.br}
+\and
+\IEEEauthorblockN{Lucas C. Cordeiro and Vicente F. de Lucena Junior}
+\IEEEauthorblockA{Graduate Program in Electrical Engineering \\
+Federal University of Amazonas \\
+Manaus, AM, Brazil\\
+Email: \{lucascordeiro,vicente\}@ufam.edu.br}
 }
-
 \maketitle
@@
 Advances in TV technology have enabled viewers to actively interact with the TV through interactive applications instead of just passively watching TV.
-Typically, the interactions occur by using a conventional remote control, which is shared by many viewers and hinders the capture of viewers' individual interactions.
+Typically, a conventional remote control is shared by many viewers and it is difficult to accurately capture the individual interaction data from each viewer.
 It is also difficult to identify the context in which the interactions occur, e.g., to manage precisely who is present in the environment and what is being watched on TV.
 In this paper, it is presented a novel architecture that facilitates the capture of viewers' individual interactions and contextual data in shared TV environments.
-This architecture uses personal devices as second screen devices with the aim of identifying viewers and capturing their interactions with interactive TV devices.
+This architecture uses personal devices as secondary screen devices with the aim of identifying viewers and capturing their interactions with interactive TV devices.
+A work in progress implementation of this architecture is briefly reported as an audio-visual content rating system for interactive TV.
 \end{abstract}
-
-\begin{IEEEkeywords}
-Interactive TV; Second Screen; Web Applications
-\end{IEEEkeywords}
 
@@
 TV watching is essentially a social activity, where groups of people with a common interest share the same space and TV set for entertainment or information.
-After many advances in technologies for interactive TV (iTV), it is possible to develop high-level applications to enrich the watching experience.
+After many advances in technologies, for interactive and digital TV, it is possible to develop high-level applications to enrich the watching experience.
 Thus, the interaction between viewers and TV became more elaborated than simply change channels and adjust sound volume.
-Many iTV products, such as Google TV, Apple TV and smart TVs from several manufactures have contributed significantly to this scenario, since they carry a wide variety of services, e.g., news, games and shopping.
+In this scenario, the data generated by the interactions of millions of viewers can provide valuable information for content providers (CPs) to improve their services~\cite{Teixeira2010}.
 
-The data that is collected from viewers during their interactions with iTV services can provide valuable information for iTV personalization, specially when it is individualized and comes from millions of viewers~\cite{Teixeira2010}.
-Unfortunately, the traditional model of interaction between viewer and TV, which uses a conventional remote control (RC), hinders to obtain this individualized data~\cite{Cesar2008}.
-The only existing remote control is shared by many viewers and it is difficult to manage who is in possession of the remote control at a given moment.
-Consequently, it is also difficult to provide high quality personalization, since the tastes that were inferred from the users in charge of the RC can be imposed on the tastes of the others.
+The conventional remote control (RC) is typically seen as the standard medium of interaction among the viewers and their TVs.
+However, this medium presents two notable problems: (1) to precisely and automatically identify the viewers who have possession of the RC; and (2) to capture contextualized viewers' interactions.
+As a result, the entire group of viewers (e.g., a family) is treated as a single viewer and the tastes that were inferred from the users in charge of the RC can be imposed on the tastes of the others.
 
-In this paper it is presented a novel secondary screen architecture to capture individualized viewers' interactions with iTV services.
-Personal devices (e.g., smartphones and tablets) are used as secondary screens with the aim of capturing personal interactions from each viewer that is actively interacting with the iTV terminal.In addition, the interactive capabilities of personal devices can provide a powerful interface for users~\cite{Lee2013}.
+Given the limitations of the conventional RC, a novel secondary screen architecture to identify viewers who are interacting with interactive TV (iTV) services is presented in this paper.
+Personal devices (e.g., smartphones and tablets) are used as secondary screens with the aim of capturing personal and contextualized interactions from each active viewer in iTV environments.
+By using proper machine learning algorithms, the personal and contextualized data can reveal important information about viewers~\cite{Kim2012,Shin2009}.
+Examples of applications of machine learning algorithms for in viewers' data include viewers classification, tastes prediction,
 
-\section{Overview of the Second Screen Architecture}
+Since personal devices are almost ubiquitous, they stand out as a powerful mechanism to identify viewers by means of wireless communication technologies~\cite{Cabarcos2011}.
+In the architecture described in this paper, personal devices play an additional role as a medium of interaction since they are easy and intuitive to use in iTV environments, as pointed out in~\cite{Courtois2012}.
+An approach described in~\cite{Teixeira2010} has shown an architecture to capture viewers' interactions through a conventional RC.
+The architecture described in this paper goes further and captures individual and contextualized data.
+Since the TV environment is shared, the data about nearby viewers is captured to contextualize the interaction with the purpose of helping to infer the behaviors of one viewer toward others ones.
 
-The overall architecture of the system is presented in \figurename~\ref{fig_architecture}.
-This architecture consists of tree parts that interact with each other: (1) second screen devices, which are the personal devices; an (2) iTV terminal; and (3) the web application.
-The web application is hosted in the content provider (CP) infrastructure, which comprises the computing resources (e.g., application server, database services and storage services) that are necessary to host web applications.
+\section{Secondary Screen Architecture}
 
-% Trocar essa figura
+Initially, the CP infrastructure must provide an authentication mechanism so that viewers and their interactive TV (iTV) devices (e.g., set-top boxes or smart TVs) can be authenticated by the CP in order to use the secondary screen services.
+Once one iTV device is authenticated, the viewers in the TV environment must link their user accounts up with the iTV device.
+As a result, the viewer can use their personal device as a secondary screen, for that iTV device, in order to capture the viewers' interactions.
+
+The architecture to precisely capture viewers interactions and the context is shown in~\figurename~\ref{fig_architecture}.
+This architecture consists of tree subsystems that interact with each other: (1) secondary screen devices, which might be implemented in personal devices; an (2) iTV device; and (3) the CP infrastructure, which includes an application hosting infrastructure.
+The dashed arrows represent the data flowing from the viewer to the CP infrastructure, while dotted arrows denote the flow from CP infrastructure to the viewers.
+The continuous arrows represent the data flow inside the CP infrastructure.
+Since applications in the TV domain have the potential to reach out millions of users at the same time, it is interesting to consider the use of a cloud based platform in the CP infrastructure~\cite{Lee2010}.
+
 \begin{figure}[!t]
 	\centering
-	\includegraphics[width=3.5in]{img/architecture-overview.pdf}
-	\caption{Architecture to capture viewers' interactions. The data is captured and stored in the CP infrastructure.}
+	\includegraphics[width=3.5in]{img/architecture.pdf}
+	\caption{Architecture to capture audience data. The viewers' interactions and the contextual information are captured and stored in the CP infrastructure.}
 	\label{fig_architecture}
@@
 
-The web application contains web services that are designed for interoperability between the second screen devices, the iTV terminal and the CP infrastructure.
-This web application contains business logic for storing and processing the interaction data obtained from the viewers.
-By using the web services, second screen clients can perform requests to the web application with the purpose of obtaining resources and sending data.
+A presence service detects physically nearby secondary screen devices that have the services necessary to establish a communication with an iTV device.
+Modern commercial operating systems for personal devices enable service discovery by using peer-to-peer or WiFi communication.
+This feature is particularly appropriated for the TV environment, once the secondary screen device must be wirelessly linked up with the iTV device.
+The devices are synchronized in order to update the contextual data about the viewer profile and the amount of people using a secondary device in the room.
+This contextual data can be used to learn how a viewer behaves towards other viewers.
 
-In the web application, an authentication mechanism must be provided so that the viewers can connect their second screen devices and iTV terminals (e.g., set-top boxes or smart TVs) with an user account.
-Once the authentication is completed, the second screen device can be used to interact with the iTV terminal.
-By using the authentication mechanism and considering the second screen client as being a personal device, every user interacting with the iTV terminal can be properly identified.
+Viewers can use their secondary screen device to interact with TV by evaluating, commenting, sharing, or even recommending TV content to other viewers.
+The data related to those actions are naturally provided by the viewers, as part of the interaction, when using interactive applications (developed by the CP) installed in their secondary screen devices.
+The interactions and the contextual data are represented by the audience data, which is sent to CP infrastructure through Internet.
+Then, a Data Writer service stores the audience data related to each particular viewer and the iTV device.
 
-Once the second screen devices are authenticated, a Presence Detector Service (PDS) in the iTV terminal detects physically nearby devices that have the services that are necessary to establish a communication with an iTV terminal.
-A Profile Handler Service (PHS) performs profile management when new second screen devices are found in the iTV environment.
-Thus, only authorized devices can communicate with the iTV terminals.
-This feature is particularly appropriate for the TV environment, since the second screen devices are usually connected with an iTV terminal via a short range wireless communication technology, e.g., WiFi or Bluetooth.
+The Data Writer service in the application hosting infrastructure consists basically of web services that provide interoperability among secondary screen devices, iTV devices and the CP infrastructure.
+The main advantage of this approach is the independence of operating system and programming languages used in the personal devices.
 
-\section{Implementation of the iTV Environment using Second Screen}
-
-The architecture described in this paper combines several concepts and technologies, e.g.
-second screen, interactive television, mobile applications, device to device communication and web technologies.
-An implementation of this architecture was developed using the Java programming language in both server and client.
-However, the proposed approach does not require uniquely the Java programming language and its use does not subtract the generality of the architecture.
-Indeed, the architecture can be implemented in other languages provided that they are compatible with the development platforms.
-
-A open source web framework~\cite{Walls2011} was used to develop the server side of the architecture, i.e., the web application.
-This framework facilitates the development of the front controller pattern in web-based applications.
-Thus, all the requests are tunneled through a single entry point, which is an efficient way to implement a command-based mechanism.
-This mechanism not only routes and dispatches commands to appropriate handlers, but also exposes a flexible and scalable API structure.
-
-Table~\ref{tab_web_services_design} describes the design of the Unified Resource Identifiers (URIs) to provide a mechanism of interaction between second screen devices and the web application.
-The CRUD operations can be requested by requested by the second screen client by using a proper resource and selecting a proper HTTP method.
-The HTTP methods that were used for CRUD were the following: POST, used when it is necessary to create a new resource; GET, used when it is needed a response with the resource specified in the request; PUT, employed to update a resource; and DELETE, used to delete a resource.
-
-\begin{table}[!t]
-	\label{tab_web_services_design}
-	\renewcommand{\arraystretch}{1.3}
-	\caption{Description of the strategies to handle CRUD by combining HTTP methods and URIs.}
-	\centering
-	\begin{tabular}{|l|c|p{3.5cm}|}
-		\hline
-		\bfseries Resource & \bfseries HTTP Method & \bfseries Description of the operation
-		\tabularnewline\hline
-		/resource & GET& Retrieves a list of the specified resource 
-		\tabularnewline\hline
-		/resource/\emph{\{id\}} & GET & Retrieves a specific resource identified by \emph{\{id\}}
-		\tabularnewline\hline
-		/resource & POST & Creates a new resource 
-		\tabularnewline
-		\hline
-		/resource/\emph{\{id\}} & PUT & Updates a resource identified by \{id\}
-		\tabularnewline
-		\hline
-		/resource/\emph{\{id\}} & DELETE & Deletes a resource identified by \emph{\{id\}}
-		\tabularnewline
-		\hline
-	\end{tabular}
-\end{table}
-
-The second screen clients can retrieve the content from the server in order to present personalized content to the client.
-This content can be obtained by using a GET request to the specified resource.
-Second screen clients can also perform requests to send content to the web-application.
-A special use of this type of request is to track viewers' interactions.
-Every interaction can be sent to the web application for further processing in order to provide content customization.
-
-An important feature of this implementation is the possibility of deployment under a cloud-computing environment to support requests from millions of iTV viewers~\cite{Lee2010}.
-In a cloud-computing environment, the data obtained from viewers can be stored in NoSQL databases, which can provide flexibility and scalability in the second screen implementation.
-
-The communication between second screen devices and iTV terminals can be accomplished by using some communication protocol.
-Usual communication protocols in this type communication include the Anymote protocol, which provides commands for communication between Android devices and iTV terminal, and the push notification mechanism, which enables the communication by means of an intermediary server on the Internet.
+Once the audience data is stored, the CP can use this data to derive valuable information about users' tastes.
+This information can be used to provide personalized content to viewers based on the current context.
+Thus, the personalized content can be delivered via Internet to the secondary screen device or even to the TV, depending on the context.
 
@@
 
-An implementation of a second screen architecture is described in this paper.
-It was developed in a commercial cloud computing platform, in which the web application is deployed.
-An experimental application that consists of a content rating system was developed to store  viewers' interactions.
-This application tracks the viewers' behaviors during their interactions and stores their ratings towards audio visual content.
-Some web services were developed to provide interoperability among the secondary screen devices and the web application.
-Thus, the interactions that are captured during the use of the secondary screen application are sent to the web application by means of HTTP requests to the web services.
+The use of personal devices as secondary screens arises as a new trend in TV environments.
+This paper presents an architecture to accurately capture audience data by using secondary screen devices and iTV devices.
+This architecture take advantage of the ubiquity and interactive capabilities of secondary screen devices to identify viewers and provide feedback for CPs through viewers' interactions and contextual information.
+The captured data relates to a single viewer in a given context.
+By using proper machine learning algorithms, this data can provide valuable information about viewers.
+Knowing this information can be very useful for CPs to improve their services aiming to increase the audience and to enrich user experience.
+As an example, CPs can deliver personalized content for a single person as well as for groups of people by using content recommendation algorithms.
+
+An implementation of the architecture that is described in this paper is under development in a commercial cloud computing platform, which plays the role of the application hosting infrastructure.
+A web application that consists of a content rating system was developed to store  viewers' evaluations toward audio visual content, and a secondary screen client is under development.
+Some RESTful web services were developed to provide interoperability among the secondary screen devices and the web application.
+Thus, the interactions that are captured during the use of the secondary screen application are sent to the CP by means of HTTP requests to RESTful web services.
+Although the secondary screen application is not fully finished yet, some specific features have been tested to capture viewers' interactions as they occur.
+
+The architecture described in this paper is also technically feasible to implement in Digital TV (DTV) systems and can be particularly interesting for CPs in DTV standards, e.g., the Brazilian (ISDB-TB) and the Japanese(ISDB-T).
+These standards support the use of mobile devices as part of the system.
+Since the mobile devices support the implementation of the DTV middleware, a cooperative exhibition of the DTV application can be managed with the iTV device, thus enabling the capture of viewers' interactions and contextual data.
 

### 2014-08-20 src/paper.tex
Versão enviada para a revisão do Lucas

@@
 \newcommand{\blurb}[1]{\marginpar{{\tt !!!}}{\tt [... #1 ...]}}
+\renewcommand{\baselinestretch}{0.99}
 
@@
 By using proper machine learning algorithms, the personal and contextualized data can reveal important information about viewers~\cite{Kim2012,Shin2009}.
-Examples of applications of machine learning algorithms for in viewers' data include viewers classification, tastes prediction,
+Examples of applications of machine learning algorithms in viewers' data include viewers classification, tastes prediction, targeted advertising and content recommendation.
 
@@
 
-An implementation of the architecture that is described in this paper is under development in a commercial cloud computing platform, which plays the role of the application hosting infrastructure.
+An implementation of the architecture that is described in this paper a commercial cloud computing platform, which plays the role of the application hosting infrastructure.
 A web application that consists of a content rating system was developed to store  viewers' evaluations toward audio visual content, and a secondary screen client is under development.

### 2014-08-21 src/paper.tex
minor changes

--- 2014-08-20 src/paper.tex
@@
 \begin{abstract}
-Advances in TV technology have enabled viewers to actively interact with the TV through interactive applications instead of just passively watching TV.
+Advances in TV technology have enabled viewers to actively interact with the TV via interactive applications instead of just passively watching TV.
 Typically, a conventional remote control is shared by many viewers and it is difficult to accurately capture the individual interaction data from each viewer.
@@
 Personal devices (e.g., smartphones and tablets) are used as secondary screens with the aim of capturing personal and contextualized interactions from each active viewer in iTV environments.
-By using proper machine learning algorithms, the personal and contextualized data can reveal important information about viewers~\cite{Kim2012,Shin2009}.
-Examples of applications of machine learning algorithms for in viewers' data include viewers classification, tastes prediction,
+By using proper machine learning algorithms, the personal and contextualized data can reveal important information about viewers' tastes~\cite{Kim2012,Shin2009}.
 
@@
 In the architecture described in this paper, personal devices play an additional role as a medium of interaction since they are easy and intuitive to use in iTV environments, as pointed out in~\cite{Courtois2012}.
-An approach described in~\cite{Teixeira2010} has shown an architecture to capture viewers' interactions through a conventional RC.
+An approach described in~\cite{Teixeira2010} has shown an architecture to capture viewers' interactions via a conventional RC.
 The architecture described in this paper goes further and captures individual and contextualized data.
@@
 The architecture to precisely capture viewers interactions and the context is shown in~\figurename~\ref{fig_architecture}.
-This architecture consists of tree subsystems that interact with each other: (1) secondary screen devices, which might be implemented in personal devices; an (2) iTV device; and (3) the CP infrastructure, which includes an application hosting infrastructure.
+This architecture consists of tree subsystems that interact with each other: (1) secondary screen devices, which might be implemented in personal devices; (2) an iTV device; and (3) the CP infrastructure, which includes an application hosting infrastructure.
 The dashed arrows represent the data flowing from the viewer to the CP infrastructure, while dotted arrows denote the flow from CP infrastructure to the viewers.
@@
 This contextual data can be used to learn how a viewer behaves towards other viewers.
+For example, a viewer behaves in a particular way while watching TV alone and the same viewer behaves in a different way while watching TV in group.
 
@@
 The data related to those actions are naturally provided by the viewers, as part of the interaction, when using interactive applications (developed by the CP) installed in their secondary screen devices.
-The interactions and the contextual data are represented by the audience data, which is sent to CP infrastructure through Internet.
+The interactions and the contextual data are represented by the audience data, which is sent to CP infrastructure through the Internet.
 Then, a Data Writer service stores the audience data related to each particular viewer and the iTV device.
@@
 This information can be used to provide personalized content to viewers based on the current context.
-Thus, the personalized content can be delivered via Internet to the secondary screen device or even to the TV, depending on the context.
+Thus, the personalized content can be delivered through the Internet to the secondary screen device or even to the TV, depending on the context.
 
@@
 This paper presents an architecture to accurately capture audience data by using secondary screen devices and iTV devices.
-This architecture take advantage of the ubiquity and interactive capabilities of secondary screen devices to identify viewers and provide feedback for CPs through viewers' interactions and contextual information.
+This architecture takes advantage of the ubiquity and interactive capabilities of secondary screen devices to identify viewers and provide feedback for CPs through viewers' interactions and contextual information.
 The captured data relates to a single viewer in a given context.
 By using proper machine learning algorithms, this data can provide valuable information about viewers.
-Knowing this information can be very useful for CPs to improve their services aiming to increase the audience and to enrich user experience.
+Knowing this information can be very useful for CPs to improve their services, aiming to increase the audience and to enrich user experience.
 As an example, CPs can deliver personalized content for a single person as well as for groups of people by using content recommendation algorithms.
@@
 An implementation of the architecture that is described in this paper is under development in a commercial cloud computing platform, which plays the role of the application hosting infrastructure.
-A web application that consists of a content rating system was developed to store  viewers' evaluations toward audio visual content, and a secondary screen client is under development.
+A web application that consists of a content rating system was developed to store  viewers' evaluations toward audio visual content; a secondary screen client is under development.
 Some RESTful web services were developed to provide interoperability among the secondary screen devices and the web application.
@@
 
-The architecture described in this paper is also technically feasible to implement in Digital TV (DTV) systems and can be particularly interesting for CPs in DTV standards, e.g., the Brazilian (ISDB-TB) and the Japanese(ISDB-T).
+The architecture described in this paper is also technically feasible to implement in Digital TV (DTV) systems and can be particularly interesting for CPs in DTV standards, e.g., the Brazilian (ISDB-TB) and the Japanese (ISDB-T).
 These standards support the use of mobile devices as part of the system.

### 2014-08-22 src/paper.tex
Agradecimentos adicionados no paper
Modificação no final da conclusão para ajustar o comprimento do texto

--- 2014-08-20 src/paper.tex
@@
 \documentclass[conference,a4paper]{IEEEtran}
+\IEEEoverridecommandlockouts
 
@@
 \newcommand{\blurb}[1]{\marginpar{{\tt !!!}}{\tt [... #1 ...]}}
-\renewcommand{\baselinestretch}{0.99}
+\renewcommand{\baselinestretch}{0.98}
 
@@
 \author{\IEEEauthorblockN{Ricardo E. V. de S. Rosa}
-\IEEEauthorblockA{Graduate Program in Electrical Engineering \\ Federal University of Minas Gerais \\ Av. Antônio Carlos 6627, 31270-901, \\Belo Horizonte, MG, Brazil\\
+\thanks{This research was supported by SAMSUNG, CAPES, FAPEAM, and CNPq grants.}\IEEEauthorblockA{Graduate Program in Electrical Engineering \\ Federal University of Minas Gerais \\ Av.
+Antônio Carlos 6627, 31270-901, \\Belo Horizonte, MG, Brazil\\
 Email: ricardoerikson@ufmg.br}
@@
 }
+
 \maketitle
@@
 
-A presence service detects physically nearby secondary screen devices that have the services necessary to establish a communication with an iTV device.
-Modern commercial operating systems for personal devices enable service discovery by using peer-to-peer or WiFi communication.
+A presence service detects physically nearby secondary screen devices that are able to establish a communication with an iTV device.
+Modern operating systems for personal devices enable service discovery by using peer-to-peer or WiFi communication.
 This feature is particularly appropriated for the TV environment, once the secondary screen device must be wirelessly linked up with the iTV device.
@@
 
-The architecture described in this paper is also technically feasible to implement in Digital TV (DTV) systems and can be particularly interesting for CPs in DTV standards, e.g., the Brazilian (ISDB-TB) and the Japanese(ISDB-T).
+The architecture described in this paper is also technically feasible to be implemented in Digital TV (DTV) systems and can be particularly interesting for content providers working with the Brazilian (ISDB-TB) or the Japanese(ISDB-T) DTV standard.
 These standards support the use of mobile devices as part of the system.

### 2014-08-25 src/paper.tex
Modificações na seção web sevices api design and implementation

--- 2014-08-20 src/paper.tex
@@
 \newcommand{\blurb}[1]{\marginpar{{\tt !!!}}{\tt [... #1 ...]}}
-\renewcommand{\baselinestretch}{0.99}
 
@@
 
-\title{A Secondary Screen Architecture to Accurately Capture Viewers' Interactions in an iTV Environment}
+\title{A Second Screen Architecture for Capturing Viewers' Interaction Data in TV Environments}
+\author{Ricardo~Erikson~V.~de~S.~Rosa, 
+	Vicente~Ferreira~de~Lucena~Jr.,\IEEEmembership{Member,IEEE,}
+	and~Lucas~Carvalho~Cordeiro
+	%
+\thanks{Ricardo Erikson V.
+de S.
+Rosa is with the Graduate Program in Electrical Engineering, Federal University of Minas Gerais, Av.
+Antônio Carlos 6627, 31270-901, Belo Horizonte, MG, Brazil(e-mail: ricardoerikson@ufmg.br)}%
+\thanks{Vicente Ferreira de Lucena Jr. is with the PPGEE, PPGI, and CETELI - Electronics and Information Technology R\&D Center at UFAM, Manaus, Amazonas, Brazil (e-mail: vicente@ufam.edu.br)}%
+\thanks{Lucas Carvalho Cordeiro is with the PPGEE-UFAM, and CETELI - Electronics and Information Technology R\&D Center, Manaus, Amazonas, Brazil (e-mail:lucascordeiro@ufam.edu.br)}}
 
-\author{\IEEEauthorblockN{Ricardo E. V. de S. Rosa}
-\IEEEauthorblockA{Graduate Program in Electrical Engineering \\ Federal University of Minas Gerais \\ Av. Antônio Carlos 6627, 31270-901, \\Belo Horizonte, MG, Brazil\\
-Email: ricardoerikson@ufmg.br}
-\and
-\IEEEauthorblockN{Lucas C. Cordeiro and Vicente F. de Lucena Junior}
-\IEEEauthorblockA{Graduate Program in Electrical Engineering \\
-Federal University of Amazonas \\
-Manaus, AM, Brazil\\
-Email: \{lucascordeiro,vicente\}@ufam.edu.br}
-}
 \maketitle
@@
 Advances in TV technology have enabled viewers to actively interact with the TV through interactive applications instead of just passively watching TV.
-Typically, a conventional remote control is shared by many viewers and it is difficult to accurately capture the individual interaction data from each viewer.
+Typically, the interactions occur by using a conventional remote control, which is shared by many viewers and hinders the capture of viewers' individual interactions.
 It is also difficult to identify the context in which the interactions occur, e.g., to manage precisely who is present in the environment and what is being watched on TV.
 In this paper, it is presented a novel architecture that facilitates the capture of viewers' individual interactions and contextual data in shared TV environments.
-This architecture uses personal devices as secondary screen devices with the aim of identifying viewers and capturing their interactions with interactive TV devices.
-A work in progress implementation of this architecture is briefly reported as an audio-visual content rating system for interactive TV.
+This architecture uses personal devices as second screen devices with the aim of identifying viewers and capturing their interactions with interactive TV devices.
+An implementation of this architecture is reported as an audio-visual content rating system for interactive TV.
 \end{abstract}
+
+\begin{IEEEkeywords}
+Second screen devices, Interactive Television, Viewers' Interactions
+\end{IEEEkeywords}
 
@@
 
-TV watching is essentially a social activity, where groups of people with a common interest share the same space and TV set for entertainment or information.
-After many advances in technologies, for interactive and digital TV, it is possible to develop high-level applications to enrich the watching experience.
+TV watching is essentially a social activity, where groups of people with a common interest share the same space and TV set for entertainment or information~\cite{Masthoff2004}.
+After many advances in technologies for interactive and digital TV, it is possible to develop high-level applications to enrich the watching experience.
 Thus, the interaction between viewers and TV became more elaborated than simply change channels and adjust sound volume.
-In this scenario, the data generated by the interactions of millions of viewers can provide valuable information for content providers (CPs) to improve their services~\cite{Teixeira2010}.
+In this scenario, .
 
@@
 Personal devices (e.g., smartphones and tablets) are used as secondary screens with the aim of capturing personal and contextualized interactions from each active viewer in iTV environments.
-By using proper machine learning algorithms, the personal and contextualized data can reveal important information about viewers~\cite{Kim2012,Shin2009}.
-Examples of applications of machine learning algorithms in viewers' data include viewers classification, tastes prediction, targeted advertising and content recommendation.
+By using proper machine learning algorithms, the personal and contextualized data can reveal important information about viewers' tastes~\cite{Kim2012,Shin2009}.
 
@@
 
-\section{Secondary Screen Architecture}
+The sections of this paper are arranged as follows: \ldots
 
-Initially, the CP infrastructure must provide an authentication mechanism so that viewers and their interactive TV (iTV) devices (e.g., set-top boxes or smart TVs) can be authenticated by the CP in order to use the secondary screen services.
-Once one iTV device is authenticated, the viewers in the TV environment must link their user accounts up with the iTV device.
-As a result, the viewer can use their personal device as a secondary screen, for that iTV device, in order to capture the viewers' interactions.
+\section{Background}
 
-The architecture to precisely capture viewers interactions and the context is shown in~\figurename~\ref{fig_architecture}.
-This architecture consists of tree subsystems that interact with each other: (1) secondary screen devices, which might be implemented in personal devices; an (2) iTV device; and (3) the CP infrastructure, which includes an application hosting infrastructure.
-The dashed arrows represent the data flowing from the viewer to the CP infrastructure, while dotted arrows denote the flow from CP infrastructure to the viewers.
-The continuous arrows represent the data flow inside the CP infrastructure.
-Since applications in the TV domain have the potential to reach out millions of users at the same time, it is interesting to consider the use of a cloud based platform in the CP infrastructure~\cite{Lee2010}.
+The key concepts and some relevant related works that underpin the proposed architecture are described in this section.
+Section~\ref{ssec_ss_usage} presents the usage of second screen devices in iTV environments.
+Section~\ref{ssec_capture_viewers_int_data} describes the current status of research on the capture of viewers' interactions.
+The identification of viewers in iTV environments is described in Section~\ref{ssec_ident_viewers_itv}.
+
+\subsection{Second Screen Usage in iTV}
+\label{ssec_ss_usage}
+
+Nowadays, TV viewers rarely just sit passively and watch TV.
+In many situations, they use second screen devices to perform activities related to TV watching.
+In the context of TV, second screen refers to the use of a computing device, which is usually a smartphone or a tablet PC, to enhance the viewing experience.
+
+Research studies on second screen have been conducted in TV environments since late 1990s.
+In 1996, Robertson~\emph{et~al}.~\cite{Robertson1996} developed a prototype where handheld devices were used to control TV and to provide additional content (e.g., textual information) related to the audio-visual content presented in the TV screen.
+More recent studies show that viewers evaluated positively the use of second screen devices to control the TV and to enrich the TV content~\cite{Cesar2008,Cesar2011,Tsekleves2011}.
+Cesar~\emph{et~al.}~\cite{Cesar2009} contributed identifying and modeling the requirements to implement an interactive TV architecture based on second screen devices.
+
+A review of the related works shown that the usage of personal devices as second screen devices has two main advantages.
+Firstly, the visual elements (e.g., an EPG guide) that are commonly shown on the TV screen can be presented on the second screen device in such a way that the viewing experience of other viewers is not affected.
+Secondly, the second screen device can enhance the capabilities of a conventional remote control by providing more sophisticated interactions, e.g., gesture and voice commands.
+
+\subsection{Capture of Viewers' Interaction Data}
+\label{ssec_capture_viewers_int_data}
+
+The interactions between viewers and TV can be classified as implicit and explicit.
+Implicit interactions are part of the natural actions of watching TV e.g., changing channels and adjust sound volume.
+In explicit interactions, viewers contribute by actively performing actions that are not usual in the their experience from conventional TV.
+Such actions include among other things, evaluating audio-visual content, commenting and sharing opinion about a given content through iTV applications.
+Both kinds of interaction are important for services that rely on audience generated data, e.g., content personalization, targeted advertising or content improvement.
+
+Some researchers proposed mechanisms to capture viewers' interactions in TV environments~\cite{Teixeira2010,Basilio2013,DosSantos2013}.
+Teixeira~\emph{et~al.}~\cite{Teixeira2010} described an infrastructure to capture both implicit and explicit interactions that are performed by means of a conventional remote control.
+The remote control events are recorded in XML format using elements to describe contextual information, implicit and explicit interactions.
+Basílio~\emph{et~al.}~\cite{Basilio2013} described two methods for capturing viewers interactions: (1) one using a middleware extension, which adds new functionalities to an existing standard; and (2) another one using applications that are added to the broadcaster data stream, which requires no changes in the middleware.
+
+Another research study conduced by Santos and Bianchini~\cite{DosSantos2013}, described a advertising system for interactive TV, which uses viewers interactions in order to generate targeted advertising.
+Basically, the described system implicitly captures metadata (e.g., the category of a TV program) information when interactions take place.
+
+The data generated by viewers' interactions can provide valuable information to improve TV personalization, especially when data comes from millions of viewers.
+For instance, viewers can change quickly the channel when a TV program loses their interest.
+Knowing when the channel change occurs can help producers to improve the content in order to increase viewers interest.
+
+\subsection{Viewers' Identification in iTV}
+\label{ssec_ident_viewers_itv}
+
+Since the TV environment can be shared by many viewers (e.g., a family or a group of friends), audience identification becomes an important feature for interactive TV systems.
+Many modern computational systems consist of identifying viewers in order to provide personalized services.
+Unfortunately, the traditional model of interaction, which uses a conventional remote control, hinders the identification of viewers.
+The only existing remote control is shared by many viewers and it is difficult to manage who is in possession of the remote control at a given moment.
+
+Some research studies have addressed automated technologies for viewers' identification in interactive TV environments.
+Cabarcos~\emph{et~al.}~\cite{Cabarcos2011} described an architecture that uses the connectivity capabilities of smartphones to detect viewers nearby the TV terminal at every moment.
+Each smartphone has a hardware identifier(e.g., the Bluetooth MAC address), which is associated to a viewer identifier according to an association table.
+Hwang~\emph{et~al.}~\cite{Hwang2007} presented a method based on face recognition to identify viewers in interactive TV environments.
+The method uses sensors to capture images of the audience and then employs images processing techniques and machine learning algorithm for both detecting and recognizing the faces of the viewers.
+
+\section{Overview of the Second Screen Architecture}
+
+The architecture presented in this paper applies to the capture of audience interactions by means of the use of second screen devices in interactive TV environments.
+Personal devices are used as second screen devices to enrich the interactivity between viewers and TV at the same time that they facilitate the capture of viewers' interactions.
 
@@
 	\centering
-	\includegraphics[width=3.5in]{img/architecture.pdf}
-	\caption{Architecture to capture audience data. The viewers' interactions and the contextual information are captured and stored in the CP infrastructure.}
+	\includegraphics[width=3.5in]{img/architecture-overview.pdf}
+	\caption{Architecture to capture audience interactions. The viewers' interactions and the contextual information are captured and stored in the CP infrastructure.}
 	\label{fig_architecture}
@@
 
-A presence service detects physically nearby secondary screen devices that have the services necessary to establish a communication with an iTV device.
-Modern commercial operating systems for personal devices enable service discovery by using peer-to-peer or WiFi communication.
-This feature is particularly appropriated for the TV environment, once the secondary screen device must be wirelessly linked up with the iTV device.
-The devices are synchronized in order to update the contextual data about the viewer profile and the amount of people using a secondary device in the room.
+An overview of the architecture described in this paper is presented in \figurename~\ref{fig_architecture}.
+This architecture consists of tree subsystems that interact with each other: (1) second screen devices, which are personal devices; an (2) iTV terminal; and (3) the CP infrastructure.
+The CP infrastructure includes an application hosting infrastructure, which comprises the computing resources (e.g., application server, database services and storage services) that are necessary to host web applications.
+The dashed arrows represent the data flowing from the viewer to the CP infrastructure, while dotted arrows denote the flow from CP infrastructure to the viewers.
+The continuous arrows represent the data flow inside the CP infrastructure.
+
+A web application is hosted in the application hosting infrastructure.
+This application contains web services that are designed for interoperability between the second screen devices, the iTV terminal and the CP infrastructure.
+A authentication mechanism must be provided so that the viewers can connect their second screen devices and interactive TV terminals (e.g., set-top boxes or smart TVs) with an user account.
+Once the authentication is completed, the second screen device can be used to interact with the iTV terminal.
+
+A presence service detects physically nearby second screen devices that have the services necessary to establish a communication with an iTV terminal.
+This feature is particularly appropriate for the TV environment, once second screen devices are usually connected with an iTV terminal via WiFi.
+The devices are synchronized in order to update the contextual data about the viewer profile, amount of people using a second screen devices in the room and content being watched.
 This contextual data can be used to learn how a viewer behaves towards other viewers.
+For example, a viewer behaves in a particular way while watching TV alone and the same viewer behaves in a different way while watching in group.
+
+The second screen devices and iTV terminals have client applications that consume the web services in order to present the content to the viewers.
+These client applications also use the web services as an interface to send interaction data to the CP.
+
+% descrever a possibilidade de autenticação em outra parte do texto: A possible way to do so is to generate a registration code for each equipment
+
+% commentar sobre a sincronização de contexto entre iTV terminals e second screen devices.
+
+\section{Second Screen Architecture}
 
@@
 
+\subsection{Application Hosting Infrastructure}
+
+The application hosting infrastructure (AHI) consists of a server environment that provides the necessary computing resources to run a web application and make it available to client applications.
+This infrastructure can be owned by the CP or it can be hired from a third party service provider.
+As a matter of fact, the web applications hosted in the AHI are responsible for providing the interactive content that can be accessed by the second screen clients.
+In addition, the web applications offer features to store and to process interaction data obtained from the viewers.
+
+The TV domain has the potential of facing very different audience ratings.
+As a result, the amount of viewers using a particular iTV service can drastically change during the day.
+So, an important requirement for shared iTV services is to dynamically scale up and down the computing resources according to workload demands.
+Researchers have proposed novel ideas based on cloud-computing environments to tackle this requirement~\cite{Lee2010,Lai2011}.
+Since the architecture presented in this paper is intended to support a potentially large amount of viewers, the use of a cloud-computing infrastructure is considered in order to handle the requests from client applications.
+
+\subsection{Web Application}
+\label{ssub_web_application}
+
+The web applications include the business logic, models and controllers that are necessary for the proper functioning of the second screen client application.
+The content that is presented in the client can be requested through HTTP requests to web services designed for this purpose.
+In addition, the data captured by second screen clients in the iTV environment can also be sent to the CP by using HTTP requests.
+
+The HTTP requests handling is carried out by using the front controller architectural design pattern~\cite{Buschmann2007}.
+It is employed to centralize requests processing in a single component, which facilitates the design and implementation of services for different second screen clients.
+As a result, client applications can access the same service even when they are developed for different operating systems in different programming languages.
+
+\subsection{iTV Environment}
+
+The iTV environment consists of an iTV terminal, which can be an set-top box or some iTV terminal with some sort of communication technology e.g., WiFi or Bluetooth.
+This environment has some services to facilitate the communication with second screen devices.
+The first one is the Presence Detector Service (PDS), which is in charge of discovering nearby second screen devices.
+The PDS is connected with another service known as Profile Handler Service (PHS), which aims to perform profile management at each moment when new second screen devices are found in the iTV environment.
+
+% describe a device to device communication
+
+\subsection{Second Screen Devices}
+
+The second screen devices are clients that communicate with the web application in order to provide interactive capabilities with the content being watched in the TV terminal.
+The communication between second screen clients and the web application is made through HTTP requests to the web services.
+Basically, there are two types of communication between clients and the web application: retrieval the content from the web server and transmission of the captured interactions to the web server.
+Since these types communication occurs between client and a web server, Internet connection is required to accomplish the requests.
+
+The retrieval of content from the web server happens when the client application performs a specially designed HTTP request aiming to obtain the content from a specified resource.
+The content obtained from this request is then shown by the client application on the second screen device.
+
+The interactions captured from the viewers are transmitted to the web server by using the same principle of the retrieval of content.
+A HTTP request is sent to the web service.
+However, in this case, the request is designed to submit the data to a specified resource in order to be processed by the web application.
+
+\section{Secondary Screen Implementation}
+
+%The architecture described in this paper combines several concepts and technologies, e.g.
+second screen, interactive television, mobile applications, device to device communication and web technologies.
+An implementation of the architecture described in this paper was developed using the Java programming language in both sides of the architecture.
+However, the proposed approach does not require uniquely the Java programming language and the its use does not subtract the generality of the architecture.
+Indeed, the architecture can be implemented in other languages provided that they are compatible with the development platforms and with the concepts described in this paper.
+
+\subsection{Web Services API Design and Implementation}
+
+A open source web framework~\cite{Walls2011} was used to develop the server side of the architecture, i.e., the web application.
+This framework facilitates the development of the front controller pattern in web-based application.
+Thus, all the requests are tunneled through a single entry point, which is an efficient mechanism to implement an API to access the web services.
+This mechanism not only routes and dispatches the requests to appropriate handlers, but also exposes a flexible and scalable API structure, which can be consumed by second screen devices using different programming languages.
+
+\subsection{Web Service Consumption}
+
+\subsection{iTV and Secondary Screen Devices Communication}
+
+\subsection{Experiment}
+
+ - Content rating systems
+
 \section{Conclusions}
@@
 
-An implementation of the architecture that is described in this paper a commercial cloud computing platform, which plays the role of the application hosting infrastructure.
+An implementation of the architecture that is described in this paper is under development in a commercial cloud computing platform, which plays the role of the application hosting infrastructure.
 A web application that consists of a content rating system was developed to store  viewers' evaluations toward audio visual content, and a secondary screen client is under development.

### 2014-09-08 src/paper.tex
Algumas modificações na seção de design dos web services
Web services design and implementation finished

@@
 
-\section{Secondary Screen Implementation}
+\section{Implementation of the Secondary Screen Architecture}
 
@@
 
-A open source web framework~\cite{Walls2011} was used to develop the server side of the architecture, i.e., the web application.
+The second screen devices and the iTV terminals need to communicate with the web application in order to send viewer's interactions and retrieve content from the server.
+As defined in the Section~\ref{ssub_web_application}, this communication occurs by means of HTTP requests to the web services of the web application.
+Thus, an API containing the Uniform Resource Identifiers (URIs) and the HTTP methods available for each URI is needed.
+
+\begin{table}
+	\caption{Strategies to perform HTTP requests by combining URIs and HTTP methods.}
+	\label{table_web_serv_design}
+	\centering
+	\begin{tabular}{|m{1.5cm}|>{\centering}m{1cm}|m{5cm}|}
+	\hline
+	\textbf{URI}&\textbf{HTTP method}&\textbf{Description of the operation}\tabularnewline
+	\hline
+	/resources & GET & Retrieves a collection of resources of the same type that is specified in the URI. \tabularnewline
+	\hline
+	/resources/\{id\} & GET & Retrieves a single resource identified by a given \emph{\{id\}}.\tabularnewline
+	\hline
+	/resources & POST & Creates a new resource of the same type that is specified in the URI.\tabularnewline
+	\hline
+	/resources/\{id\} & PUT & Updates a resource identified by a given \emph{\{id\}}.\tabularnewline
+	\hline
+	/resources/\{id\} & DELETE & Deletes a resource identified by a given \emph{\{id\}}.
+	\tabularnewline
+	\hline
+	\end{tabular}
+\end{table}
+
+Table~\ref{table_web_serv_design} describes how the web services API are designed by combining URIs and HTTP methods.
+The URI specifies the resource of interest and the HTTP method indicates the desired action to be performed on the resource.
+Since each HTTP method is intended for a different purpose, it is necessary to identify what actions apply to each resource.
+
+A open source web framework~\cite{Walls2011} was used to develop the server side of the architecture, i.e., the web services.
 This framework facilitates the development of the front controller pattern in web-based application.
@@
 \subsection{Web Service Consumption}
+
+Since there is a well-defined web services API to access resources in the web application, it is possible to send HTTP requests from second screen devices and iTV terminals.
+A base URI which contains the domain of the web application (e.g., \emph{http://example.com}) is used as part of the API.
+Thus, the implementation of the HTTP requests in the client devices must use the entire URI, including the resource of interest.
+
+The effect of the HTTP request depends on the type of HTTP method that is used, which typically can be GET, POST, PUT, or DELETE.
+Requests using GET must only retrieve data from the server.
+Depending on the URI, a request using GET should retrieve either a single element or a collection of elements of the specified resource.
+Requests using POST demand that the server accept a new resource identified by the URI.
+Requests using PUT modify a resource identified by the URI.
+The requests using the DELETE method are intended to delete a resource specified by the URI.
+
+The proper combination of URIs and HTTP methods is used to manipulate resources in HTTP requests.
+It is possible to implement CRUD (create, read, update and delete) operations that are intended for communication between the second screen devices and the web application.
+As a result, the client devices can consume the web services and perform the desired actions on the resources.As an example, a request can be used to send viewers' interactions to the web server, or to retrieve content that is personalized according past interactions of the viewer.
 

### 2014-09-08 src/paper.tex
WIP: Experiment

@@
 
-A open source web framework~\cite{Walls2011} was used to develop the server side of the architecture, i.e., the web services.
+A open source web framework~\cite{Wheeler2013} was used to develop the server side of the architecture, i.e., the web services.
 This framework facilitates the development of the front controller pattern in web-based application.
@@
 
- - Content rating systems
+To motivate the need for using this architecture in second screen-based iTV environments, a scenario that enables the viewer to evaluate TV programs was developed.
+This scenario uses mobile devices (second screen devices) to present content related to what is being watched on TV.
+The content is shown in a native application developed for a commercial operating system and this second screen application communicates with a web server through Internet.
+
+The viewers' interactions captured in the application are sent
 

### 2014-09-09 src/paper.tex
WIP: descrição das requisições HTTP

@@
 	\hline
-	\textbf{URI}&\textbf{HTTP method}&\textbf{Description of the operation}\tabularnewline
+	\textbf{URI}&\textbf{HTTP method}&\textbf{Description of the HTTP request}\tabularnewline
 	\hline
@@
 Depending on the URI, a request using GET should retrieve either a single element or a collection of elements of the specified resource.
+Responses bodies for requests using GET methods typically are in JSON format.
+However, it is also possible to provide other Internet media types, e.g., XML or Atom.
 Requests using POST demand that the server accept a new resource identified by the URI.
@@
 To motivate the need for using this architecture in second screen-based iTV environments, a scenario that enables the viewer to evaluate TV programs was developed.
-This scenario uses mobile devices (second screen devices) to present content related to what is being watched on TV.
-The content is shown in a native application developed for a commercial operating system and this second screen application communicates with a web server through Internet.
+This scenario uses mobile devices (second screen devices) to present content related to what is being watched on TV and capture viewers' interactions.
+To develop this scenario it was used a commercial cloud-computing environment intended for emulating the CP infrastructure and a mobile device with a commercial operating system.
 
-The viewers' interactions captured in the application are sent
+The content that is presented in the mobile device is retrieved from a web server through Internet and is shown in an application, which was specially designed for second screen purposes.
+In this application, every time a viewer evaluates a TV program, the data obtained from this interaction is captured and sent to the web application through a HTTP request.
+This request carries data such as \emph{viewerId} (id of the viewer), \emph{programId} (id of the TV program), and \emph{rating} (a value from 1 to 5).
+
+It is important to note that every type of interaction can be captured, e.g., viewers' navigation between TV programs, amount of time spent viewing information about a TV program on the secondary screen, time stamp of the interaction, and etc.
+However, it is necessary to attach triggers to the application so that every time an interaction occurs, the data obtained from this specific interaction is sent to the web application in a HTTP request.
+
+The following URIs and HTTP methods describe some of the HTTP requests for viewers resources that were developed in this experiment:
+
+\begin{itemize}
+	\item \emph{/viewers} - POST: create a new viewer, as identified by the resource that is present in the URI.
+The parameters that are carried in this type of request are \emph{viewerId}, \emph{username}, and \emph{password}.
+	\item \emph{/viewers/\{viewerId\}} - GET: retrieves a specific viewer identified by \{viewerId\} in the URI.
+This request retrieves information about the viewer, which in this case are only \emph{viewerId} and \emph{username}, since is not safe to include a \emph{password} in the response body.
+	\item \emph{/viewers/\{viewerId\}/evaluations} - GET: Retrieve a collection of evaluations that were performed by a viewer identified by \{viewerId\}.
+Each evaluation contains a \emph{programId}, \emph{viewerId}, and a \emph{rating}.
+\end{itemize}
+
+HTTP requests for TV programs are identified as follows:
+
+\begin{itemize}
+	\item 
+\end{itemize}
 

### 2014-09-09 src/paper.tex
WIP: Experiment

@@
 
+The web services API consists in handling three types of resources: viewers, TV programs and programs evaluations.
 The following URIs and HTTP methods describe some of the HTTP requests for viewers resources that were developed in this experiment:
@@
 \begin{itemize}
-	\item \emph{/viewers} - POST: create a new viewer, as identified by the resource that is present in the URI.
+	\item \emph{/viewers} - POST: creates a new viewer, as identified by the resource that is present in the URI.
 The parameters that are carried in this type of request are \emph{viewerId}, \emph{username}, and \emph{password}.
@@
 This request retrieves information about the viewer, which in this case are only \emph{viewerId} and \emph{username}, since is not safe to include a \emph{password} in the response body.
-	\item \emph{/viewers/\{viewerId\}/evaluations} - GET: Retrieve a collection of evaluations that were performed by a viewer identified by \{viewerId\}.
+	\item \emph{/viewers/\{viewerId\}/evaluations} - GET: retrieves a collection of evaluations that were performed by a viewer identified by \{viewerId\}.
 Each evaluation contains a \emph{programId}, \emph{viewerId}, and a \emph{rating}.
@@
 HTTP requests for TV programs are identified as follows:
+
+\begin{itemize}
+	\item \emph{/programs} - POST: creates a new TV program, so that it is available to be presented in the second screen clients.
+In this experiment, an on-line videos platform was used provide videos to second screen clients.
+	\item \emph{/programs} - GET: retrieves a collection of TV programs that are available on the CP.
+A TV program representation contains a \emph{programId}, a \emph{videoCode} (id of the video on the videos platform), and \emph{title} (a title for the video representing the program).
+	\item \emph{/program/\{programId\}} - GET: retrieves a single program identified by \emph{programId}.
+\end{itemize}
+
+Resources related to the evaluations performed by viewers are handled by the following HTTP requests:
 

### 2014-09-10 src/paper.tex
WIP: Experiments - mostrar figura com código gerado

@@
 A TV program representation contains a \emph{programId}, a \emph{videoCode} (id of the video on the videos platform), and \emph{title} (a title for the video representing the program).
-	\item \emph{/program/\{programId\}} - GET: retrieves a single program identified by \emph{programId}.
+	\item \emph{/program/\{programId\}} - GET: retrieves a single program identified by \emph{\{programId\}}.
 \end{itemize}
@@
 \begin{itemize}
-	\item 
+	\item \emph{/evaluations} - POST: creates a new evaluation. This HTTP request requires three attributes: \emph{viewerId}, \emph{programId}, and \emph{rating}.
+	\item \emph{evaluations/viewers/\{viewerId\}/programs/\{programId\}} - GET: retrieves an evaluation of a viewer identified by \emph{\{viewerId\}} on a given TV program identified by \emph{\{programId\}}.
 \end{itemize}
+
+Some of the HTTP requests
 

### 2014-09-10 src/paper.tex
WIP: response in JSON format

@@
 
-Some of the HTTP requests
+Some of the HTTP requests were created only to add content to the web application, so that the client devices could access this content.
+The HTTP requests that use a GET method typically are used in the client devices to retrieve content from the web application.
+In this experiment, it was employed the JSON format to
+
+\begin{figure}
+\begin{verbatim}
+{
+    id: 5722646637445120,
+    viewer: {
+        id: 5629499534213120,
+        username: "user1"
+    },
+    program: {
+        id: 5649391675244544,
+        videoCode: "ZuQ4obQAN8",
+        title: "A Program Title"
+    },
+    rating: 4
+}
+\end{verbatim}
+\caption{Example of response to an HTTP request in JSON format.}
+\label{fig_response_json}
+\end{figure}
 

### 2014-09-11 src/paper.tex
Figure about JSON finished

@@
 	\item \emph{/evaluations} - POST: creates a new evaluation. This HTTP request requires three attributes: \emph{viewerId}, \emph{programId}, and \emph{rating}.
-	\item \emph{evaluations/viewers/\{viewerId\}/programs/\{programId\}} - GET: retrieves an evaluation of a viewer identified by \emph{\{viewerId\}} on a given TV program identified by \emph{\{programId\}}.
+	\item \emph{/evaluations/viewers/\{viewerId\}/programs/\{programId\}} - GET: retrieves an evaluation of a viewer identified by \emph{\{viewerId\}} on a given TV program identified by \emph{\{programId\}}.
 \end{itemize}
@@
 Some of the HTTP requests were created only to add content to the web application, so that the client devices could access this content.
-The HTTP requests that use a GET method typically are used in the client devices to retrieve content from the web application.
-In this experiment, it was employed the JSON format to
+The client devices typically use GET method in HTTP requests to retrieve content from the web application.
+In this experiment, the responses to the HTTP requests were provided in the JSON format.
+\figurename{~\ref{fig_response_json}} presents a JSON representation of a resource obtained with an HTTP request to \emph{/evaluations/viewers/\{viewerId\}/programs/\{programId\}} using a GET method.
+The implementation of this request searches for a resource of the type evaluation that has a \emph{\{viewerId\}} equals to 5629499534213120, and a \emph{\{programId\}} equals to 5649391675244544.
 
@@
 \end{verbatim}
-\caption{Example of response to an HTTP request in JSON format.}
+\caption{Response to an HTTP request using GET method in JSON format. The code shown in this figure is a JSON representation of a resource of the type evaluation.}
 \label{fig_response_json}

### 2014-09-11 src/paper.tex
WIP: Devices communication

@@
 \title{A Second Screen Architecture for Capturing Viewers' Interaction Data in TV Environments}
-\author{Ricardo~Erikson~V.~de~S.~Rosa, 
-	Vicente~Ferreira~de~Lucena~Jr.,\IEEEmembership{Member,IEEE,}
-	and~Lucas~Carvalho~Cordeiro
+\author{Ricardo~Erikson~V.~de~S.~Rosa,\IEEEmembership{Student Member, IEEE,} 
+	and Vicente~Ferreira~de~Lucena~Jr.,\IEEEmembership{Senior Member,IEEE}
+	% and~Lucas~Carvalho~Cordeiro
 	%
@@
 \thanks{Vicente Ferreira de Lucena Jr. is with the PPGEE, PPGI, and CETELI - Electronics and Information Technology R\&D Center at UFAM, Manaus, Amazonas, Brazil (e-mail: vicente@ufam.edu.br)}%
-\thanks{Lucas Carvalho Cordeiro is with the PPGEE-UFAM, and CETELI - Electronics and Information Technology R\&D Center, Manaus, Amazonas, Brazil (e-mail:lucascordeiro@ufam.edu.br)}}
+% \thanks{Lucas Carvalho Cordeiro is with the PPGEE-UFAM, and CETELI - Electronics and Information Technology R\&D Center, Manaus, Amazonas, Brazil (e-mail:lucascordeiro@ufam.edu.br)}
+}
 
@@
 
-\subsection{Experiment}
+An important feature of the architecture described in this paper is the possibility of performing communication between iTV and secondary screen devices.
+This type of communication enable the discovery of devices that are physically nearby, i.e., devices on the same local network or not.
+Since people are typically carry their phones with them, it is possible to manage who is present in the environment at a given moment.
+
+Industry provided several means to provide communication among devices.
+
+\section{Experiment: Evaluation of TV Programs using Second Screen}
+
+% Enfatizar a interação de avaliar um programa
 

### 2014-09-14 src/paper.tex
WIP: Reviewing introduction

@@
 Rosa is with the Graduate Program in Electrical Engineering, Federal University of Minas Gerais, Av.
-Antônio Carlos 6627, 31270-901, Belo Horizonte, MG, Brazil(e-mail: ricardoerikson@ufmg.br)}%
+Ant\^{o}nio Carlos 6627, 31270-901, Belo Horizonte, MG, Brazil(e-mail: ricardoerikson@ufmg.br)}%
 \thanks{Vicente Ferreira de Lucena Jr. is with the PPGEE, PPGI, and CETELI - Electronics and Information Technology R\&D Center at UFAM, Manaus, Amazonas, Brazil (e-mail: vicente@ufam.edu.br)}%
@@
 
-TV watching is essentially a social activity, where groups of people with a common interest share the same space and TV set for entertainment or information~\cite{Masthoff2004}.
-After many advances in technologies for interactive and digital TV, it is possible to develop high-level applications to enrich the watching experience.
-Thus, the interaction between viewers and TV became more elaborated than simply change channels and adjust sound volume.
-In this scenario, .
+TV watching is essentially a social activity, where groups of people gather around with the common interest of sharing the same space and TV set for entertainment or information~\cite{Masthoff2004}.
+After many advances in technologies for interactive and digital TV, it is possible to develop high-level content to enrich the watching experience.
+Thus, the traditional TV, where viewers simply change channels and adjust sound volume, turned into a more elaborated TV model.
+In this new model, viewers can access Internet and interact with applications and services while watching TV programs.
 
-The conventional remote control (RC) is typically seen as the standard medium of interaction among the viewers and their TVs.
-However, this medium presents two notable problems: (1) to precisely and automatically identify the viewers who have possession of the RC; and (2) to capture contextualized viewers' interactions.
-As a result, the entire group of viewers (e.g., a family) is treated as a single viewer and the tastes that were inferred from the users in charge of the RC can be imposed on the tastes of the others.
+Capturing the interactions of millions of viewers who are using interactive iTV applications and services can provide valuable information for content providers (CPs)~\cite{Teixeira2010}.
+However, it lacks a proper device to enable the capture of these interactions.
+The conventional remote control (RC), which is typically seen as the standard medium of interaction among the viewers and their TVs, became very limited for the iTV capabilities~\cite{Jeong2014}.
+In addition, this medium presents two notable problems: (1) to precisely and automatically identify the viewers who have possession of the RC; and (2) to capture contextualized viewers' interactions.
+As a result, it is difficult to properly identify the viewers in the environment, and the entire group (e.g., a family) is treated as a single viewer that is interacting with the iTV terminal.
 
-Given the limitations of the conventional RC, a novel secondary screen architecture to identify viewers who are interacting with interactive TV (iTV) services is presented in this paper.
-Personal devices (e.g., smartphones and tablets) are used as secondary screens with the aim of capturing personal and contextualized interactions from each active viewer in iTV environments.
-By using proper machine learning algorithms, the personal and contextualized data can reveal important information about viewers' tastes~\cite{Kim2012,Shin2009}.
+Given the limitations of the conventional RC, a novel second screen architecture to identify viewers who are interacting with iTV applications and services is presented in this paper.
+Personal devices (e.g., smartphones and tablets) are used as second screen devices with the aim of capturing personal and contextualized interactions from active viewers in iTV environments.
+As a result, viewers that interact with an iTV terminal can be properly identified while using a second screen device, which enables to precisely capture their interactions.
+
+An important application... By using proper machine learning algorithms, the personal and contextualized data can reveal important information about viewers' tastes~\cite{Kim2012,Shin2009}.
+
+As an example, the CPs can obtain the user ratings given to TV programs, amount of viewers and user profiles.
+By using proper machine learning algorithms, CPs can infer viewer tastes from audience data in order to deliver personalized content~\cite{Kim2012,Shin2009}.
+However, identifying viewers and capturing their individual interactions is still a challenging activity in the traditional TV environment, which consists of the TV and viewers interacting through a remote control (RC).
+
+% In this research field, a typical approach is to use personal devices (e.g., smartphones and tablets) as secondary screens~\cite{Courtois2012}, since they are almost ubiquitous and present good graphical capabilities.
+The connectivity capabilities of smartphones were used in~\cite{Cabarcos2011}, where the Bluetooth technology was used to detect nearby users.
+With respect to interaction capture, an approach described in~\cite{Teixeira2010} shown an architecture to capture contextualized viewers interactions through a conventional RC.
+In this paper, we present an architecture that facilitates the identification of viewers, which is useful to contextualize the interactions.
+In addition, the identification of viewers facilitates the capture of their individual interactions.
 
@@
 
-Viewers can use their secondary screen device to interact with TV by evaluating, commenting, sharing, or even recommending TV content to other viewers.
-The data related to those actions are naturally provided by the viewers, as part of the interaction, when using interactive applications (developed by the CP) installed in their secondary screen devices.
+Viewers can use their second screen device to interact with TV by evaluating, commenting, sharing, or even recommending TV content to other viewers.
+The data related to those actions are naturally provided by the viewers, as part of the interaction, when using interactive applications (developed by the CP) installed in their second screen devices.
 The interactions and the contextual data are represented by the audience data, which is sent to CP infrastructure through Internet.
@@
 
-The Data Writer service in the application hosting infrastructure consists basically of web services that provide interoperability among secondary screen devices, iTV devices and the CP infrastructure.
+The Data Writer service in the application hosting infrastructure consists basically of web services that provide interoperability among second screen devices, iTV devices and the CP infrastructure.
 The main advantage of this approach is the independence of operating system and programming languages used in the personal devices.
@@
 This information can be used to provide personalized content to viewers based on the current context.
-Thus, the personalized content can be delivered via Internet to the secondary screen device or even to the TV, depending on the context.
+Thus, the personalized content can be delivered via Internet to the second screen device or even to the TV, depending on the context.
 
@@
 
-\section{Implementation of the Secondary Screen Architecture}
+\subsection{iTV and Second Screen Devices Communication}
+
+An important feature of the architecture described in this paper is the possibility of performing communication between iTV and second screen devices.
+This type of communication enable the discovery of devices that are physically nearby, i.e., devices on the same local network or not~\cite{Hellman2013}.
+Since people are typically carrying their phones with them, it is possible to manage who is present in the environment at a given moment.
+
+The mobile devices industry provided several means to enable communication among devices (e.g., mobile devices and iTV terminals), which usually include network service discovery and peer-to-peer (P2P) connections.
+A useful way to find out nearby devices that are able to communicate is by using network service discovery.
+It consists of an application that can broadcast its name and connection information on the local network and identifies other devices running a service with which it can communicate.
+
+An interesting P2P communication technology that is being adopted by many mobile devices manufactures is the Wi-Fi Direct~\cite{Hellman2013}.
+This technology enables Wi-Fi devices to connect directly without requiring an intermediate dedicated access point.
+This type of communication is based on Wi-Fi and is very similar to Bluetooth, except that it uses a higher speed and a range beyond the Bluetooth capabilities.
+
+% Colocar um parágrafo falando da importância de saber quem está por perto.
+
+\section{Implementation of the Second Screen Architecture}
 
@@
 
-\subsection{iTV and Secondary Screen Devices Communication}
-
-An important feature of the architecture described in this paper is the possibility of performing communication between iTV and secondary screen devices.
-This type of communication enable the discovery of devices that are physically nearby, i.e., devices on the same local network or not.
-Since people are typically carry their phones with them, it is possible to manage who is present in the environment at a given moment.
-
-Industry provided several means to provide communication among devices.
-
 \section{Experiment: Evaluation of TV Programs using Second Screen}
@@
 
-It is important to note that every type of interaction can be captured, e.g., viewers' navigation between TV programs, amount of time spent viewing information about a TV program on the secondary screen, time stamp of the interaction, and etc.
+It is important to note that every type of interaction can be captured, e.g., viewers' navigation between TV programs, amount of time spent viewing information about a TV program on the second screen, time stamp of the interaction, and etc.
 However, it is necessary to attach triggers to the application so that every time an interaction occurs, the data obtained from this specific interaction is sent to the web application in a HTTP request.
+
+\subsection{Implementation of the Web Application}
 
@@
 
+\subsection{Implementation of the Second Screen Application}
+
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
+tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
 \section{Conclusions}
 
-The use of personal devices as secondary screens arises as a new trend in TV environments.
-This paper presents an architecture to accurately capture audience data by using secondary screen devices and iTV devices.
-This architecture take advantage of the ubiquity and interactive capabilities of secondary screen devices to identify viewers and provide feedback for CPs through viewers' interactions and contextual information.
+The use of personal devices as second screens arises as a new trend in TV environments.
+This paper presents an architecture to accurately capture audience data by using second screen devices and iTV devices.
+This architecture take advantage of the ubiquity and interactive capabilities of second screen devices to identify viewers and provide feedback for CPs through viewers' interactions and contextual information.
 The captured data relates to a single viewer in a given context.
@@
 An implementation of the architecture that is described in this paper is under development in a commercial cloud computing platform, which plays the role of the application hosting infrastructure.
-A web application that consists of a content rating system was developed to store  viewers' evaluations toward audio visual content, and a secondary screen client is under development.
-Some RESTful web services were developed to provide interoperability among the secondary screen devices and the web application.
-Thus, the interactions that are captured during the use of the secondary screen application are sent to the CP by means of HTTP requests to RESTful web services.
-Although the secondary screen application is not fully finished yet, some specific features have been tested to capture viewers' interactions as they occur.
+A web application that consists of a content rating system was developed to store  viewers' evaluations toward audio visual content, and a second screen client is under development.
+Some RESTful web services were developed to provide interoperability among the second screen devices and the web application.
+Thus, the interactions that are captured during the use of the second screen application are sent to the CP by means of HTTP requests to RESTful web services.
+Although the second screen application is not fully finished yet, some specific features have been tested to capture viewers' interactions as they occur.
 
@@
 
+% Nowadays, viewers rarely just sit passively and watch TV.
+In many situations, they use their personal devices as a secondary screen to perform activities related to TV watching, which end up enriching and immersing the user into the TV experience.
+
 \bibliographystyle{IEEEtran}

### 2014-09-15 src/paper.tex
WIP: Background

@@
 Given the limitations of the conventional RC, a novel second screen architecture to identify viewers who are interacting with iTV applications and services is presented in this paper.
-Personal devices (e.g., smartphones and tablets) are used as second screen devices with the aim of capturing personal and contextualized interactions from active viewers in iTV environments.
-As a result, viewers that interact with an iTV terminal can be properly identified while using a second screen device, which enables to precisely capture their interactions.
+Personal devices (e.g., smartphones and tablets) are used as second screen devices with the aim of capturing personal and contextualized interactions from viewers in an iTV environment.
+Furthermore, the usage of second screen devices offers other advantage.
+It opens up possibilities to provide viewers with extra information without disturbing the collective experience.
+For example, each viewer can independently browse through different information about programs and actors in their own devices.
+As a result, viewers that interact with an iTV terminal by using a second screen device can be properly identified, which enables to precisely capture their interactions.
 
-An important application... By using proper machine learning algorithms, the personal and contextualized data can reveal important information about viewers' tastes~\cite{Kim2012,Shin2009}.
+To illustrate the second screen architecture described in this paper, it was developed a scenario where viewers can use a second screen device to evaluate TV programs.
+In this scenario, it is possible to browse through TV programs and view extra information about a programs of interest.
+Viewers' interactions are captured during the use of the second screen device and the data is sent to a remote web application.
+The applications that were developed in this scenario were tested in commercial platforms that are commonly used by developers and manufacturers as a production environment for both free and paid applications.
 
-As an example, the CPs can obtain the user ratings given to TV programs, amount of viewers and user profiles.
-By using proper machine learning algorithms, CPs can infer viewer tastes from audience data in order to deliver personalized content~\cite{Kim2012,Shin2009}.
-However, identifying viewers and capturing their individual interactions is still a challenging activity in the traditional TV environment, which consists of the TV and viewers interacting through a remote control (RC).
+Important applications for the data generated during the viewers' interactions with second screen devices are in the field of content personalization.
+By using proper machine learning algorithms, the data can reveal important information about viewers tastes~\cite{Kim2012,Shin2009}.
+As a result, CPs can deliver to the right viewer at the right time the right content, which can include targeted advertising, and content recommendation ~\cite{DosSantos2013}.
 
-% In this research field, a typical approach is to use personal devices (e.g., smartphones and tablets) as secondary screens~\cite{Courtois2012}, since they are almost ubiquitous and present good graphical capabilities.
-The connectivity capabilities of smartphones were used in~\cite{Cabarcos2011}, where the Bluetooth technology was used to detect nearby users.
+% The contribution of this paper focuses on the consumption stage, and not so much on the different algorithms for content recommendation ... // Do not focus on content unrelated activity.
+
+The rest of this paper is organized as follows: first, Section~\ref{sec_background} presents a background of the concepts related to the proposed architecture.
+In Section~\ref{sec_overview} an overview of the proposed architecture is described.
+Section~\ref{sec_architecture} presents the architecture to capture interactions data from viewers, and Section~\ref{sec_implementation} describes how the architecture was implemented.
+Finally, Section~\ref{sec_experiment} contains 	an implementation scenario, and Section~\ref{sec_conclusions} provides the conclusions of this work.
+
+\section{Background}
+\label{sec_background}
+
+% The connectivity capabilities of smartphones were used in~\cite{Cabarcos2011}, where the Bluetooth technology was used to detect nearby users.
 With respect to interaction capture, an approach described in~\cite{Teixeira2010} shown an architecture to capture contextualized viewers interactions through a conventional RC.
@@
 
-Since personal devices are almost ubiquitous, they stand out as a powerful mechanism to identify viewers by means of wireless communication technologies~\cite{Cabarcos2011}.
+% Since personal devices are almost ubiquitous, they stand out as a powerful mechanism to identify viewers by means of wireless communication technologies~\cite{Cabarcos2011}.
 In the architecture described in this paper, personal devices play an additional role as a medium of interaction since they are easy and intuitive to use in iTV environments, as pointed out in~\cite{Courtois2012}.
@@
 Since the TV environment is shared, the data about nearby viewers is captured to contextualize the interaction with the purpose of helping to infer the behaviors of one viewer toward others ones.
-
-The sections of this paper are arranged as follows: \ldots
-
-\section{Background}
 
@@
 In many situations, they use second screen devices to perform activities related to TV watching.
-In the context of TV, second screen refers to the use of a computing device, which is usually a smartphone or a tablet PC, to enhance the viewing experience.
+In the context of TV, second screen refers to the use of a computing device as an extended remote control in order to enhance the viewing experience~\cite{Cesar2009}.
+A typical approach is to use personal devices (e.g., smartphones and tablet PCs) as second screens, since they are almost ubiquitous and present good graphical capabilities~\cite{Courtois2012}.
 
 Research studies on second screen have been conducted in TV environments since late 1990s.
-In 1996, Robertson~\emph{et~al}.~\cite{Robertson1996} developed a prototype where handheld devices were used to control TV and to provide additional content (e.g., textual information) related to the audio-visual content presented in the TV screen.
-More recent studies show that viewers evaluated positively the use of second screen devices to control the TV and to enrich the TV content~\cite{Cesar2008,Cesar2011,Tsekleves2011}.
-Cesar~\emph{et~al.}~\cite{Cesar2009} contributed identifying and modeling the requirements to implement an interactive TV architecture based on second screen devices.
+In 1996, Robertson~\emph{et~al}.~\cite{Robertson1996} developed a prototype where handheld devices were used to control TV and to provide additional content (e.g., textual information) related to the audio-visual content presented on TV screen.
+The handheld device can be used both as a stand-alone service and in conjunction with TV.
+In the latest case, the TV only responds to handheld commands and presents audio-visual content.
+
+More recent studies show that viewers evaluated positively the use of second screen devices to control and to enrich the TV content~\cite{Cesar2008,Cesar2011,Tsekleves2011}.
+Cesar~\emph{et~al.}~\cite{Cesar2009} contributed identifying and modeling the requirements to implement an iTV architecture based on second screen devices.
+Through business analysis
+
+an working implementation of this architecture, the authors were able to conclude that second screen devices are considered as a value-added service
 
@@
 \section{Overview of the Second Screen Architecture}
+\label{sec_overview}
 
@@
 
-\section{Second Screen Architecture}
+\section{Architecture to Capture Viewers' Interactions}
+\label{sec_architecture}
 
@@
 \section{Implementation of the Second Screen Architecture}
+\label{sec_implementation}
 
@@
 \section{Experiment: Evaluation of TV Programs using Second Screen}
+\label{sec_experiment}
 
@@
 \section{Conclusions}
+\label{sec_conclusions}
 

### 2014-09-16 src/paper.tex
Primeira versão do artigo enviada para o professor

@@
 
-% The connectivity capabilities of smartphones were used in~\cite{Cabarcos2011}, where the Bluetooth technology was used to detect nearby users.
-With respect to interaction capture, an approach described in~\cite{Teixeira2010} shown an architecture to capture contextualized viewers interactions through a conventional RC.
-In this paper, we present an architecture that facilitates the identification of viewers, which is useful to contextualize the interactions.
-In addition, the identification of viewers facilitates the capture of their individual interactions.
-
-% Since personal devices are almost ubiquitous, they stand out as a powerful mechanism to identify viewers by means of wireless communication technologies~\cite{Cabarcos2011}.
-In the architecture described in this paper, personal devices play an additional role as a medium of interaction since they are easy and intuitive to use in iTV environments, as pointed out in~\cite{Courtois2012}.
-An approach described in~\cite{Teixeira2010} has shown an architecture to capture viewers' interactions through a conventional RC.
-The architecture described in this paper goes further and captures individual and contextualized data.
-Since the TV environment is shared, the data about nearby viewers is captured to contextualize the interaction with the purpose of helping to infer the behaviors of one viewer toward others ones.
-
 The key concepts and some relevant related works that underpin the proposed architecture are described in this section.
@@
 In the context of TV, second screen refers to the use of a computing device as an extended remote control in order to enhance the viewing experience~\cite{Cesar2009}.
-A typical approach is to use personal devices (e.g., smartphones and tablet PCs) as second screens, since they are almost ubiquitous and present good graphical capabilities~\cite{Courtois2012}.
+A typical approach is to use personal devices (e.g., smartphones and tablet PCs) as second screens, since they easy and intuitive to use in iTV environments~\cite{Courtois2012}.
 
@@
 Cesar~\emph{et~al.}~\cite{Cesar2009} contributed identifying and modeling the requirements to implement an iTV architecture based on second screen devices.
-Through business analysis
-
-an working implementation of this architecture, the authors were able to conclude that second screen devices are considered as a value-added service
+A business analysis revealed that second screen devices are considered as a value-added service for iTV.
+The same study concluded that users that tested the working implementation proved to be attracted by the possibility of using a second screen device in the home environment.
 
 A review of the related works shown that the usage of personal devices as second screen devices has two main advantages.
-Firstly, the visual elements (e.g., an EPG guide) that are commonly shown on the TV screen can be presented on the second screen device in such a way that the viewing experience of other viewers is not affected.
+Firstly, the visual elements (e.g., an EPG guide) that are commonly shown on TV screen can be presented on the second screen device in such a way that the viewing experience of other viewers is not affected.
 Secondly, the second screen device can enhance the capabilities of a conventional remote control by providing more sophisticated interactions, e.g., gesture and voice commands.
@@
 The remote control events are recorded in XML format using elements to describe contextual information, implicit and explicit interactions.
-Basílio~\emph{et~al.}~\cite{Basilio2013} described two methods for capturing viewers interactions: (1) one using a middleware extension, which adds new functionalities to an existing standard; and (2) another one using applications that are added to the broadcaster data stream, which requires no changes in the middleware.
+Bas\'{i}lio~\emph{et~al.}~\cite{Basilio2013} described two methods for capturing viewers interactions: (1) one using a middleware extension, which adds new functionalities to an existing standard; and (2) another one using applications that are added to the broadcaster data stream, which requires no changes in the middleware.
 
-Another research study conduced by Santos and Bianchini~\cite{DosSantos2013}, described a advertising system for interactive TV, which uses viewers interactions in order to generate targeted advertising.
-Basically, the described system implicitly captures metadata (e.g., the category of a TV program) information when interactions take place.
+Another research study conduced by Santos and Bianchini~\cite{DosSantos2013}, described a advertising system for interactive TV, which uses viewers' interactions to generate targeted advertising.
+Basically, the methodology that was employed in this work is to capture implicit and explicit interactions from viewers during the use of the proposed system.
+When browsing through TV programs, metadata information of the programs is stored and linked to the viewer profile in order to generate targeted advertising.
 
-The data generated by viewers' interactions can provide valuable information to improve TV personalization, especially when data comes from millions of viewers.
+The data generated by viewers' interactions can provide valuable feedback to improve the quality of content and iTV services.
+This feedback is even more valuable when it comes from millions of viewers.
 For instance, viewers can change quickly the channel when a TV program loses their interest.
@@
 
-Since the TV environment can be shared by many viewers (e.g., a family or a group of friends), audience identification becomes an important feature for interactive TV systems.
-Many modern computational systems consist of identifying viewers in order to provide personalized services.
+Since the TV environment can be shared by many viewers (e.g., a family or a group of friends), audience identification becomes an important feature for iTV systems.
+Many modern computational systems consist of identifying viewers in order to provide personalized content and services.
 Unfortunately, the traditional model of interaction, which uses a conventional remote control, hinders the identification of viewers.
@@
 
-Some research studies have addressed automated technologies for viewers' identification in interactive TV environments.
+Some research studies have addressed automated technologies for viewers' identification in iTV environments.
 Cabarcos~\emph{et~al.}~\cite{Cabarcos2011} described an architecture that uses the connectivity capabilities of smartphones to detect viewers nearby the TV terminal at every moment.
@@
 
-The architecture presented in this paper applies to the capture of audience interactions by means of the use of second screen devices in interactive TV environments.
+% The architecture described in this paper goes further and captures individual and contextualized data.
+Since the TV environment is shared, the data about nearby viewers is captured to contextualize the interaction with the purpose of helping to infer the behaviors of one viewer toward others ones.
+
+The architecture presented in this paper applies to the capture of audience interactions by means of the use of second screen devices in iTV environments.
 Personal devices are used as second screen devices to enrich the interactivity between viewers and TV at the same time that they facilitate the capture of viewers' interactions.
@@
 	\includegraphics[width=3.5in]{img/architecture-overview.pdf}
-	\caption{Architecture to capture audience interactions. The viewers' interactions and the contextual information are captured and stored in the CP infrastructure.}
+	\caption{Architecture to capture audience interactions. The viewers' interactions and the contextual information are captured and stored in the AHI.}
 	\label{fig_architecture}
@@
 An overview of the architecture described in this paper is presented in \figurename~\ref{fig_architecture}.
-This architecture consists of tree subsystems that interact with each other: (1) second screen devices, which are personal devices; an (2) iTV terminal; and (3) the CP infrastructure.
-The CP infrastructure includes an application hosting infrastructure, which comprises the computing resources (e.g., application server, database services and storage services) that are necessary to host web applications.
-The dashed arrows represent the data flowing from the viewer to the CP infrastructure, while dotted arrows denote the flow from CP infrastructure to the viewers.
-The continuous arrows represent the data flow inside the CP infrastructure.
+This architecture consists of tree subsystems that interact with each other: (1) second screen devices, which are personal devices; an (2) iTV terminal; and (3) the Application Hosting Infrastructure (AHI).
+The AHI includes an application hosting infrastructure, which comprises the computing resources (e.g., application server, database services and storage services) that are necessary to store the content, which is accessed by second screen devices and iTV terminals.
+The dashed arrows represent the data flowing from the viewer to the AHI, while dotted arrows denote the flow from AHI to the viewers.
+Since the AHI is usually in a remote location far from viewers' houses, the communication with it is made by performing remote requests to its servers, while the communication between the second screen devices and the iTV terminal is done within the local network through a local synchronization.
 
-A web application is hosted in the application hosting infrastructure.
-This application contains web services that are designed for interoperability between the second screen devices, the iTV terminal and the CP infrastructure.
-A authentication mechanism must be provided so that the viewers can connect their second screen devices and interactive TV terminals (e.g., set-top boxes or smart TVs) with an user account.
+The content in the AHI is made accessible by means of a web application.
+This application contains web services that are designed for interoperability between the second screen devices, the iTV terminal and the AHI.
+A authentication mechanism must be provided so that the viewers can sign in their second screen devices and interactive TV terminals with an user account, which enables the viewer identification.
 Once the authentication is completed, the second screen device can be used to interact with the iTV terminal.
+This method presents an advantage when compared to the method described by Cabarcos~\emph{et~al.}~\cite{Cabarcos2011}.
+Since the viewer is associated to a user account an not a physical device, it is only necessary to sign in on the new device to use it as a second screen.
 
-A presence service detects physically nearby second screen devices that have the services necessary to establish a communication with an iTV terminal.
+% A presence service detects 
+In this architecture, the physically nearby second screen devices that have the services necessary to establish a communication with an iTV terminal are detected.
 This feature is particularly appropriate for the TV environment, once second screen devices are usually connected with an iTV terminal via WiFi.
-The devices are synchronized in order to update the contextual data about the viewer profile, amount of people using a second screen devices in the room and content being watched.
+The devices are synchronized in order to send commands and update the contextual data about the iTV environment ,e.g., viewer profile, amount of people using a second screen devices in the room and content being watched.
 This contextual data can be used to learn how a viewer behaves towards other viewers.
@@
 The second screen devices and iTV terminals have client applications that consume the web services in order to present the content to the viewers.
-These client applications also use the web services as an interface to send interaction data to the CP.
-
-% descrever a possibilidade de autenticação em outra parte do texto: A possible way to do so is to generate a registration code for each equipment
-
-% commentar sobre a sincronização de contexto entre iTV terminals e second screen devices.
+These client applications also use the web services as an interface to send viewers' interaction data to the AHI.
 
@@
 
-Viewers can use their second screen device to interact with TV by evaluating, commenting, sharing, or even recommending TV content to other viewers.
-The data related to those actions are naturally provided by the viewers, as part of the interaction, when using interactive applications (developed by the CP) installed in their second screen devices.
-The interactions and the contextual data are represented by the audience data, which is sent to CP infrastructure through Internet.
-Then, a Data Writer service stores the audience data related to each particular viewer and the iTV device.
+Viewers can use their second screen devices to interact with TV by evaluating, commenting, sharing, or even recommending TV content to other viewers.
+The data related to those actions are naturally provided by the viewers, as part of the interaction, when using interactive applications installed in their second screen devices.
+The interactions and the contextual data are represented by the audience data, which is sent to AHI through Internet.
+Then, the audience data related to each particular viewer and the iTV device is stored in the AHI.
 
-The Data Writer service in the application hosting infrastructure consists basically of web services that provide interoperability among second screen devices, iTV devices and the CP infrastructure.
-The main advantage of this approach is the independence of operating system and programming languages used in the personal devices.
+\begin{figure}[!t]
+	\centering
+	\includegraphics[width=3.5in]{img/architecture-detailed.pdf}
+	\caption{Architecture to capture audience interactions. The viewers' interactions and the contextual information are captured and stored in the AHI.}
+	\label{fig_arch_detail}
+\end{figure}
+
+The mechanism used to store data in the AHI is the use of HTTP requests.
+\figurename{~\ref{fig_arch_detail}} presents a more detailed illustration of the architecture presented in this paper.
+As shown in the figure, the communication between client devices and AHI is performed by means of HTTP requests.
+The core of these requests consists basically of a Web Services Container (WSC), which have the web services that provide interoperability among second screen devices, iTV devices and the AHI.
+The main advantage of this approach is the independence of operating system and programming languages used in the client devices.
 
@@
 This information can be used to provide personalized content to viewers based on the current context.
-Thus, the personalized content can be delivered via Internet to the second screen device or even to the TV, depending on the context.
+Thus, the personalized content can be delivered over the Internet to the second screen device or even to the TV, depending on the context.
 
-\subsection{Application Hosting Infrastructure}
+\subsection{Application Hosting Infrastructure (AHI)}
 
-The application hosting infrastructure (AHI) consists of a server environment that provides the necessary computing resources to run a web application and make it available to client applications.
+The AHI consists of a server environment that provides the necessary computing resources to run a web application and make it available to client applications.
 This infrastructure can be owned by the CP or it can be hired from a third party service provider.
@@
 
-\subsection{Web Application}
+\subsection{Web Services Container (WSC)}
 \label{ssub_web_application}
 
-The web applications include the business logic, models and controllers that are necessary for the proper functioning of the second screen client application.
+The WSC is part of the web application and includes the business logic, and controllers that are necessary for the proper functioning of the second screen client application.
 The content that is presented in the client can be requested through HTTP requests to web services designed for this purpose.
-In addition, the data captured by second screen clients in the iTV environment can also be sent to the CP by using HTTP requests.
+In addition, the data captured by second screen clients in the iTV environment can also be sent to the AHI by using HTTP requests.
 
@@
 
+Every HTTP request that is made to the web application is handled by the front controller, as can be seen in \figurename{~\ref{fig_arch_detail}}.
+The front controller send this request to a handler mapping, which is responsible for identifying the controller that can handle the request.
+Once the request is completed by the controller, a response is sent to the client that performed the request.
+Typically, this response is given in a machine readable format such as JSON or XML.
+
 \subsection{iTV Environment}
 
-The iTV environment consists of an iTV terminal, which can be an set-top box or some iTV terminal with some sort of communication technology e.g., WiFi or Bluetooth.
+The iTV environment consists of an iTV terminal, which can be a set-top box or an iTV terminal with some sort of communication technology e.g., WiFi or Bluetooth.
 This environment has some services to facilitate the communication with second screen devices.
 The first one is the Presence Detector Service (PDS), which is in charge of discovering nearby second screen devices.
-The PDS is connected with another service known as Profile Handler Service (PHS), which aims to perform profile management at each moment when new second screen devices are found in the iTV environment.
+The PDS is connected with another service known as Profile Handler Service (PHS), which aims at performing profile management at each moment when new second screen devices are found in the iTV environment.
 
@@
 This scenario uses mobile devices (second screen devices) to present content related to what is being watched on TV and capture viewers' interactions.
-To develop this scenario it was used a commercial cloud-computing environment intended for emulating the CP infrastructure and a mobile device with a commercial operating system.
+To develop this scenario it was used a commercial cloud-computing environment intended for emulating the AHI and a mobile device with a commercial operating system.
 
@@
 
-\subsection{Implementation of the Web Application}
+% \subsection{Implementation of the Web Application}
 
@@
 
-\subsection{Implementation of the Second Screen Application}
+% \subsection{Implementation of the Second Screen Application}
 
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
-tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
-quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
-consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
-cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
-proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+% Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
+% tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
+% quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo
+% consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse
+% cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+% proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 
@@
 
-An implementation of the architecture that is described in this paper is under development in a commercial cloud computing platform, which plays the role of the application hosting infrastructure.
-A web application that consists of a content rating system was developed to store  viewers' evaluations toward audio visual content, and a second screen client is under development.
-Some RESTful web services were developed to provide interoperability among the second screen devices and the web application.
-Thus, the interactions that are captured during the use of the second screen application are sent to the CP by means of HTTP requests to RESTful web services.
-Although the second screen application is not fully finished yet, some specific features have been tested to capture viewers' interactions as they occur.
+A scenario of the architecture was implemented under a commercial cloud computing platform, which plays the role of the AHI.
+This scenario consists of a web application which contains a content rating system to store  viewers' evaluations toward audio visual content, and a second screen client.
+Web services to provide interoperability among the second screen devices and the web application were developed in such a way that the interactions that are captured during the use of the second screen application are sent to the CP by means of HTTP requests to web services.
 
-The architecture described in this paper is also technically feasible to implement in Digital TV (DTV) systems and can be particularly interesting for CPs in DTV standards, e.g., the Brazilian (ISDB-TB) and the Japanese(ISDB-T).
+The architecture described in this paper is also technically feasible to implement in Digital TV systems and can be particularly interesting for CPs in DTV standards, e.g., the Brazilian (ISDB-TB) and the Japanese(ISDB-T).
 These standards support the use of mobile devices as part of the system.
